### 2023-02-03 products/manuscript/Manuscript.qmd
Project Part 1

---
title: "Example Manuscript Template for a Data Analysis Project"
subtitle: ""
author: Andreas Handel
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../dataanalysis_template_references.bib
csl: ../apa.csl
---

The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. 
You don't need to have exactly these sections, but the content covering those sections should be addressed.

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

# Summary/Abstract
Globally, *Salmonella* is a significant public health threat. We evaluated the persistence of *Salmonella* on various surface materials (e.g., picnic tables and bird feeders). Our studies suggest that *Salmonella* can have prolonged persistence on shared spaces. However, persistence depends on surface material and precautionary measures should be adopted to reduce exposure.

{{< pagebreak >}}

# Introduction (Why *Salmonella*?)

*Salmonella enterica* infections are a significant public health threat, responsible for over 93 million annual cases of human illness worldwide. In the United States alone, 1.35 million cases of salmonellosis and 420 deaths are reported annually. Most cases of human salmonellosis are caused by food-borne *Salmonella* strains associated with produce or undercooked meat. However, a rising subset of human infections are often associated with unidentified environmental exposures or contact with animals.

The American white ibis (Eudocimus albus), forms large nesting colonies in natural wetlands. However, due to habitat destruction in Everglades National Park, this gregarious member of the Pelecaniformes, has begun to urbanize parks with abundant anthropgenic food and water resources throughout South Florida. Infection with *Salmonella* spp. in American white ibis is well documented. It has been found that ibis can harbor diverse *Salmonella* strains and shed it at a higher prevalence than their natural counterparts. Studies have genetically matched the *Salmonella* shed by ibis to human salmonellosis cases in South Florida during the same time period ibis were sampled.

Similarly, in 2021, following an epidemiological investigation of 29 human cases of *Salmonella enterica* serovar Typhimurium (*S.* Typhimurium), the Centers for Disease Control and Prevention (CDC) documented a relationship between human cases and a concurrent avian salmonellosis outbreak. Of the 29 cases, 14 individuals were hospitalized and the outbreak was likely due to contact with bird feeders, sick or dead wild birds, or pets that had contact with wild birds.

Persistence of *Salmonella* can be affected by ambient temperature, presence/absence of a biofilm, and environmental nutrient conditions. Under ideal conditions (e.g. optimal temperature 35-43°C, pH 7-7.5, etc.), *Salmonella* is known to persist in the environment for extended periods of time. While *Salmonella* persistence has been examined on soil and poultry litter, there is a lack of understanding of the persistence of *Salmonella* on various surfaces including wood, plastic, and surfaces lined with antimicrobial coatings.

## Description of data and data source

The data selected for my MADA project is persistence and prevalence data from two trial studies. The first trial was conducted in South Florida where picnic tables, under the same pavilion and exposed to conditions known to affect *Salmonella*, were selected, swabbed daily for a week and shipped back to the University of Georgia for *Salmonella* isolation. The second trial was conducted in Athens, Georgia, at Whitehall Forest in a modified shade house. Bird feeders of different materials (e.g., cedar, plastic, plastic coated, and wood coated) were hung in triplicate and seeded with an isolate of *S.* Typhimurium isolated from a songbird.

The data from the first trial in South Florida was collected by an undergraduate student of Dr. Sonia Hernandez and provided to me by Dr. Hernandez. All data from the Feeder Trial was collected by me, Kimberly Perez under the guidance of Dr. Sonia Hernandez.

The Picnic Table Trial data is composed of several worksheets embedded in one excel file. After viewing the file (and running it below), there seems to be 192 observations. I am not sure how the Picnic Table Trial was collected but hope to visit with Dr. Hernandez regarding data collection methods soon. Feeder Trial data was collected and noted upon confirmation of *Salmonella* from processed samples. There are 144 observations for the Feeder Trial which spanned nearly 2 months. For both trials, we were attempting to measure *Salmonella* persistence. The Picnic Table Trial collected data in an attempt to measure prevalence as well. (Please Note: Aforementioned data was imported under "Data import and cleaning" section.)

## Questions/Hypotheses to be addressed

Given the uncertainty of *Salmonella's* persistence on various surface materials, from my data I would like to answer:
- What is the prevalence of *Salmonella* over time? 
- How long does *Salmonella* persist on different surface materials (e.g., wood, plastic, and those with an antimicrobial coating) exposed to various climatic factors (e.g., rain, sun, changing temperatures)

From these questions, I hope to better inform the public and public health practitioners on safer feeder handling methods and precautions to take when frequenting public locations where humans and wildlife (e.g., ibis) interact.

To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above (here `dataanalysis_template_references.bib`) and have the right bibtex key. Then you can include like this:

Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a]

{{< pagebreak >}}

# Methods
Coming soon...
_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._

## Data aquisition
_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._

The data from the first trial in South Florida was collected by an undergraduate student of Dr. Sonia Hernandez and provided to me by Dr. Hernandez. All data from the Feeder Trial was collected by me, Kimberly Perez under the guidance of Dr. Sonia Hernandez.

# Data import and cleaning

## Picnic Table Data 
```{r}
library(readxl)
PTD<- read_excel("C:/GitHub/MADA/KimberlyPerez-MADA-project/data/raw_data/Picnic_Table_Data.xlsx")
View(PTD)
summary(PTD)
```
## Feeder Data
```{r}
library(readcsv)
FD<- read.csv("C:/GitHub/MADA/KimberlyPerez-MADA-project/data/raw_data/Feeder Data.csv")
View(FD)
summary(FD)
```

_Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along._

## Statistical analysis
_Explain anything related to your statistical analyses._

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

_Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project._

@tbl-summarytable shows a summary of the data.

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/summarytable.rds")
knitr::kable(resulttable)
```

## Basic statistical analysis

While I am unsure of proper analysis, I may start by calculating probabilities. It may also be helpful to run a regression or linear model. Other than this, I am unsure of how to start the analysis process for the Picnic Table data.

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
knitr::include_graphics(here("results","height_weight_stratified.png"))
```

## Full analysis

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","resulttable2.rds"))
knitr::kable(resulttable2)
```

{{< pagebreak >}}

# Discussion

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.

{{< pagebreak >}}

# References

### 2023-02-09 products/manuscript/Manuscript.qmd
Updated with here() command

@@
 library(readxl)
-PTD<- read_excel("C:/GitHub/MADA/KimberlyPerez-MADA-project/data/raw_data/Picnic_Table_Data.xlsx")
+PTD<- read_excel(here("data", "raw_data", "Picnic_Table_Data.xlsx"))
 View(PTD)
@@
 library(readcsv)
-FD<- read.csv("C:/GitHub/MADA/KimberlyPerez-MADA-project/data/raw_data/Feeder Data.csv")
+FD<- read.csv(here("data", "raw_data", "Feeder Data.csv"))
 View(FD)

### 2023-02-10 products/manuscript/Manuscript.qmd
Run and Render
#Note to self: Be careful with package names and functions

@@
 ```{r}
-library(readcsv)
-FD<- read.csv(here("data", "raw_data", "Feeder Data.csv"))
+library(readr)
+FD<- read_csv(here("data", "raw_data", "Feeder Data.csv"))
 View(FD)

### 2023-03-10 code/analysis_code/exploratory_analysis.qmd
Project Part 2

---
title: "Combined Data Cleaning and Exploratory Analysis Script for Salmonella Environmental Persistence Informs Management Relevant to Avian and Public Health"
author: "Kimberly M. Perez"
date: "10/28/2021"
output: html_document
---
# **Data Processing Methodology**
Before attention can be focused on the data analysis portion, removal of unnecessary variables and visualizing data will provide a clearer picture of the data. This section will focus on data cleaning and wrangling. First, both datasets must be loaded. Prior to loading any data into RStudio, ensure that a new Quarto or RMarkdown File is established and linked to the preferred working directory. Basic packages need to be loaded to read in the data. Such packages to download include read_csv(), read_excel(), and here(), each of which can be loaded by utilizing the library() function (e.g., library(here)). Next, the datasets from both the Ibis and Feeder trials are loaded into RStudio utilizing the appropriate function (e.g., read_csv() for .csv files and read_excel() for .xlsx files). After data is loaded into R, it is best to get acquainted with the data by utilizing basic functions such as glimpse(). Glimpse provides me with insight into my data such as column names, number of rows and how each row is classified (e.g., character or numeric). After viewing the data I can work on cleaning the data by removing unnecessary variables. I would like to make the data a little more uniform and easier to work with. With that, I will change "Yes" to 1, and *NAs* to 0. I believe this will help me further in the analysis process. Given the variablility in the column names, I will then clean the names up to make them easier to include in my code as I move forward. With all of that, it is finally time to visualize the data by making some tables and/or graphs.

Here, some more packages were loaded that I forsaw using in this section. Before visualization began, the glimpse() function was utilized to view the recently cleaned data. Given the uncertainty of *Salmonella's* persistence on various surface materials and the information these trials hoped to provide, the summary() function was utilized to provide a quick glimpse of basic result summaries for positive *Salmonella* samples and if the pile present was the original pile or not. Next, I continued to explore these two variables of interest by plotting each using a stacked bar graph via ggplot().

I followed the same workflow for the Feeder Data, however, this dataset required less wrangling and did not provide as many exploratory options as the main focus was on feeder type, date, and the presence/absence of *Salmonella* for each day the feeders were sampled.

## **Data import and cleaning**

## **Initial Package Loading**
```{r}
# load a few R packages
library(here)
library(knitr)
library(dplyr)
library(readxl)
```

## **Picnic Table Data Loading**
```{r}
PTD<- read_excel(here("data", "raw_data", "Picnic_Table_Data.xlsx"))
View(PTD)
summary(PTD)
```
## **Picnic Table Data Wrangling**
```{r}
glimpse(PTD)

PTD1<- PTD %>% select(-c(Day, `Shippng Date`, Notes, `STAB TUBE NUMBER`, `ORIGIN OF FECES`, ...13)) #Removing columns

PTD1$`Original PP still intact?`<- ifelse(PTD1$`Original PP still intact?`=="Yes", 1, 0)

PTD2<- PTD1 %>% 
    mutate(`Original PP still intact?` = ifelse(is.na(`Original PP still intact?`), 0, `Original PP still intact?`)) #NAs to 0 for column `Original PP still intact?`

PTD3<- PTD2 %>% 
    mutate(`Pos orNeg`= ifelse(is.na(`Pos orNeg`), 0, `Pos orNeg`)) #NAs to 0 for column `Pos orNeg`

glimpse(PTD3) #Current variable names

PTD_clean<- PTD3 %>% #Renaming column names
  rename(
    ID= `Sample ID`,
    Sample_Type= `Sample Type (Pooled or Persistence)`,
    Salmonella_Positive= `Pos orNeg`,
    Original_Persistence_Pile= `Original PP still intact?`)

glimpse(PTD_clean)
```

## **Picnic Table Data Exploration and Visualization**
```{r}
#After cleaning I have decided I need to download a few more packages for the data visualization

library(gtsummary) #Tables 
library(plotly) #I may try for an interactive plot, this is a good resource to use
library(ggplot2) #For basic visualization

PTDsumt1<- summary(PTD_clean$Salmonella_Positive) #Quick glimpse of samples, non-specific (e.g., persistence/pool) based on Salmonella positivity
PTDsumt1

PTDsumt2<- summary(PTD_clean$Original_Persistence_Pile) #Quick glimpse of samples, non-specific (e.g., persistence/pool) based on Persistence Piles (e.g., OG or not)

PTDH<- ggplot(PTD_clean, 
              aes(fill=Sample_Type, y= Salmonella_Positive, x= Date)) + 
  geom_bar (position="stack", stat="identity") +
  geom_vline(xintercept= as.numeric (PTD_clean$Date[c(95)]), linetype= "dotted", color="red", lwd=1)

PTDH #Basic Plot: Positive Salmonella Samples by Sample Type via stacked plot, red line denotes decrease, will tap Dr. Hernandez for more information on this decline (e.g., weather, shipping, etc.)

PTD_Quad<- ggplot(PTD_clean, 
              aes(fill=Quadrant, y= Salmonella_Positive, x= Date)) + 
  geom_bar (position="stack", stat="identity") +
  geom_vline(xintercept= as.numeric (PTD_clean$Date[c(95)]), linetype= "dotted", color="green", lwd=1)

PTD_Quad #Basic Plot: Positive Salmonella Samples (pooled v. persistence) by Quadrant via stacked plot green line denotes decrease, will tap Dr. Hernandez for more information on this decline (e.g., weather, shipping, etc.)  
```
## **Saving the Figures and Tables from Picnic Table Data**

```{r}
#Here is where I will save my final tables and figures
```
We will continue loading, cleaning, and exploring the data we will use for the final project

## **Feeder Data Loading**
```{r}
library(readr)
FD<- read_csv(
  here("data", "raw_data", "Feeder Data.csv"))

View(FD)

summary(FD)

str(FD)

glimpse(FD)
```

## **Feeder Data Wrangling**
```{r}
FD1<- FD %>% select(-c(`Absence (0)/Presence (1)`, `...5`)) #Removing columns

glimpse(FD1) #Quick glimpse at the data, I can determine if I should wrangle the data a bit more

FD_clean<- FD1 #Data looked clean so I will rename it to use with my data exploration and visualization process below 
```

**Picnic Table Data Exploration and Visualization**
```{r}
summary(FD_clean$Absence_0_Presence_1) # Utilizing summary() for a quick look at result summaries- not much info here

FD_plot<- ggplot(FD_clean, #Given my limited dataset I will focus my visualization on feeder type and presence/absence of Salmonella 
              aes(fill=Feeder_Type, y=Absence_0_Presence_1, x= Collection_Date)) + 
  geom_bar (position="stack", stat="identity")

FD_plot  
```

## **Saving the Figures and Tables from Feeder Data**

```{r}
#Here is where I will save my final tables and figures to my results folder/subfolders 
```

### 2023-03-10 products/manuscript/Manuscript.qmd
Project Part 2

@@
 ---
-title: "Example Manuscript Template for a Data Analysis Project"
+title: "Salmonella Environmental Persistence Informs Management Relevant to Avian and Public Health"
 subtitle: ""
@@
 
-The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. 
-You don't need to have exactly these sections, but the content covering those sections should be addressed.
+# **Summary/Abstract**
 
-This uses MS Word as output format.
-[See here](https://quarto.org/docs/output-formats/ms-word.html) for more information.
-You can switch to other formats, like html or pdf.
-See [the Quarto documentation](https://quarto.org/) for other formats.
-
-```{r, echo=FALSE, message=FALSE}
-# load a few R packages
-library(here)
-library(knitr)
-```
-
-# Summary/Abstract
 Globally, *Salmonella* is a significant public health threat.
@@
 
-{{< pagebreak >}}
-
-# Introduction (Why *Salmonella*?)
+# **Introduction (Why *Salmonella*?)**
 
@@
 
-## Description of data and data source
+## **Description of data and data source**
 
@@
 
-## Questions/Hypotheses to be addressed
+## **Questions/Hypotheses to be addressed**
 
-Given the uncertainty of *Salmonella's* persistence on various surface materials, from my data I would like to answer:
-- What is the prevalence of *Salmonella* over time? 
+Given the uncertainty of *Salmonella's* persistence on various surface materials, from my data I would like to answer: - What is the prevalence of *Salmonella* over time?
 - How long does *Salmonella* persist on different surface materials (e.g., wood, plastic, and those with an antimicrobial coating) exposed to various climatic factors (e.g., rain, sun, changing temperatures)
@@
 
-To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above (here `dataanalysis_template_references.bib`) and have the right bibtex key.
-Then you can include like this:
+# **Methods**
 
-Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a]
+## **Ibis Picnic Table Trial Experimental Design and Sample Collection Methodology**
 
-{{< pagebreak >}}
+The ibis picnic table trial was conducted at Dreher Park, a public park in Palm Beach County frequented by white ibis who are known to carry *Salmonella*.
+Because humans and ibis both utilize the wooden picnic tables at the park, three tables under a single pavilion with similar environmental conditions (e.g., exposure to sunlight, rainfall, etc.) were selected.
+All tables were divided into four quadrants (e.g., tabletop plus bench) and ibis were encouraged to feed on the picnic tables until they deficated.
+Pictures were taken of every quadrant daily to facilitate the counting of new fecal piles.
+If feces were not deposited in a quadrant, fresh feces from the concrete pad was collected and placed onto the quadrant.
+To determine persistence, one large fecal pile from each quadrant was selected as the "persistence pile" and monitored and sampled daily for 8 days.
+To determine table prevalence, a pooled sample was taken daily from every quadrant by swabbing the surface picnic table surface with a sampling sponge, avoiding the designated "persistence pile".
+To maintain cleanliness for other park patrons, new piles were not smeared with the sampling sponge.
+Samples were placed into 20 mL of dulcitol selenite and maintained at room temperature prior to being shipped to the University of Georgia (UGA), every 2-4 days, for processing.
 
-# Methods
-Coming soon...
-_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._
+## **Feeder Trial Experimental Design and Sample Collection Methodology**
 
-## Data aquisition
-_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._
+Our Feeder trial aimed to assess the persistence of *Salmonella* on four types of feeders (cedar, plastic, cedar + antimicrobial coating, and plastic + antimicrobial coating) in Georgia.
+18 g of fresh chicken feces was collected, homogenized, 1 g was aliquoted into individual 15 mL test tubes, then sterilized through autoclaving.
+The individual alioquots were then inoculated with 1.5x10\^7 mL of a *Salmonella* Typhimurium isolate responsible for a previous avian salmonellosis outbreak.
+Samples were then transported to a modified shade house at Whitehall Forest where four feeder types, represented in triplicate and randomized, were seeded with 1 g of inoculated feces and 0.5 mL of sterile physiologic saline.
+Feeders were then swabbed daily for one week, every other day for one week, and weekly for an additional week.
+Swabs were embedded into 10 mL of Dulcitol Selenite broth and transported back to UGA for processing.
 
+## **Sample Processing Methodology**
+After collection, 100 uL of the sample vortexed in dulcitol selentite was inoculated into 10 mL of Rappaport Vassiliadis (RV) broth.
+The broth was incubated for 24 hours at 37c.
+The RV broth was then streaked onto XLT-4 agar plates and incubated at 37c for 24 hours, then an additional 24 hours on the lab bench at room temperature.
+If colonies morphometrically consistent with *Salmonella* were present, one colony per plate was chosen and a stab in Luria Broth (LB) agar was created.
+To confirm the presence of *Salmonella*, patch plates were made on ChromAgar using the stabbed colonies.
+If the patches turned magenta, that indicated positive *Salmonella* samples.
+A blue color indicated something other than *Salmonella.*
+
+## **Data acquisition**
 The data from the first trial in South Florida was collected by an undergraduate student of Dr.
@@
 
-# Data import and cleaning
+## **Data Processing Methodology**
+Before attention can be focused on the data analysis portion, removal of unnecessary variables and visualizing data will provide a clearer picture of the data.
+This section will focus on data cleaning and wrangling.
+First, both datasets must be loaded.
+Prior to loading any data into RStudio, ensure that a new Quarto or RMarkdown File is established and linked to the preferred working directory.
+Basic packages need to be loaded to read in the data.
+Such packages to download include read_csv(), read_excel(), and here(), each of which can be loaded by utilizing the library() function (e.g., library(here)).
+Next, the datasets from both the Ibis and Feeder trials are loaded into RStudio utilizing the appropriate function (e.g., read_csv() for .csv files and read_excel() for .xlsx files).
+After data is loaded into R, it is best to get acquainted with the data by utilizing basic functions such as glimpse().
+Glimpse provides me with insight into my data such as column names, number of rows and how each row is classified (e.g., character or numeric).
+After viewing the data I can work on cleaning the data by removing unnecessary variables.
+I would like to make the data a little more uniform and easier to work with.
+With that, I will change "Yes" to 1, and *NAs* to 0.
+I believe this will help me further in the analysis process.
+Given the variablility in the column names, I will then clean the names up to make them easier to include in my code as I move forward.
+With all of that, it is finally time to visualize the data by making some tables and/or graphs.
 
-## Picnic Table Data 
+Here, some more packages were loaded that I forsaw using in this section.
+Before visualization began, the glimpse() function was utilized to view the recently cleaned data.
+Given the uncertainty of *Salmonella's* persistence on various surface materials and the information these trials hoped to provide, the summary() function was utilized to provide a quick glimpse of basic result summaries for positive *Salmonella* samples and if the pile present was the original pile or not.
+Next, I continued to explore these two variables of interest by plotting each using a stacked bar graph via ggplot().
+
+I followed the same workflow for the Feeder Data, however, this dataset required less wrangling and did not provide as many exploratory options as the main focus was on feeder type, date, and the presence/absence of *Salmonella* for each day the feeders were sampled.
+
+# **Data import and cleaning**
+
+## **Initial Package Loading**
 ```{r}
+# load a few R packages
+library(here)
+library(knitr)
+library(dplyr)
 library(readxl)
+```
+
+## **Picnic Table Data Loading**
+```{r}
 PTD<- read_excel(here("data", "raw_data", "Picnic_Table_Data.xlsx"))
@@
 ```
-## Feeder Data
+## **Picnic Table Data Wrangling**
+```{r}
+glimpse(PTD)
+
+PTD1<- PTD %>% select(-c(Day, `Shippng Date`, Notes, `STAB TUBE NUMBER`, `ORIGIN OF FECES`, ...13)) #Removing columns
+
+PTD1$`Original PP still intact?`<- ifelse(PTD1$`Original PP still intact?`=="Yes", 1, 0)
+
+PTD2<- PTD1 %>% 
+    mutate(`Original PP still intact?` = ifelse(is.na(`Original PP still intact?`), 0, `Original PP still intact?`)) #NAs to 0 for column `Original PP still intact?`
+
+PTD3<- PTD2 %>% 
+    mutate(`Pos orNeg`= ifelse(is.na(`Pos orNeg`), 0, `Pos orNeg`)) #NAs to 0 for column `Pos orNeg`
+
+glimpse(PTD3) #Current variable names
+
+PTD_clean<- PTD3 %>% #Renaming column names
+  rename(
+    ID= `Sample ID`,
+    Sample_Type= `Sample Type (Pooled or Persistence)`,
+    Salmonella_Positive= `Pos orNeg`,
+    Original_Persistence_Pile= `Original PP still intact?`)
+
+glimpse(PTD_clean)
+```
+## **Picnic Table Data Exploration and Visualization**
+```{r}
+library(gtsummary)
+library(plotly)
+library(ggplot2)
+
+summary(PTD_clean$Salmonella_Positive) #Quick glimpse of samples, non-specific (e.g., persistence/pool) based on Salmonella positivity
+
+summary(PTD_clean$Original_Persistence_Pile) #Quick glimpse of samples, non-specific (e.g., persistence/pool) based on Persistence Piles (e.g., OG or not)
+
+PTDH<- ggplot(PTD_clean, 
+              aes(fill=Sample_Type, y= Salmonella_Positive, x= Date)) + 
+  geom_bar (position="stack", stat="identity") +
+  geom_vline(xintercept= as.numeric (PTD_clean$Date[c(95)]), linetype= "dotted", color="red", lwd=1)
+
+PTDH #Basic Plot: Positive Salmonella Samples by Sample Type via stacked plot, red line denotes decrease, will tap Dr. Hernandez for more information on this decline (e.g., weather, shipping, etc.)
+
+PTD_Quad<- ggplot(PTD_clean, 
+              aes(fill=Quadrant, y= Salmonella_Positive, x= Date)) + 
+  geom_bar (position="stack", stat="identity") +
+  geom_vline(xintercept= as.numeric (PTD_clean$Date[c(95)]), linetype= "dotted", color="green", lwd=1)
+
+PTD_Quad #Basic Plot: Positive Salmonella Samples (pooled v.
+persistence) by Quadrant via stacked plot green line denotes decrease, will tap Dr.
+Hernandez for more information on this decline (e.g., weather, shipping, etc.)
+```
+
+## **Feeder Data Loading**
 ```{r}
 library(readr)
-FD<- read_csv(here("data", "raw_data", "Feeder Data.csv"))
+FD<- read_csv(
+  here("data", "raw_data", "Feeder Data.csv"))
 View(FD)
+
 summary(FD)
+
+str(FD)
+
+glimpse(FD)
 ```
+## **Feeder Data Wrangling**
+```{r}
+FD1<- FD %>% select(-c(`Absence (0)/Presence (1)`, `...5`)) #Removing columns
 
-_Write code that reads in the file and cleans it so it's ready for analysis.
-Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts.
-If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g.
-as supplement in a paper).
-All materials, including files that contain code, should be commented well so everyone can follow along._
+glimpse(FD1) #Quick glimpse at the data, I can determine if I should wrangle the data a bit more
 
-## Statistical analysis
-_Explain anything related to your statistical analyses._
+FD_clean<- FD1 #Data looked clean so I will rename it to use with my data exploration and visualization process below 
+```
+**Picnic Table Data Exploration and Visualization**
+```{r}
+summary(FD_clean$Absence_0_Presence_1) # Utilizing summary() for a quick look at result summaries- not much info here
+
+FD_plot<- ggplot(FD_clean, #Given my limited dataset I will focus my visualization on feeder type and presence/absence of Salmonella 
+              aes(fill=Feeder_Type, y=Absence_0_Presence_1, x= Collection_Date)) + 
+  geom_bar (position="stack", stat="identity")
+
+FD_plot  
+```
+{{< pagebreak >}}
+
+# **Future Project PARTS for KP to populate**
+
+# **Statistical analysis**
+
+# **Results**
+
+## **Exploratory/Descriptive analysis**
+
+## **Basic statistical analysis**
+
+After the modules following Part 1, I am leaning towards running a GLM for both given the binary responses for Salmonella persistence (e.g., presence v.
+absence).
+I need to meet with Dr.
+Hernandez to clear up some questions I have on the picnic table data, specifically regarding the introduction of new feces.
+I am curious if I should include other data into the mix to make my analysis more robust.
+At the moment, the variables of interest for both are limited (e.g., two for the picnic table data, and one primary variable of interest for the feeder trial).
+
+## **Full analysis**
+
+# **Discussion**
+
+## **Summary and Interpretation**
+
+## **Strengths and Limitations**
+
+## **Conclusions**
 
@@
 
-# Results
-
-## Exploratory/Descriptive analysis
-
-_Use a combination of text/tables/figures to explore and describe your data.
-Show the most important descriptive results here.
-Additional ones should go in the supplement.
-Even more can be in the R and Quarto files that are part of your project._
-
-@tbl-summarytable shows a summary of the data.
-
-Note the loading of the data providing a **relative** path using the `../../` notation.
-(Two dots means a folder up).
-You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path.
-You can also use the `here` R package to create paths.
-See examples of that below.
-
-```{r}
-#| label: tbl-summarytable
-#| tbl-cap: "Data summary table."
-#| echo: FALSE
-resulttable=readRDS("../../results/summarytable.rds")
-knitr::kable(resulttable)
-```
-
-## Basic statistical analysis
-
-While I am unsure of proper analysis, I may start by calculating probabilities.
-It may also be helpful to run a regression or linear model.
-Other than this, I am unsure of how to start the analysis process for the Picnic Table data.
-
-_To get some further insight into your data, if reasonable you could compute simple statistics (e.g.
-simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable.
-Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._
-
-@fig-result shows a scatterplot figure produced by one of the R scripts.
-
-```{r}
-#| label: fig-result
-#| fig-cap: "Height and weight stratified by gender."
-#| echo: FALSE
-knitr::include_graphics(here("results","height_weight_stratified.png"))
-```
-
-## Full analysis
-
-_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc.
-This might again be code that is best placed in one or several separate R scripts that need to be well documented.
-You want the code to produce figures and data ready for display as tables, and save those.
-Then you load them here._
-
-Example @tbl-resulttable2 shows a summary of a linear model fit.
-
-```{r}
-#| label: tbl-resulttable2
-#| tbl-cap: "Linear model fit table."
-#| echo: FALSE
-resulttable2 = readRDS(here("results","resulttable2.rds"))
-knitr::kable(resulttable2)
-```
-
-{{< pagebreak >}}
-
-# Discussion
-
-## Summary and Interpretation
-_Summarize what you did, what you found and what it means._
-
-## Strengths and Limitations
-_Discuss what you perceive as strengths and limitations of your analysis._
-
-## Conclusions
-_What are the main take-home messages?_
-
-_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_
-
-This paper [@leek2015] discusses types of analyses.
-
-These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.
-
-Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header.
-Many more style files for almost any journal [are available](https://www.zotero.org/styles).
-You also specify the location of your bibtex reference file in the YAML.
-You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.
-
-{{< pagebreak >}}
-
-# References
+# **References**

### 2023-03-10 products/manuscript/supplement/Supplementary_Material.qmd
Project Part 2

---
title: "Supplement to Example Manuscript Template for a Data Analysis Project"

subtitle: ""
author: Andreas Handel
date: "`r Sys.Date()`"
format:
  pdf:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../dataanalysis_template_references.bib
csl: ../../vancouver.csl
---

# **Overview**
Overview of what I will include for readers needs to be included here.

# **Code and file information**

Here is where I will provide a step-by-step document so others will be able to follow along and easily reproduce my code. I will utilize some of the text I included in the main text/Manuscript.qmd and will expand upon it as I work through this project.

{{< pagebreak >}}

# **Additional Method Details**
{{< pagebreak >}}

# **Additional results**
{{< pagebreak >}}

# **Discussion**
{{< pagebreak >}}

# **References**

### 2023-03-10 code/processing_code/processingcode.R

@@ -1,151 +1,100 @@
 ###############################
-# processing script
+# Processing Picnic Table and Feeder Data Script
 #
-#this script loads the raw data, processes and cleans it 
-#and saves it as Rds file in the processed_data folder
+# The script I use below will load the raw data, process and clean it for both 
+# The picnic table data and the feeder data will also be saved as a RDS file in the processed_data folder
+# I will also include full text and code in my manuscript in the meantime. 
 #
-# Note the ## ---- name ---- notation
-# This is done so one can pull in the chunks of code into the Quarto document
-# see here: https://bookdown.org/yihui/rmarkdown-cookbook/read-chunk.html
+# KP USE AS REFERENCE AS YOU ARE COMPLETING PROJECT https://bookdown.org/yihui/rmarkdown-cookbook/read-chunk.html
 
 
-## ---- packages --------
-#load needed packages. make sure they are installed.
-library(readxl) #for loading Excel files
-library(dplyr) #for data processing/cleaning
-library(tidyr) #for data processing/cleaning
-library(skimr) #for nice visualization of data 
-library(here) #to set paths
+## ---- Packages --------
+# Let's load a few R packages for processing
+library(knitr)
+library(dplyr) #Loading in for data cleaning
+library(readxl) #This will load in excel files (picnic table )
+library(here) #Very important as this will circumvent any absolute path issues, makes more reproducible, easily set paths
 
-## ---- loaddata --------
-#path to data
-#note the use of the here() package and not absolute paths
-data_location <- here::here("data","raw_data","exampledata.xlsx")
+## ---- Load Picnic Table Data --------
+PTD<- read_excel(here("data", "raw_data", "Picnic_Table_Data.xlsx"))
 
-#load data. 
-#note that for functions that come from specific packages (instead of base R)
-# I often specify both package and function like so
-#package::function() that's not required one could just call the function
-#specifying the package makes it clearer where the function "lives",
-#but it adds typing. You can do it either way.
-rawdata <- readxl::read_excel(data_location)
+## ---- ExploreData --------
+# Let's take a look at the data we just loaded in
+View(PTD)
 
-# We can look in the Codebook (second tab) for a variable explanation
-codebook <- readxl::read_excel(data_location, sheet ="Codebook")
-print(codebook)
+summary(PTD)
 
+glimpse(PTD)
 
[…]

### 2023-03-18 code/analysis_code/exploratory_analysis.qmd
Push Project Part 3

@@
 PTDsumt2<- summary(PTD_clean$Original_Persistence_Pile) #Quick glimpse of samples, non-specific (e.g., persistence/pool) based on Persistence Piles (e.g., OG or not)
+PTDsumt2
 
@@
 PTDH #Basic Plot: Positive Salmonella Samples by Sample Type via stacked plot, red line denotes decrease, will tap Dr. Hernandez for more information on this decline (e.g., weather, shipping, etc.) 
+ggsave(PTDH, file="Picnic_Table_PosSamp.png")
 
@@
 Hernandez for more information on this decline (e.g., weather, shipping, etc.)
+ggsave(PTD_Quad, file="Picnic_Table_Quad.png")
+
 ```
-## **Saving the Figures and Tables from Picnic Table Data**
-
-```{r}
-#Here is where I will save my final tables and figures
-```
-We will continue loading, cleaning, and exploring the data we will use for the final project
-
 ## **Feeder Data Loading**
@@
 FD_plot  
+ggsave(FD_plot, file="Feeder_PA.png")
 ```
-
-## **Saving the Figures and Tables from Feeder Data**
-
-```{r}
-#Here is where I will save my final tables and figures to my results folder/subfolders 
-```

### 2023-03-18 products/manuscript/Manuscript.qmd
Manuscript Update (Attempt 2)
Uploaded manuscript not updated one!
Push Project Part 3

@@
 ## **Sample Processing Methodology**
+
 After collection, 100 uL of the sample vortexed in dulcitol selentite was inoculated into 10 mL of Rappaport Vassiliadis (RV) broth.
@@
 ## **Data acquisition**
+
 The data from the first trial in South Florida was collected by an undergraduate student of Dr.
@@
 ## **Data Processing Methodology**
+
 Before attention can be focused on the data analysis portion, removal of unnecessary variables and visualizing data will provide a clearer picture of the data.
@@
 
-# **Data import and cleaning**
+**Picnic Table Data Exploration and Visualization**
 
-## **Initial Package Loading**
-```{r}
-# load a few R packages
-library(here)
-library(knitr)
-library(dplyr)
-library(readxl)
-```
+![Positive Salmonella Samples by Type and Date](images/Picnic_Table_PosSamp-04.png) Positive *Salmonella* Samples by Sample Type and Date
 
-## **Picnic Table Data Loading**
-```{r}
-PTD<- read_excel(here("data", "raw_data", "Picnic_Table_Data.xlsx"))
-View(PTD)
-summary(PTD)
-```
-## **Picnic Table Data Wrangling**
-```{r}
-glimpse(PTD)
+![Positive *Salmonella* Samples by Quadrant and Date](images/Picnic_Table_Quad-01.png)
 
-PTD1<- PTD %>% select(-c(Day, `Shippng Date`, Notes, `STAB TUBE NUMBER`, `ORIGIN OF FECES`, ...13)) #Removing columns
+**Bird Feeder Data Exploration and Visualization**
 
-PTD1$`Original PP still intact?`<- ifelse(PTD1$`Original PP still intact?`=="Yes", 1, 0)
+![Positive *Salmonella* Samples by feeder type and date.](images/Picnic_Table_PosSamp.png)
 
-PTD2<- PTD1 %>% 
-    mutate(`Original PP still intact?` = ifelse(is.na(`Original PP still intact?`), 0, `Original PP still intact?`)) #NAs to 0 for column `Original PP still intact?`
-
-PTD3<- PTD2 %>% 
-    mutate(`Pos orNeg`= ifelse(is.na(`Pos orNeg`), 0, `Pos orNeg`)) #NAs to 0 for column `Pos orNeg`
-
-glimpse(PTD3) #Current variable names
-
-PTD_clean<- PTD3 %>% #Renaming column names
-  rename(
-    ID= `Sample ID`,
-    Sample_Type= `Sample Type (Pooled or Persistence)`,
-    Salmonella_Positive= `Pos orNeg`,
-    Original_Persistence_Pile= `Original PP still intact?`)
-
-glimpse(PTD_clean)
-```
-## **Picnic Table Data Exploration and Visualization**
-```{r}
-library(gtsummary)
-library(plotly)
-library(ggplot2)
-
-summary(PTD_clean$Salmonella_Positive) #Quick glimpse of samples, non-specific (e.g., persistence/pool) based on Salmonella positivity
-
-summary(PTD_clean$Original_Persistence_Pile) #Quick glimpse of samples, non-specific (e.g., persistence/pool) based on Persistence Piles (e.g., OG or not)
-
-PTDH<- ggplot(PTD_clean, 
-              aes(fill=Sample_Type, y= Salmonella_Positive, x= Date)) + 
-  geom_bar (position="stack", stat="identity") +
-  geom_vline(xintercept= as.numeric (PTD_clean$Date[c(95)]), linetype= "dotted", color="red", lwd=1)
-
-PTDH #Basic Plot: Positive Salmonella Samples by Sample Type via stacked plot, red line denotes decrease, will tap Dr. Hernandez for more information on this decline (e.g., weather, shipping, etc.)
-
-PTD_Quad<- ggplot(PTD_clean, 
-              aes(fill=Quadrant, y= Salmonella_Positive, x= Date)) + 
-  geom_bar (position="stack", stat="identity") +
-  geom_vline(xintercept= as.numeric (PTD_clean$Date[c(95)]), linetype= "dotted", color="green", lwd=1)
-
-PTD_Quad #Basic Plot: Positive Salmonella Samples (pooled v.
-persistence) by Quadrant via stacked plot green line denotes decrease, will tap Dr.
-Hernandez for more information on this decline (e.g., weather, shipping, etc.)
-```
-
-## **Feeder Data Loading**
-```{r}
-library(readr)
-FD<- read_csv(
-  here("data", "raw_data", "Feeder Data.csv"))
-View(FD)
-
-summary(FD)
-
-str(FD)
-
-glimpse(FD)
-```
-## **Feeder Data Wrangling**
-```{r}
-FD1<- FD %>% select(-c(`Absence (0)/Presence (1)`, `...5`)) #Removing columns
-
-glimpse(FD1) #Quick glimpse at the data, I can determine if I should wrangle the data a bit more
-
-FD_clean<- FD1 #Data looked clean so I will rename it to use with my data exploration and visualization process below 
-```
-**Picnic Table Data Exploration and Visualization**
-```{r}
-summary(FD_clean$Absence_0_Presence_1) # Utilizing summary() for a quick look at result summaries- not much info here
-
-FD_plot<- ggplot(FD_clean, #Given my limited dataset I will focus my visualization on feeder type and presence/absence of Salmonella 
-              aes(fill=Feeder_Type, y=Absence_0_Presence_1, x= Collection_Date)) + 
-  geom_bar (position="stack", stat="identity")
-
-FD_plot  
-```
 {{< pagebreak >}}
@@
 
+## **Ibis Trial**
+
+Of 96 deposited feces, we obtained 27% positive.
+*Salmonella* persisted for an average of 3 days, and maximum of 4 days on
+wooden picnic tables.
+
+## *Feeder Trial*
+We confirmed successful inoculation of all
+feeders with *Salmonella* (presence on day post inoculation \[DPI\]
+0). Further, results indicate that *Salmonella* may persist on
+some feeders up to DPI 5 and current analysis indicates that *Salmonella* persisted
+on both coated and uncoated plastic feeders at a higher rate than wood and wood
+coated. Figure 2 depicts the percent of samples positive based on feeder type
+throughout the study period. While plastic and plastic coated feeders had an
+overall higher percent positivity, a spike in percent positivity in wood feeders occurred on DPI 5 after a rain event. Further analyses should be conducted to determine whether environmental
+conditions (e.g., rain event) had a significant influence on persistence.
+
+Our samples persisted for an average of 2 days (feeders), 3 days (picnic tables) with 4 (picnic tables)
+and 5 (feeders) being the maximum number of days persisted. The varied
+persistence during both trials could be as a result of a lack of moisture, the
+fluctuation in ambient temperature during sampling, and water activity. Our studies
+suggest that *Salmonella* can have prolonged persistence on shared spaces,
+persistence depends on surface material, and precautionary measures (i.e.,
+increased hygiene) should be adopted to reduce human exposure.
+
 ## **Exploratory/Descriptive analysis**
@@
 
-After the modules following Part 1, I am leaning towards running a GLM for both given the binary responses for Salmonella persistence (e.g., presence v.
+After the modules following Part 1, I am leaning towards running a GLM for both given the binary responses for *Salmonella* persistence (e.g., presence v.
 absence).

### 2023-03-18 code/analysis_code/statistical_analysis.R

@@ -1,21 +1,27 @@
 ###############################
-# analysis script
+# Analysis script
 #
 #this script loads the processed, cleaned data, does a simple analysis
 #and saves the results to the results folder
 
-#load needed packages. make sure they are installed.
-library(ggplot2) #for plotting
+#Load needed packages
+library(tidyverse)
+ibrary(ggplot2) #for plotting
 library(broom) #for cleaning up output from lm()
 library(here) #for data loading/saving
+library(tidymodels)
 
-#path to data
+#Path to Feeder Data
 #note the use of the here() package and not absolute paths
-data_location <- here::here("data","processed_data","processeddata.rds")
+data_location_FD <- here::here("data","processed_data","FDprocessed.RDS")
 
-#load data. 
-mydata <- readRDS(data_location)
+#Lets do the same for the Picnic Table Data 
+data_location_PT <- here::here("data","processed_data","PTprocessed.RDS")
 
+#Let's load the data we saved above  
+FD_Data <-readRDS(data_location_FD)
+
+PT_Data <-readRDS(data_location_PT)
 
 ######################################
 #Data fitting/statistical analysis
@@ -23,34 +29,57 @@ mydata <- readRDS(data_location)
 
 ############################
 #### First model fit
-# fit linear model using height as outcome, weight as predictor
+# fit linear model using Salmonella_Pos as outcome, Table as predictor
+lmfit3.0 <- lm(Salmonella_Positive~Table, PT_Data) 
+lmfit3.0
+plot(lmfit3.0)
 
-lmfit1 <- lm(Height ~ Weight, mydata)  
+# Here we place results from the above fit into a data frame via tidy function
+lmtable1 <- broom::tidy(lmfit3.0)
 
-# place results from fit into a data frame with the tidy function
-lmtable1 <- broom::tidy(lmfit1)
-
-#look at fit results
+#Let's check out the results
 print(lmtable1)
 
 # save fit results table  
-table_file1 = here("results", "resulttable1.rds")
-saveRDS(lmtable1, file = table_file1)
+Table1_PD= here("results", "PD_results_table.rds") #KP Double check path going to MADA Project not folder
+saveRDS(lmtable1, file = "Table1_PD")
 
 ############################
 #### Second model fit
-# fit linear model using height as outcome, weight and gender as predictor
 
-lmfit2 <- lm(Height ~ Weight + Gender, mydata)  
[…]

### 2023-03-18 code/processing_code/processingcode.R

@@ -1,7 +1,7 @@
 ###############################
 # Processing Picnic Table and Feeder Data Script
 #
-# The script I use below will load the raw data, process and clean it for both 
+# The script I use below will load the raw data, process and clean it for both. 
 # The picnic table data and the feeder data will also be saved as a RDS file in the processed_data folder
 # I will also include full text and code in my manuscript in the meantime. 
 #

### 2023-03-20 code/analysis_code/statistical_analysis.R
Continuing Project Work

@@ -6,10 +6,18 @@
 
 #Load needed packages
 library(tidyverse)
-ibrary(ggplot2) #for plotting
+library(ggplot2) #for plotting
 library(broom) #for cleaning up output from lm()
 library(here) #for data loading/saving
 library(tidymodels)
+library(tidyr)
+library(gtsummary)
+library(dplyr)
+library(vip)
+library(performance)
+library(jtools)
+library(broom)
+library(broom.mixed)
 
 #Path to Feeder Data
 #note the use of the here() package and not absolute paths
@@ -23,12 +31,76 @@ FD_Data <-readRDS(data_location_FD)
 
 PT_Data <-readRDS(data_location_PT)
 
+
+glimpse (FD_Data)
+
+glimpse(PT_Data)
 ######################################
 #Data fitting/statistical analysis
 ######################################
 
 ############################
-#### First model fit
+### Let's Begin with Some Linear Model Fitting; I will try this multiple ways, with several predictors!
+lmog<-linear_reg() %>% 
+  set_engine("lm")
+
+lmfit_PT<- lmog %>% #Training; outcome: Salmonella Pos and Main Predictor Table
+  fit(Salmonella_Positive~Table, PT_Data) 
+
+lmfit_PT
+
+glimpse(lmfit_PT)
+
+#Way 1
+glance(lmfit_PT)
+
+#Way 2
+tidy(lmfit_PT)
+
+TableTidy_PT= here("results", "PD_results_TableTidy.rds") #KP Double check path going to MADA Project not folder
+saveRDS(lmfit_PT, file = "TableTidy_PD")
+
+plot_summs(lmfit_PT)
+
+PlotSum_PD1= here("results", "PD_results_PlotSum.rds") #KP Double check path going to MADA Project not folder
+saveRDS(lmall_PT, file = "PlotSum1_PD")
+
+#Fitting other specific indicators
+lmog_1<-linear_reg() %>% 
+  set_engine("lm")
+
+lmfit_PT1<- lmog_1 %>% #Training; outcome: Salmonella Pos and Main Predictor Table
+  fit(Salmonella_Positive~Original_Persistence_Pile+Table, PT_Data) 
+
+lmfit_PT1
+
+glimpse(lmfit_PT1)
+
+#Way 1
+glance(lmfit_PT1)
+
+#Way 2
+tidy(lmfit_PT1)
+
+plot_summs(lmfit_PT1)
+
[…]

### 2023-04-08 1. Data Cleaning Script/DataCleaning.qmd
Project: Revamp! [IP]

---
editor: 
  markdown: 
    wrap: 72
---

# **Processing Picnic Table and Feeder Data Script**

The script I use below will load the raw data, process and clean it for
both.

The picnic table data and the feeder data will also be saved as a RDS
file in the processed_data folder.

I will also include full text and code in my manuscript in the meantime.

# **Loading Packages**
```{r, message=FALSE, warning=FALSE}
# Let's load a few R packages for processing
library(knitr)
library(dplyr) #Loading in for data cleaning
library(readxl) #This will load in excel files (picnic table )
library(here) #Very important as this will circumvent any absolute path issues, makes more reproducible, easily set paths
library(readr) #Easy to get data from csv files
```

# **1a. Loading: Picnic Table Data**
```{r}
PTD<- read_excel(here("0. Raw Data", "Picnic_Table_Data.xlsx"))
```

# **2a. Exploring: Picnic Table Data**
Let's take a look at the data we just loaded in.
```{r}
View(PTD)

summary(PTD)

glimpse(PTD)
```

# **3a. Cleaning: Picnic Table Data**

There is a lot to clean here. Overall, the excel seems a bit messy and
disorganized. From what I saw in step 2a, I want to:

-   Replace NAs

-   Keep the dates, type of sample (e.g., persistence/pooled), positive
    samples

-   Rename columns

-   Check how the cleaning process is going
```{r}
 PTD1<- PTD %>% 
  select(-c(Day, `Shippng Date`, Notes,`STAB TUBE NUMBER`, `ORIGIN OF FECES`, ...13)) #Removing columns
```

```{r}
PTD1$`Original PP still intact?`<- ifelse(PTD1$`Original PP still intact?`=="Yes", 1, 0)

PTD2<- PTD1 %>% mutate(`Original PP still intact?` =
    ifelse(is.na(`Original PP still intact?`), 0,
    `Original PP still intact?`)) #NAs to 0 for column`Original PP still intact?`

PTD3<- PTD2 %>% mutate(`Pos orNeg`= ifelse(is.na(`Pos orNeg`), 0,
`Pos orNeg`)) #NAs to 0 for column `Pos orNeg`

glimpse(PTD3) #Current variable names

PTD_clean<- PTD3 %>% #Renaming column names 
  rename( ID= `Sample ID`, Sample_Type= `Sample Type (Pooled or Persistence)`,
  Salmonella_Positive=`Pos orNeg`, Original_Persistence_Pile= `Original PP still intact?`)

glimpse(PTD_clean)
```

## **4a. Save the Data!**
The cleaned data looks much better! I would like to pick a location to save this portion as a file. In this step, I will save the file as an RDS to the **2. Cleaned Data** folder under the name "PTprocessed".
```{r}
save_data_location <- here::here("2. Clean Data","PTprocessed.rds")
saveRDS(PTD_clean, file = save_data_location)
```

# **Let's do it all again, this time with the Feeder Data**
# **1b. Loading: Feeder Data**
```{r}
FD<- read_csv(here("0. Raw Data", "Feeder Data.csv"))
```

# **2b. Exploring: Feeder Data**
Let's take a look at the data we just loaded into R!
```{r}
View(FD)

summary(FD)

str(FD)

glimpse(FD)
```

# **3b. Cleaning: Feeder Data**
Unlike the picnic table data, this is a fairly clean and short dataset. I think we can clean this with a few lines of code.
```{r}
FD1<- FD %>% select(-c(`Absence (0)/Presence (1)`, `...5`)) #Removing columns

glimpse(FD1) #Quick glimpse at the data, I can determine if I should wrangle the data a bit more

FD_clean<- FD1 #Data looked clean so I will rename it to use with my data exploration and visualization process below
```

## **4b. Save the Data!**
The cleaned data looks good. I would like to pick a location to save this portion as a file. I will save the file as an RDS to the **2. Cleaned Data** folder under the name "FDprocessed".
```{r}
save_data_location2<- here::here("2. Clean Data","FDprocessed.rds")
saveRDS(FD_clean, file = save_data_location2)
```

### 2023-04-08 3. Analysis Script/StatisticalAnalysis.qmd
Project: Revamp 2 [IP-SA .qmd REDO]

---
title: "Combined Data Cleaning and Statistical Analysis Script for Manuscript: Salmonella Environmental Persistence Informs Management Relevant to Avian and Public Health"
author: "Kimberly M. Perez"
date: "05/01/2023"
output: html_document
---

# **Analysis script**
This script loads the cleaned data, does a simple analysis, and saves the results to the "4. Results" folder

## **Loading Packages**
```{r}
#Load needed packages
library(tidyverse)
library(ggplot2) #for plotting
library(broom) #for cleaning up output from lm()
library(here) #for data loading/saving
library(tidymodels)
library(tidyr)
library(gtsummary)
library(dplyr)
library(vip)
library(performance)
library(jtools)
library(broom)
library(broom.mixed)
```

## **1. Loading: Picnic Table & Feeder Data**
```{r}
#Picnic Table Data
PT_Data<- readRDS(here::here("2. Clean Data","PTprocessed.rds"))

#Feeder Data 
FD_Data <-readRDS(here::here("2. Clean Data","FDprocessed.rds"))

glimpse(PT_Data)

glimpse (FD_Data)
```

# **Data Fitting/Statistical Analysis**
Let's Begin with some Linear Model Fitting; I will try this multiple ways, with several predictors.
```{r}
lmog<-linear_reg() %>% 
  set_engine("lm")

lmfit_PT<- lmog %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Salmonella_Positive~Table, PT_Data)

lmfit_PT

glimpse(lmfit_PT)

#Way 1
glance(lmfit_PT)

#Way 2
tidy(lmfit_PT)

TableTidy_PTD<- here::here("4. Results","PTD_TableTidy.png")

ggsave(lmfit_PT, file=TableTidy_PTD)

plot_summs(lmfit_PT)

PlotSum_PD1= here("results", "PD_results_PlotSum.rds") #KP Double check path going to MADA Project not folder
saveRDS(lmall_PT, file = "PlotSum1_PD")

#Fitting other specific indicators
lmog_1<-linear_reg() %>% 
  set_engine("lm")

lmfit_PT1<- lmog_1 %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Salmonella_Positive~Original_Persistence_Pile+Table, PT_Data)

lmfit_PT1

glimpse(lmfit_PT1)

#Way 1
glance(lmfit_PT1)

#Way 2
tidy(lmfit_PT1)

plot_summs(lmfit_PT1)

#Trying ALL predictors of Picnic Table dataset 
lmog_PT<- linear_reg() %>%
  set_engine("lm")

lmall_PT<- lmog_PT %>% #Training 
  fit(Salmonella_Positive~., data=PT_Data)

lmall_PT

tidy(lmall_PT)

plot_summs(lmall_PT, robust=TRUE) #KP NOTE: I need to clean this up so it is a bit more readable, must fullscreen it

PlotSum_PD1= here("results", "PD_results_PlotSum.rds") #KP Double check path going to MADA Project not folder
saveRDS(lmall_PT, file = "PlotSum1_PD")

#### Model fit another way
# fit linear model using Salmonella_Pos as outcome, Table as predictor
lmfit3.0 <- lm(Salmonella_Positive~Table, PT_Data) 
lmfit3.0
plot(lmfit3.0)

# Here we place results from the above fit into a data frame via tidy function
lmtable1 <- broom::tidy(lmfit3.0)

#Let's check out the results
print(lmtable1)

# save fit results table  
Table1_PD= here("results", "PD_results_table.rds") #KP Double check path going to MADA Project not folder
saveRDS(lmtable1, file = "Table1_PD")

#A different way Part Deux 
# I will do the same as above, however, I am now adding persistence pile
lmfit2 <- lm(Salmonella_Positive~Original_Persistence_Pile+Table, PT_Data) 
plot(lmfit2)

# Convert to a data frame via tidy function
lmtable2 <- broom::tidy(lmfit2)

#Results
print(lmtable2)

#Save  
Table2_PD = here("results", "PD_results_table2.rds")
saveRDS(lmtable2, file = "Table2_PD")

############################
#### Second model fit

#Doing the same for FEEDER DATA
lmog_FD<-linear_reg() %>% 
  set_engine("lm")

lmfit_FD<- lmog_FD %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Absence_0_Presence_1~Feeder_number, FD_Data)

lmfit_FD

glimpse(lmfit_FD)

#Way 1
glance(lmfit_FD)

#Way 2
tidy(lmfit_FD)

TableTidy_FD= here("results", "FD_results_TableTidy.rds") #KP Double check path going to MADA Project not folder
saveRDS(lmfit_FD, file = "TableTidy_FD")

plot_summs(lmfit_FD)

PlotSum_FD1= here("results", "FD_results_PlotSum.rds") #KP Double check path going to MADA Project not folder
saveRDS(lmfit_FD, file = "PlotSum1_FD")

#More fitting with some other indicators

lmog_FD2<-linear_reg() %>% 
  set_engine("lm")

lmfit_FD2<- lmog_FD2 %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Absence_0_Presence_1~Feeder_number+Feeder_Type, FD_Data)

lmfit_FD2

glimpse(lmfit_FD2)

#Way 1
glance(lmfit_FD2)

#Way 2
tidy(lmfit_FD2)

TableTidy_FD2= here("results", "FD_results_TableTidy2.rds") #KP Double check path going to MADA Project not folder
saveRDS(lmfit_FD2, file = "TableTidy_FD2")

plot_summs(lmfit_FD2)

PlotSum_FD2= here("results", "FD2_results_PlotSum.rds") #KP Double check path going to MADA Project not folder
saveRDS(lmfit_FD2, file = "PlotSum2_FD")

#Trying ALL predictors of Feeder dataset 
lmog_FD<- linear_reg() %>%
  set_engine("lm")

lmall_FD<- lmog_FD %>% #Training 
  fit(Absence_0_Presence_1~., data=FD_Data)

lmall_FD

tidy(lmall_FD)

plot_summs(lmall_FD, robust=TRUE) #KP NOTE: I need to clean this up so it is a bit more readable, must fullscreen it

PlotSum_FDALL= here("results", "FD_results_PlotSumALL.rds") #KP Double check path going to MADA Project not folder
saveRDS(lmall_FD, file = "PlotSumALL_FD")

#Let's Start with Defining our model 
#Because both out data is continuous, I am going to use a linear regression model

#After the LR, I will create my recipe, create my workflow, create model using the workflow

#And finally try the model on data
```

### 2023-04-08 3. Analysis Script/exploratory_analysis.qmd
Project: Revamp 2 [IP-SA .qmd REDO]

--- 2023-03-18 code/analysis_code/exploratory_analysis.qmd
@@
 ---
-title: "Combined Data Cleaning and Exploratory Analysis Script for Salmonella Environmental Persistence Informs Management Relevant to Avian and Public Health"
+title: "Combined Data Cleaning and Exploratory Analysis Script for Manuscript: Salmonella Environmental Persistence Informs Management Relevant to Avian and Public Health"
 author: "Kimberly M. Perez"
-date: "10/28/2021"
+date: "05/01/2023"
 output: html_document
 ---
+
 # **Data Processing Methodology**
+
 Before attention can be focused on the data analysis portion, removal of unnecessary variables and visualizing data will provide a clearer picture of the data.
@@
 Such packages to download include read_csv(), read_excel(), and here(), each of which can be loaded by utilizing the library() function (e.g., library(here)).
-Next, the datasets from both the Ibis and Feeder trials are loaded into RStudio utilizing the appropriate function (e.g., read_csv() for .csv files and read_excel() for .xlsx files).
+Next, the datasets from both the Ibis Picnic Table trial and Feeder trial are loaded into RStudio utilizing the appropriate function (e.g., read_csv() for .csv files and read_excel() for .xlsx files).
 After data is loaded into R, it is best to get acquainted with the data by utilizing basic functions such as glimpse().
@@
 
-Here, some more packages were loaded that I forsaw using in this section.
+Here, some more packages were loaded that I foresaw using in this section.
 Before visualization began, the glimpse() function was utilized to view the recently cleaned data.
@@
 
-## **Data import and cleaning**
+## **Loading Packages**
 
-## **Initial Package Loading**
 ```{r}
@@
 library(readxl)
-```
-
-## **Picnic Table Data Loading**
-```{r}
-PTD<- read_excel(here("data", "raw_data", "Picnic_Table_Data.xlsx"))
-View(PTD)
-summary(PTD)
-```
-## **Picnic Table Data Wrangling**
-```{r}
-glimpse(PTD)
-
-PTD1<- PTD %>% select(-c(Day, `Shippng Date`, Notes, `STAB TUBE NUMBER`, `ORIGIN OF FECES`, ...13)) #Removing columns
-
-PTD1$`Original PP still intact?`<- ifelse(PTD1$`Original PP still intact?`=="Yes", 1, 0)
-
-PTD2<- PTD1 %>% 
-    mutate(`Original PP still intact?` = ifelse(is.na(`Original PP still intact?`), 0, `Original PP still intact?`)) #NAs to 0 for column `Original PP still intact?`
-
-PTD3<- PTD2 %>% 
-    mutate(`Pos orNeg`= ifelse(is.na(`Pos orNeg`), 0, `Pos orNeg`)) #NAs to 0 for column `Pos orNeg`
-
-glimpse(PTD3) #Current variable names
-
-PTD_clean<- PTD3 %>% #Renaming column names
-  rename(
-    ID= `Sample ID`,
-    Sample_Type= `Sample Type (Pooled or Persistence)`,
-    Salmonella_Positive= `Pos orNeg`,
-    Original_Persistence_Pile= `Original PP still intact?`)
-
-glimpse(PTD_clean)
-```
-
-## **Picnic Table Data Exploration and Visualization**
-```{r}
-#After cleaning I have decided I need to download a few more packages for the data visualization
-
 library(gtsummary) #Tables 
@@
 library(ggplot2) #For basic visualization
+```
 
+## **1a. Loading: Picnic Table Data**
+
+```{r}
+PTD_clean<- readRDS(here::here("2. Clean Data","PTprocessed.rds"))
+View(PTD_clean)
+summary(PTD_clean)
+```
+
+## **2a. Exploration and Visualization: Picnic Table Data**
+
+```{r}
 PTDsumt1<- summary(PTD_clean$Salmonella_Positive) #Quick glimpse of samples, non-specific (e.g., persistence/pool) based on Salmonella positivity
@@
 
-PTDH<- ggplot(PTD_clean, 
+PTDH_0<- ggplot(PTD_clean, 
               aes(fill=Sample_Type, y= Salmonella_Positive, x= Date)) + 
   geom_bar (position="stack", stat="identity") +
-  geom_vline(xintercept= as.numeric (PTD_clean$Date[c(95)]), linetype= "dotted", color="red", lwd=1)
+  geom_vline(xintercept= as.numeric (PTD_clean$Date[c(95)]), linetype= "dotted", color="red", lwd=1) +
+  ggtitle("Positive Salmonella Samples by Quadrant") +
+  labs(y= "Salmonella Positive", x= "Collection Date")
 
-PTDH #Basic Plot: Positive Salmonella Samples by Sample Type via stacked plot, red line denotes decrease, will tap Dr. Hernandez for more information on this decline (e.g., weather, shipping, etc.) 
-ggsave(PTDH, file="Picnic_Table_PosSamp.png")
+PTDH<- PTDH_0 + labs(fill= "Sample Type")
+PTDH
+
+PT_viz<- here::here("4. Results","Picnic_Table_PosSamp.png")
+ggsave(PTDH, file=PT_viz)
 
@@
   geom_bar (position="stack", stat="identity") +
-  geom_vline(xintercept= as.numeric (PTD_clean$Date[c(95)]), linetype= "dotted", color="green", lwd=1)
+  geom_vline(xintercept= as.numeric (PTD_clean$Date[c(95)]), linetype= "dotted", color="green", lwd=1) +
+  ggtitle("Positive Salmonella Samples by Date and Quadrant") +
+  labs(y= "Salmonella Positive", x= "Collection Date")
 
@@
 Hernandez for more information on this decline (e.g., weather, shipping, etc.)
-ggsave(PTD_Quad, file="Picnic_Table_Quad.png")
 
+PT_viz1<- here::here("4. Results","Picnic_Table_Quad.png")
+ggsave(PTD_Quad , file=PT_viz1)
 ```
-## **Feeder Data Loading**
+
+## **1b. Loading: Feeder Data**
+
 ```{r}
 library(readr)
-FD<- read_csv(
-  here("data", "raw_data", "Feeder Data.csv"))
+FD_clean<- readRDS(here::here("2. Clean Data","FDprocessed.rds"))
 
-View(FD)
+View(FD_clean)
 
-summary(FD)
+summary(FD_clean)
 
-str(FD)
+str(FD_clean)
 
-glimpse(FD)
+glimpse(FD_clean)
 ```
 
-## **Feeder Data Wrangling**
-```{r}
-FD1<- FD %>% select(-c(`Absence (0)/Presence (1)`, `...5`)) #Removing columns
+**2b. Exploration and Visualization: Feeder Data**
 
-glimpse(FD1) #Quick glimpse at the data, I can determine if I should wrangle the data a bit more
-
-FD_clean<- FD1 #Data looked clean so I will rename it to use with my data exploration and visualization process below 
-```
-
-**Picnic Table Data Exploration and Visualization**
 ```{r}
@@
               aes(fill=Feeder_Type, y=Absence_0_Presence_1, x= Collection_Date)) + 
-  geom_bar (position="stack", stat="identity")
+  geom_bar (position="stack", stat="identity") + 
+  ggtitle("Absence or Presence of Salmonella by Date based on Feeder Type")+ 
+  labs(x= "Collection Date", y= "Salmonella Absence/Presence")
 
-FD_plot  
-ggsave(FD_plot, file="Feeder_PA.png")
+FD_plot_final<- FD_plot + labs(fill="Feeder Type")
+
+FD_plot_final
+
+FD_viz<- here::here("4. Results","Feeder_PA.png")
+ggsave(FD_plot_final, file=FD_viz)
+
 ```

### 2023-04-09 5. Manuscript/Manuscript.qmd
Revamp Continuation

--- 2023-03-18 products/manuscript/Manuscript.qmd
@@
 ---
-title: "Salmonella Environmental Persistence Informs Management Relevant to Avian and Public Health"
+title: "*Salmonella* Environmental Persistence Informs Management Relevant to Avian and Public Health"
 subtitle: ""
-author: Andreas Handel
+author: Kimberly M. Perez
 date: "`r Sys.Date()`"
@@
     highlight-style: github
-bibliography: ../dataanalysis_template_references.bib
-csl: ../apa.csl
+bibliography: Ref.bib
 ---
@@
 
-*Salmonella enterica* infections are a significant public health threat, responsible for over 93 million annual cases of human illness worldwide.
-In the United States alone, 1.35 million cases of salmonellosis and 420 deaths are reported annually.
+*Salmonella enterica* infections are a significant public health threat, responsible for over 93 million annual cases of human illness worldwide [@Majowicz2010].
+In the United States alone, 1.35 million cases of salmonellosis and 420 deaths are reported annually [@CentersforDiseaseControlandPrevention2021a].
 Most cases of human salmonellosis are caused by food-borne *Salmonella* strains associated with produce or undercooked meat.
-However, a rising subset of human infections are often associated with unidentified environmental exposures or contact with animals.
+However, a rising subset of human infections are often associated with unidentified environmental exposures or contact with animals [@CentersforDiseaseControlandPrevention2016].
 
@@
 in American white ibis is well documented.
-It has been found that ibis can harbor diverse *Salmonella* strains and shed it at a higher prevalence than their natural counterparts.
-Studies have genetically matched the *Salmonella* shed by ibis to human salmonellosis cases in South Florida during the same time period ibis were sampled.
+It has been found that ibis can harbor diverse *Salmonella* strains and shed it at a higher prevalence than their natural counterparts [@Hernandez2016].
+Studies have genetically matched the *Salmonella* shed by ibis to human salmonellosis cases in South Florida during the same time period ibis were sampled [@Hernandez2016].
 
 Similarly, in 2021, following an epidemiological investigation of 29 human cases of *Salmonella enterica* serovar Typhimurium (*S.* Typhimurium), the Centers for Disease Control and Prevention (CDC) documented a relationship between human cases and a concurrent avian salmonellosis outbreak.
-Of the 29 cases, 14 individuals were hospitalized and the outbreak was likely due to contact with bird feeders, sick or dead wild birds, or pets that had contact with wild birds.
+Of the 29 cases, 14 individuals were hospitalized and the outbreak was likely due to contact with bird feeders, sick or dead wild birds, or pets that had contact with wild birds [@CentersforDiseaseControlandPrevention2021].
 
-Persistence of *Salmonella* can be affected by ambient temperature, presence/absence of a biofilm, and environmental nutrient conditions.
+Persistence of *Salmonella* can be affected by ambient temperature, presence/absence of a biofilm, and environmental nutrient conditions [@Will1973; @Davies1995; @Barker2000; @Renner2002].
 Under ideal conditions (e.g.
@@
 If the patches turned magenta, that indicated positive *Salmonella* samples.
-A blue color indicated something other than *Salmonella.*
+A blue color indicated something other than *Salmonella* [@Maurer2015].
 

### 2023-04-09 3. Analysis Script/StatisticalAnalysis.qmd
Revamp: Analysis [IP]

@@
 ```
-
-# **Data Fitting/Statistical Analysis**
+# 2. **Data Fitting/Statistical Analysis: Picnic Table**
 Let's Begin with some Linear Model Fitting; I will try this multiple ways, with several predictors.
@@
 
-TableTidy_PTD<- here::here("4. Results","PTD_TableTidy.png")
-
-ggsave(lmfit_PT, file=TableTidy_PTD)
-
 plot_summs(lmfit_PT)
-
-PlotSum_PD1= here("results", "PD_results_PlotSum.rds") #KP Double check path going to MADA Project not folder
-saveRDS(lmall_PT, file = "PlotSum1_PD")
 
@@
 plot_summs(lmall_PT, robust=TRUE) #KP NOTE: I need to clean this up so it is a bit more readable, must fullscreen it
+```
 
-PlotSum_PD1= here("results", "PD_results_PlotSum.rds") #KP Double check path going to MADA Project not folder
-saveRDS(lmall_PT, file = "PlotSum1_PD")
-
+```{r} 
 #### Model fit another way
@@
 
-# save fit results table  
-Table1_PD= here("results", "PD_results_table.rds") #KP Double check path going to MADA Project not folder
-saveRDS(lmtable1, file = "Table1_PD")
-
 #A different way Part Deux 
@@
 print(lmtable2)
+```
 
-#Save  
-Table2_PD = here("results", "PD_results_table2.rds")
-saveRDS(lmtable2, file = "Table2_PD")
-
-############################
-#### Second model fit
-
+Second model fit
+```{r}
 #Doing the same for FEEDER DATA
@@
 
-TableTidy_FD= here("results", "FD_results_TableTidy.rds") #KP Double check path going to MADA Project not folder
-saveRDS(lmfit_FD, file = "TableTidy_FD")
-
 plot_summs(lmfit_FD)
-
-PlotSum_FD1= here("results", "FD_results_PlotSum.rds") #KP Double check path going to MADA Project not folder
-saveRDS(lmfit_FD, file = "PlotSum1_FD")
 
@@
 
-TableTidy_FD2= here("results", "FD_results_TableTidy2.rds") #KP Double check path going to MADA Project not folder
-saveRDS(lmfit_FD2, file = "TableTidy_FD2")
-
 plot_summs(lmfit_FD2)
-
-PlotSum_FD2= here("results", "FD2_results_PlotSum.rds") #KP Double check path going to MADA Project not folder
-saveRDS(lmfit_FD2, file = "PlotSum2_FD")
 
@@
 plot_summs(lmall_FD, robust=TRUE) #KP NOTE: I need to clean this up so it is a bit more readable, must fullscreen it
+```
+# **IP Analysis KP finish by 4/9!!**
 
-PlotSum_FDALL= here("results", "FD_results_PlotSumALL.rds") #KP Double check path going to MADA Project not folder
-saveRDS(lmall_FD, file = "PlotSumALL_FD")
+#**Time to event**
+```{r}
+library(finalfit)
+library(forcats)
 
-#Let's Start with Defining our model 
-#Because both out data is continuous, I am going to use a linear regression model
+ff_glimpse(PT_Data)
 
-#After the LR, I will create my recipe, create my workflow, create model using the workflow
+PT_Data= PT_Data %>%
+  mutate(
 
-#And finally try the model on data
+  )
 ```
+
+#**Cox regression**
+```{r}
+```

### 2023-04-09 5. Manuscript/Manuscript.qmd
Revamp: Analysis [IP]

@@
 bibliography: Ref.bib
+editor: 
+  markdown: 
+    wrap: 72
 ---
@@
 
-Globally, *Salmonella* is a significant public health threat.
-We evaluated the persistence of *Salmonella* on various surface materials (e.g., picnic tables and bird feeders).
-Our studies suggest that *Salmonella* can have prolonged persistence on shared spaces.
-However, persistence depends on surface material and precautionary measures should be adopted to reduce exposure.
+Globally, *Salmonella* is a significant public health threat. We
+evaluated the persistence of *Salmonella* on various surface materials
+(e.g., picnic tables and bird feeders). Our studies suggest that
+*Salmonella* can have prolonged persistence on shared spaces. However,
+persistence depends on surface material and precautionary measures
+should be adopted to reduce exposure.
 
@@
 
-*Salmonella enterica* infections are a significant public health threat, responsible for over 93 million annual cases of human illness worldwide [@Majowicz2010].
-In the United States alone, 1.35 million cases of salmonellosis and 420 deaths are reported annually [@CentersforDiseaseControlandPrevention2021a].
-Most cases of human salmonellosis are caused by food-borne *Salmonella* strains associated with produce or undercooked meat.
-However, a rising subset of human infections are often associated with unidentified environmental exposures or contact with animals [@CentersforDiseaseControlandPrevention2016].
+*Salmonella enterica* infections are a significant public health threat,
+responsible for over 93 million annual cases of human illness worldwide
+[@Majowicz2010]. In the United States alone, 1.35 million cases of
+salmonellosis and 420 deaths are reported annually
+[@CentersforDiseaseControlandPrevention2021a]. Most cases of human
+salmonellosis are caused by food-borne *Salmonella* strains associated
+with produce or undercooked meat. However, a rising subset of human
+infections are often associated with unidentified environmental
+exposures or contact with animals
+[@CentersforDiseaseControlandPrevention2016].
 
-The American white ibis (Eudocimus albus), forms large nesting colonies in natural wetlands.
-However, due to habitat destruction in Everglades National Park, this gregarious member of the Pelecaniformes, has begun to urbanize parks with abundant anthropgenic food and water resources throughout South Florida.
-Infection with *Salmonella* spp.
-in American white ibis is well documented.
-It has been found that ibis can harbor diverse *Salmonella* strains and shed it at a higher prevalence than their natural counterparts [@Hernandez2016].
-Studies have genetically matched the *Salmonella* shed by ibis to human salmonellosis cases in South Florida during the same time period ibis were sampled [@Hernandez2016].
+The American white ibis (Eudocimus albus), forms large nesting colonies
+in natural wetlands. However, due to habitat destruction in Everglades
+National Park, this gregarious member of the Pelecaniformes, has begun
+to urbanize parks with abundant anthropgenic food and water resources
+throughout South Florida. Infection with *Salmonella* spp. in American
+white ibis is well documented. It has been found that ibis can harbor
+diverse *Salmonella* strains and shed it at a higher prevalence than
+their natural counterparts [@Hernandez2016]. Studies have genetically
+matched the *Salmonella* shed by ibis to human salmonellosis cases in
+South Florida during the same time period ibis were sampled
+[@Hernandez2016].
 
-Similarly, in 2021, following an epidemiological investigation of 29 human cases of *Salmonella enterica* serovar Typhimurium (*S.* Typhimurium), the Centers for Disease Control and Prevention (CDC) documented a relationship between human cases and a concurrent avian salmonellosis outbreak.
-Of the 29 cases, 14 individuals were hospitalized and the outbreak was likely due to contact with bird feeders, sick or dead wild birds, or pets that had contact with wild birds [@CentersforDiseaseControlandPrevention2021].
+Similarly, in 2021, following an epidemiological investigation of 29
+human cases of *Salmonella enterica* serovar Typhimurium (*S.*
+Typhimurium), the Centers for Disease Control and Prevention (CDC)
+documented a relationship between human cases and a concurrent avian
+salmonellosis outbreak. Of the 29 cases, 14 individuals were
+hospitalized and the outbreak was likely due to contact with bird
+feeders, sick or dead wild birds, or pets that had contact with wild
+birds [@CentersforDiseaseControlandPrevention2021].
 
-Persistence of *Salmonella* can be affected by ambient temperature, presence/absence of a biofilm, and environmental nutrient conditions [@Will1973; @Davies1995; @Barker2000; @Renner2002].
-Under ideal conditions (e.g.
-optimal temperature 35-43°C, pH 7-7.5, etc.), *Salmonella* is known to persist in the environment for extended periods of time.
-While *Salmonella* persistence has been examined on soil and poultry litter, there is a lack of understanding of the persistence of *Salmonella* on various surfaces including wood, plastic, and surfaces lined with antimicrobial coatings.
+Persistence of *Salmonella* can be affected by ambient temperature,
+presence/absence of a biofilm, and environmental nutrient conditions
+[@Will1973; @Davies1995; @Barker2000; @Renner2002]. Under ideal
+conditions (e.g. optimal temperature 35-43°C, pH 7-7.5, etc.),
+*Salmonella* is known to persist in the environment for extended periods
+of time. While *Salmonella* persistence has been examined on soil and
+poultry litter, there is a lack of understanding of the persistence of
+*Salmonella* on various surfaces including wood, plastic, and surfaces
+lined with antimicrobial coatings.
 
@@
 
-The data selected for my MADA project is persistence and prevalence data from two trial studies.
-The first trial was conducted in South Florida where picnic tables, under the same pavilion and exposed to conditions known to affect *Salmonella*, were selected, swabbed daily for a week and shipped back to the University of Georgia for *Salmonella* isolation.
-The second trial was conducted in Athens, Georgia, at Whitehall Forest in a modified shade house.
-Bird feeders of different materials (e.g., cedar, plastic, plastic coated, and wood coated) were hung in triplicate and seeded with an isolate of *S.* Typhimurium isolated from a songbird.
+The data selected for my MADA project is persistence and prevalence data
+from two trial studies. The first trial was conducted in South Florida
+where picnic tables, under the same pavilion and exposed to conditions
+known to affect *Salmonella*, were selected, swabbed daily for a week
+and shipped back to the University of Georgia for *Salmonella*
+isolation. The second trial was conducted in Athens, Georgia, at
+Whitehall Forest in a modified shade house. Bird feeders of different
+materials (e.g., cedar, plastic, plastic coated, and wood coated) were
+hung in triplicate and seeded with an isolate of *S.* Typhimurium
+isolated from a songbird.
 
-The data from the first trial in South Florida was collected by an undergraduate student of Dr.
-Sonia Hernandez and provided to me by Dr.
-Hernandez.
-All data from the Feeder Trial was collected by me, Kimberly Perez under the guidance of Dr.
-Sonia Hernandez.
+The data from the first trial in South Florida was collected by an
+undergraduate student of Dr. Sonia Hernandez and provided to me by Dr.
+Hernandez. All data from the Feeder Trial was collected by me, Kimberly
+Perez under the guidance of Dr. Sonia Hernandez.
 
-The Picnic Table Trial data is composed of several worksheets embedded in one excel file.
-After viewing the file (and running it below), there seems to be 192 observations.
-I am not sure how the Picnic Table Trial was collected but hope to visit with Dr.
-Hernandez regarding data collection methods soon.
-Feeder Trial data was collected and noted upon confirmation of *Salmonella* from processed samples.
-There are 144 observations for the Feeder Trial which spanned nearly 2 months.
-For both trials, we were attempting to measure *Salmonella* persistence.
-The Picnic Table Trial collected data in an attempt to measure prevalence as well.
-(Please Note: Aforementioned data was imported under "Data import and cleaning" section.)
+The Picnic Table Trial data is composed of several worksheets embedded
+in one excel file. After viewing the file (and running it below), there
+seems to be 192 observations. I am not sure how the Picnic Table Trial
+was collected but hope to visit with Dr. Hernandez regarding data
+collection methods soon. Feeder Trial data was collected and noted upon
+confirmation of *Salmonella* from processed samples. There are 144
+observations for the Feeder Trial which spanned nearly 2 months. For
+both trials, we were attempting to measure *Salmonella* persistence. The
+Picnic Table Trial collected data in an attempt to measure prevalence as
+well. (Please Note: Aforementioned data was imported under "Data import
+and cleaning" section.)
 
@@
 
-Given the uncertainty of *Salmonella's* persistence on various surface materials, from my data I would like to answer: - What is the prevalence of *Salmonella* over time?
-- How long does *Salmonella* persist on different surface materials (e.g., wood, plastic, and those with an antimicrobial coating) exposed to various climatic factors (e.g., rain, sun, changing temperatures)
+Given the uncertainty of *Salmonella's* persistence on various surface
+materials, questions such as:
 
-From these questions, I hope to better inform the public and public health practitioners on safer feeder handling methods and precautions to take when frequenting public locations where humans and wildlife (e.g., ibis) interact.
+-   What is the prevalence of *Salmonella* over time?
+
+-   How long does *Salmonella* persist on different surface materials
+    (e.g., wood, plastic, and those with an antimicrobial coating)?
+
+-   Is there a greater risk of acquiring *Salmonella* from objects that
+    both humans and animals use (e.g. picnic tables in parks)?
+
+From these questions, we hope to better inform the public and public
+health practitioners on safer feeder handling methods and precautions to
+take when frequenting public locations where humans and wildlife (e.g.,
+ibis) interact.
 
@@
 
-The ibis picnic table trial was conducted at Dreher Park, a public park in Palm Beach County frequented by white ibis who are known to carry *Salmonella*.
-Because humans and ibis both utilize the wooden picnic tables at the park, three tables under a single pavilion with similar environmental conditions (e.g., exposure to sunlight, rainfall, etc.) were selected.
-All tables were divided into four quadrants (e.g., tabletop plus bench) and ibis were encouraged to feed on the picnic tables until they deficated.
-Pictures were taken of every quadrant daily to facilitate the counting of new fecal piles.
-If feces were not deposited in a quadrant, fresh feces from the concrete pad was collected and placed onto the quadrant.
-To determine persistence, one large fecal pile from each quadrant was selected as the "persistence pile" and monitored and sampled daily for 8 days.
-To determine table prevalence, a pooled sample was taken daily from every quadrant by swabbing the surface picnic table surface with a sampling sponge, avoiding the designated "persistence pile".
-To maintain cleanliness for other park patrons, new piles were not smeared with the sampling sponge.
-Samples were placed into 20 mL of dulcitol selenite and maintained at room temperature prior to being shipped to the University of Georgia (UGA), every 2-4 days, for processing.
+The ibis picnic table trial was conducted at Dreher Park, a public park
+in Palm Beach County frequented by white ibis who are known to carry
+*Salmonella*. Because humans and ibis both utilize the wooden picnic
+tables at the park, three tables under a single pavilion with similar
+environmental conditions (e.g., exposure to sunlight, rainfall, etc.)
+were selected. All tables were divided into four quadrants (e.g.,
+tabletop plus bench) and ibis were encouraged to feed on the picnic
+tables until they deficated. Pictures were taken of every quadrant daily
+to facilitate the counting of new fecal piles. If feces were not
+deposited in a quadrant, fresh feces from the concrete pad was collected
+and placed onto the quadrant. To determine persistence, one large fecal
+pile from each quadrant was selected as the "persistence pile" and
+monitored and sampled daily for 8 days. To determine table prevalence, a
+pooled sample was taken daily from every quadrant by swabbing the picnic table surface with a sampling sponge, avoiding the designated "persistence pile". To maintain cleanliness for other park
+patrons, new piles were not smeared with the sampling sponge. Samples
+were placed into 20 mL of dulcitol selenite and maintained at room
+temperature prior to being shipped to the University of Georgia (UGA),
+every 2-4 days, for processing.
 
@@
 
-Our Feeder trial aimed to assess the persistence of *Salmonella* on four types of feeders (cedar, plastic, cedar + antimicrobial coating, and plastic + antimicrobial coating) in Georgia.
-18 g of fresh chicken feces was collected, homogenized, 1 g was aliquoted into individual 15 mL test tubes, then sterilized through autoclaving.
-The individual alioquots were then inoculated with 1.5x10\^7 mL of a *Salmonella* Typhimurium isolate responsible for a previous avian salmonellosis outbreak.
-Samples were then transported to a modified shade house at Whitehall Forest where four feeder types, represented in triplicate and randomized, were seeded with 1 g of inoculated feces and 0.5 mL of sterile physiologic saline.
-Feeders were then swabbed daily for one week, every other day for one week, and weekly for an additional week.
-Swabs were embedded into 10 mL of Dulcitol Selenite broth and transported back to UGA for processing.
+Our Feeder trial aimed to assess the persistence of *Salmonella* on four
+types of feeders (cedar, plastic, cedar + antimicrobial coating, and
+plastic + antimicrobial coating) in Georgia. 18 g of fresh chicken feces
+was collected, homogenized, 1 g was aliquoted into individual 15 mL test
+tubes, then sterilized through autoclaving. The individual alioquots
+were then inoculated with 1.5x10\^7 mL of a *Salmonella* Typhimurium
+isolate responsible for a previous avian salmonellosis outbreak. Samples
+were then transported to a modified shade house at Whitehall Forest
+where four feeder types, represented in triplicate and randomized, were
+seeded with 1 g of inoculated feces and 0.5 mL of sterile physiologic
+saline. Feeders were then swabbed daily for one week, every other day
+for one week, and weekly for an additional week. Swabs were embedded
+into 10 mL of Dulcitol Selenite broth and transported back to UGA for
+processing.
 
@@
 
-After collection, 100 uL of the sample vortexed in dulcitol selentite was inoculated into 10 mL of Rappaport Vassiliadis (RV) broth.
-The broth was incubated for 24 hours at 37c.
-The RV broth was then streaked onto XLT-4 agar plates and incubated at 37c for 24 hours, then an additional 24 hours on the lab bench at room temperature.
-If colonies morphometrically consistent with *Salmonella* were present, one colony per plate was chosen and a stab in Luria Broth (LB) agar was created.
-To confirm the presence of *Salmonella*, patch plates were made on ChromAgar using the stabbed colonies.
-If the patches turned magenta, that indicated positive *Salmonella* samples.
-A blue color indicated something other than *Salmonella* [@Maurer2015].
+After collection, 100 uL of the sample vortexed in dulcitol selentite
+was inoculated into 10 mL of Rappaport Vassiliadis (RV) broth. The broth
+was incubated for 24 hours at 37c. The RV broth was then streaked onto
+XLT-4 agar plates and incubated at 37c for 24 hours, then an additional
+24 hours on the lab bench at room temperature. If colonies
+morphometrically consistent with *Salmonella* were present, one colony
+per plate was chosen and a stab in Luria Broth (LB) agar was created. To
+confirm the presence of *Salmonella*, patch plates were made on
+ChromAgar using the stabbed colonies. If the patches turned magenta,
+that indicated positive *Salmonella* samples. A blue color indicated
+something other than *Salmonella* [@Maurer2015].
 
@@
 
-The data from the first trial in South Florida was collected by an undergraduate student of Dr.
-Sonia Hernandez and provided to me by Dr.
-Hernandez.
-All data from the Feeder Trial was collected by me, Kimberly Perez under the guidance of Dr.
-Sonia Hernandez.
+The data from the Picnic Table Trial in South Florida was collected by an
+undergraduate student of Dr. Sonia M. Hernandez and provided to me by Dr.
+Hernandez. All data from the Feeder Trial was collected by me, Kimberly
+Perez under the guidance of Dr. Sonia Hernandez.
 
@@
 
-Before attention can be focused on the data analysis portion, removal of unnecessary variables and visualizing data will provide a clearer picture of the data.
-This section will focus on data cleaning and wrangling.
-First, both datasets must be loaded.
-Prior to loading any data into RStudio, ensure that a new Quarto or RMarkdown File is established and linked to the preferred working directory.
-Basic packages need to be loaded to read in the data.
-Such packages to download include read_csv(), read_excel(), and here(), each of which can be loaded by utilizing the library() function (e.g., library(here)).
-Next, the datasets from both the Ibis and Feeder trials are loaded into RStudio utilizing the appropriate function (e.g., read_csv() for .csv files and read_excel() for .xlsx files).
-After data is loaded into R, it is best to get acquainted with the data by utilizing basic functions such as glimpse().
-Glimpse provides me with insight into my data such as column names, number of rows and how each row is classified (e.g., character or numeric).
-After viewing the data I can work on cleaning the data by removing unnecessary variables.
-I would like to make the data a little more uniform and easier to work with.
-With that, I will change "Yes" to 1, and *NAs* to 0.
-I believe this will help me further in the analysis process.
-Given the variablility in the column names, I will then clean the names up to make them easier to include in my code as I move forward.
+Before attention can be focused on the data analysis portion, removal of
+unnecessary variables and visualizing data will provide a clearer
+picture of the data. This section will focus on data cleaning and
+wrangling. First, both datasets must be loaded. Prior to loading any
+data into RStudio, ensure that a new Quarto or RMarkdown File is
+established and linked to the preferred working directory. Basic
+packages need to be loaded to read in the data. Such packages to
+download include read_csv(), read_excel(), and here(), each of which can
+be loaded by utilizing the library() function (e.g., library(here)).
+Next, the datasets from both the Ibis and Feeder trials are loaded into
+RStudio utilizing the appropriate function (e.g., read_csv() for .csv
+files and read_excel() for .xlsx files). After data is loaded into R, it
+is best to get acquainted with the data by utilizing basic functions
+such as glimpse(). Glimpse provides users with insight into the data
+such as column names, number of rows and how each row is classified
+(e.g., character or numeric). After viewing the data, cleaning the data
+by removing unnecessary variables can occur.
+
+Ensuring the data is uniform by removing unnecessary variables and renaming columns will assist in the analysis process downstream.
+To begin, "Yes" responses should be coded to produce a 1, while *NAs* would produce a 0.
+Given the variability in the column names, standardizing or renaming them a preferred name I will streamline future coding or analysis.
 With all of that, it is finally time to visualize the data by making some tables and/or graphs.
 
-Here, some more packages were loaded that I forsaw using in this section.
-Before visualization began, the glimpse() function was utilized to view the recently cleaned data.
-Given the uncertainty of *Salmonella's* persistence on various surface materials and the information these trials hoped to provide, the summary() function was utilized to provide a quick glimpse of basic result summaries for positive *Salmonella* samples and if the pile present was the original pile or not.
-Next, I continued to explore these two variables of interest by plotting each using a stacked bar graph via ggplot().
+More packages were loaded that would assist in the visualization process before begining this process. Before visualization began, the glimpse() function was utilized
+to view the recently cleaned data. Given the uncertainty of
+*Salmonella's* persistence on various surface materials and the
+information these trials hoped to provide, the summary() function was
+utilized to provide a quick glimpse of basic result summaries for
+positive *Salmonella* samples and if the pile present was the original
+pile or not. Exploration continued on these two variables of
+interest by plotting each using a stacked bar graph via ggplot().
 
-I followed the same workflow for the Feeder Data, however, this dataset required less wrangling and did not provide as many exploratory options as the main focus was on feeder type, date, and the presence/absence of *Salmonella* for each day the feeders were sampled.
+The same workflow for the Feeder Data was utilized, however, this dataset
+required less wrangling and did not provide as many exploratory options
+as the main focus was on feeder type, date, and the presence/absence of
+*Salmonella* for each day the feeders were sampled.
 
@@
 
-![Positive Salmonella Samples by Type and Date](images/Picnic_Table_PosSamp-04.png) Positive *Salmonella* Samples by Sample Type and Date
+![Positive Salmonella Samples by Type and
+Date](images/Picnic_Table_PosSamp-04.png) Positive *Salmonella* Samples
+by Sample Type and Date
 
-![Positive *Salmonella* Samples by Quadrant and Date](images/Picnic_Table_Quad-01.png)
+![Positive *Salmonella* Samples by Quadrant and
+Date](images/Picnic_Table_Quad-01.png)
 
@@
 
-![Positive *Salmonella* Samples by feeder type and date.](images/Picnic_Table_PosSamp.png)
+![Positive *Salmonella* Samples by feeder type and
+date.](images/Picnic_Table_PosSamp.png)
 
 {{< pagebreak >}}
-
-# **Future Project PARTS for KP to populate**
 
@@
 
+### **Persistence**
+On Day 1 of the study, four fecal piles per table (one per quadrant, n = 12) were designated for tracking *Salmonella* persistence.
+**Salmonella** was isolated from three of the 12 designated persistence study piles on Day 1, allowing the persistence to be tracked on these three piles.
+Persistence, expressed as the number of days that a positive fecal pile on Day 1 remained positive, ranged from 1 to 4 days with an average persistence of 3 days.
+The probability of an already positive sample to remain positive the next day declined consistently throughout the study period.
+
+### **Prevalence**
 Of 96 deposited feces, we obtained 27% positive.
-*Salmonella* persisted for an average of 3 days, and maximum of 4 days on
-wooden picnic tables.
+*Salmonella* persisted for an average of 3 days, and maximum of 4 days on wooden picnic tables.
+*Salmonella* was isolated from all tables, at least once.
+While there appears to be a significant positive correlation between the number of new fecal piles and the probability of yielding a positive pooled sample.
+The probabilities of a positive sample varied throughout the week, with day 4 having the lowest probability of positive *Salmonella* isolation and day 7 having the most error.
 
-## *Feeder Trial*
-We confirmed successful inoculation of all
-feeders with *Salmonella* (presence on day post inoculation \[DPI\]
-0). Further, results indicate that *Salmonella* may persist on
-some feeders up to DPI 5 and current analysis indicates that *Salmonella* persisted
-on both coated and uncoated plastic feeders at a higher rate than wood and wood
-coated. Figure 2 depicts the percent of samples positive based on feeder type
-throughout the study period. While plastic and plastic coated feeders had an
-overall higher percent positivity, a spike in percent positivity in wood feeders occurred on DPI 5 after a rain event. Further analyses should be conducted to determine whether environmental
-conditions (e.g., rain event) had a significant influence on persistence.
+## **Feeder Trial**
+We confirmed successful inoculation of all feeders with *Salmonella*
+(presence on day post inoculation \[DPI\] 0). Further, results indicate
+that *Salmonella* may persist on some feeders up to DPI 5 and current
+analysis indicates that *Salmonella* persisted on both coated and
+uncoated plastic feeders at a higher rate than wood and wood coated.
+Figure 2 depicts the percent of samples positive based on feeder type
+throughout the study period. While plastic and plastic coated feeders
+had an overall higher percent positivity, a spike in percent positivity
+in wood feeders occurred on DPI 5 after a rain event. Further analyses
+should be conducted to determine whether environmental conditions (e.g.,
+rain event) had a significant influence on persistence.
 
-Our samples persisted for an average of 2 days (feeders), 3 days (picnic tables) with 4 (picnic tables)
-and 5 (feeders) being the maximum number of days persisted. The varied
-persistence during both trials could be as a result of a lack of moisture, the
-fluctuation in ambient temperature during sampling, and water activity. Our studies
-suggest that *Salmonella* can have prolonged persistence on shared spaces,
-persistence depends on surface material, and precautionary measures (i.e.,
-increased hygiene) should be adopted to reduce human exposure.
+Our samples persisted for an average of 2 days (feeders), 3 days (picnic
+tables) with 4 (picnic tables) and 5 (feeders) being the maximum number
+of days persisted. The varied persistence during both trials could be
+a result of a lack of moisture, the fluctuation in ambient temperature
+during sampling, and water activity. Our studies suggest that
+*Salmonella* can have prolonged persistence on shared spaces,
+persistence depends on surface material, and precautionary measures
+(i.e., increased hygiene) should be adopted to reduce human exposure.
 
@@
 
-After the modules following Part 1, I am leaning towards running a GLM for both given the binary responses for *Salmonella* persistence (e.g., presence v.
-absence).
-I need to meet with Dr.
-Hernandez to clear up some questions I have on the picnic table data, specifically regarding the introduction of new feces.
-I am curious if I should include other data into the mix to make my analysis more robust.
-At the moment, the variables of interest for both are limited (e.g., two for the picnic table data, and one primary variable of interest for the feeder trial).
+After the modules following Part 1, I am leaning towards running a GLM
+for both given the binary responses for *Salmonella* persistence (e.g.,
+presence v. absence). I need to meet with Dr. Hernandez to clear up some
+questions I have on the picnic table data, specifically regarding the
+introduction of new feces. I am curious if I should include other data
+into the mix to make my analysis more robust. At the moment, the
+variables of interest for both are limited (e.g., two for the picnic
+table data, and one primary variable of interest for the feeder trial).
 

### 2023-04-24 3. Analysis Script/StatisticalAnalysis.qmd
Analysis Edits continued...

@@
 ```{r}
+#KP needs to edit this! Not running as expected....
+
 library(finalfit)
 library(forcats)
+library(survival)
 
-ff_glimpse(PT_Data)
+glimpse(PT_Data)
 
-PT_Data= PT_Data %>%
+as.Date(PT_Data$Date, format= "%Y-%M-%D")
+
+PT_Data$Date<- as.numeric(PT_Data$Date)
+
+PT_Data = PT_Data %>% 
   mutate(
+    Salmonella_Positive_os = ifelse(Salmonella_Positive == 0,1,0),
+      status_crr= case_when(
+        `Salmonella_Positive` == 1~0,
+        `Salmonella_Positive` == 1~1))
 
-  )
+dependent_crr= "Surv(Date, status_crr)"
+explanatory= c("Table", "Quadrant", "Sample_Type")
+
+PT_Data %>% finalfit(dependent_crr, explanatory)
+
+PT_Data %>% 
+   finalfit(dependent_crr, explanatory, add_dependent_label = FALSE) %>% 
+    rename("Overall survival" = label) %>% 
+    rename(" " = levels) %>% 
+    rename("  " = all) 
 ```
 
-#**Cox regression**
+#**Cox regression: Picnic Table Data**
 ```{r}
+
+as.Date(PT_Data$Date, format= "%Y-%M-%D")
+
+PT_Data$Date<- as.numeric(PT_Data$Date)
+
+coxph(Surv(Date, Salmonella_Positive) ~ Sample_Type, PT_Data)
+
+coxph(Surv(Date, Salmonella_Positive) ~ Sample_Type, PT_Data) %>%
+  tbl_regression(exp=TRUE)
 ```
+#**Cox regression: Feeder Data**
+```{r}
+library(lubridate)
+
+#Date needs to be fixed in this DS
+
+FD_Data <- FD_Data %>%
+  mutate(Collection_Date=lubridate::dmy(Collection_Date))
+
+FD_Data$Collection_Date<- as.numeric(FD_Data$Collection_Date)
+
+coxph(Surv(Collection_Date, Absence_0_Presence_1 ~ Sample_Type, FD_Data))
+
+coxph(Surv(Date, Salmonella_Positive) ~ Sample_Type, PT_Data) %>%
+  tbl_regression(exp=TRUE)
+```

### 2023-04-26 3. Analysis Script/StatisticalAnalysis.qmd
Edits Continued (Analysis)

@@
 library(broom) #for cleaning up output from lm()
+library(dotwhisker)
 library(here) #for data loading/saving
@@
 lmfit_PT<- lmog %>% #Training; outcome: Salmonella Pos and Main Predictor Table
-  fit(Salmonella_Positive~Table, PT_Data)
+  fit(Salmonella_Positive~Date, PT_Data)
 
@@
 
+#Checking model performance: Picnic Table
+check_model(lmfit_PT$fit)
+
+tidy(lmfit_PT) %>%
+    dwplot(dot_args = list(size = 2, color = "coral1"),
+         whisker_args = list(color = "coral1"),
+         vline = geom_vline(xintercept = 0, color = "darkred", linetype = 4))
+
 #Fitting other specific indicators
@@
 
+model_lm<- linear_reg()
+
 lmfit_PT1<- lmog_1 %>% #Training; outcome: Salmonella Pos and Main Predictor Table
-  fit(Salmonella_Positive~Original_Persistence_Pile+Table, PT_Data)
+  fit(Salmonella_Positive~Sample_Type+Date, PT_Data)
 
@@
 
+#Checking model performance: Picnic Table
+check_model(lmfit_PT1$fit)
+
+tidy(lmfit_PT1) %>%
+    dwplot(dot_args = list(size = 2, color = "coral1"),
+         whisker_args = list(color = "coral1"),
+         vline = geom_vline(xintercept = 0, color = "darkred", linetype = 4))
+
 #Trying ALL predictors of Picnic Table dataset 
@@
 
-plot_summs(lmall_PT, robust=TRUE) #KP NOTE: I need to clean this up so it is a bit more readable, must fullscreen it
+#Do not want all predictors...
+
+#Checking model performance: Picnic Table
+check_model(lmall_PT$fit)
 ```
@@
 
+#Checking model performance: FEEDER
+check_model(lmfit_FD$fit)
+
 #More fitting with some other indicators
@@
 ```
-# **IP Analysis KP finish by 4/9!!**
+# **Generalized Linear Model: Picnic Table** 
+```{r}
+lmog<-linear_reg() %>% 
+  set_engine("glm")
 
+lmfit_PT_glm<- lmog %>% #Training; outcome: Salmonella Pos and Main Predictor Table
+  fit(Salmonella_Positive~Table, PT_Data)
+
+lmfit_PT_glm
+
+glimpse(lmfit_PT_glm)
+
+#Way 1
+glance(lmfit_PT_glm)
+
+#Way 2
+tidy(lmfit_PT_glm)
+
+plot_summs(lmfit_PT_glm)
+
+#Fitting other specific indicators
+lmog_glm1<-linear_reg() %>% 
+  set_engine("glm")
+
+lmfit_PTglm1<- lmog_glm1 %>% #Training; outcome: Salmonella Pos and Main Predictor Table
+  fit(Salmonella_Positive~Original_Persistence_Pile+Table, PT_Data)
+
+lmfit_PTglm1
+
+glimpse(lmfit_PTglm1)
+
+#Way 1
+glance(lmfit_PTglm1)
+
+#Way 2
+tidy(lmfit_PTglm1)
+
+plot_summs(lmfit_PTglm1)
+
+#Fitting other specific indicators
+lmog_glm2<-linear_reg() %>% 
+  set_engine("glm")
+
+lmfit_PTglm2<- lmog_glm2 %>% #Training; outcome: Salmonella Pos and Main Predictor Table
+  fit(Salmonella_Positive~Quadrant+Table, PT_Data)
+
+lmfit_PTglm2
+
+glimpse(lmfit_PTglm2)
+
+#Way 1
+glance(lmfit_PTglm2)
+
+#Way 2
+tidy(lmfit_PTglm2)
+
+plot_summs(lmfit_PTglm2)
+
+#Trying ALL predictors of Picnic Table dataset 
+lmog_PTGLM<- linear_reg() %>%
+  set_engine("glm")
+
+lmall_PTGLM<- lmog_PTGLM %>% #Training 
+  fit(Salmonella_Positive~., data=PT_Data)
+
+lmall_PTGLM
+
+tidy(lmall_PTGLM)
+```
+# **Generalized Linear Model: Picnic Table (Persistence & Prevalence)** 
+```{r}
+PT_GLM_PER <- glm(Salmonella_Positive ~ Date, data = PT_Data, family = binomial)
+
+PT_GLM_PER
+
+PT_GLM_PREV <- glm(Salmonella_Positive ~ Sample_Type + as.factor(Date) + Table, data= PT_Data, family = binomial)
+
+PT_GLM_PREV
+```
+# **Generalized Linear Model: Feeder Data**
+```{r}
+lmog_FD<-linear_reg() %>% 
+  set_engine("glm")
+
+lmfit_FD_glm<- lmog_FD %>% #Training; outcome: Salmonella Pos and Main Predictor Table
+  fit(Absence_0_Presence_1~Feeder_number, FD_Data)
+
+lmfit_FD_glm
+
+glimpse(lmfit_FD_glm)
+
+#Way 1
+glance(lmfit_FD_glm)
+
+#Way 2
+tidy(lmfit_FD_glm)
+
+plot_summs(lmfit_FD_glm)
+
+#Fitting other specific indicators
+lmog_FD_GLM<-linear_reg() %>% 
+  set_engine("glm")
+
+lmfit_FD_GLM<- lmog_FD_GLM %>% #Training; outcome: Salmonella Pos and Main Predictor Table
+  fit(Absence_0_Presence_1~Feeder_number+Feeder_Type, FD_Data)
+
+lmfit_FD_GLM
+
+glimpse(lmfit_FD_GLM)
+
+#Way 1
+glance(lmfit_FD_GLM)
+
+#Way 2
+tidy(lmfit_FD_GLM)
+
+plot_summs(lmfit_FD_GLM)
+
+#Trying ALL predictors of Picnic Table dataset 
+lmog_FDGLM_all<- linear_reg() %>%
+  set_engine("glm")
+
+lmall_FDGLM_all1<- lmog_FDGLM_all %>% #Training 
+  fit(Absence_0_Presence_1~., FD_Data)
+
+lmall_FDGLM_all1
+
+tidy(lmall_FDGLM_all1)
+```
+
+# **Generalized Linear Model: Feeder Data (Persistence & Prevalence)**
+```{r}
+FD_GLM_PER <- glm(Absence_0_Presence_1 ~ Collection_Date, data = FD_Data, family = binomial)
+
+FD_GLM_PER
+
+FD_GLM_PREV <- glm(Absence_0_Presence_1 ~ Feeder_Type + as.factor(Collection_Date), data= FD_Data, family = binomial)
+
+FD_GLM_PREV
+```
 #**Time to event**
 ```{r}
-#KP needs to edit this! Not running as expected....
-
 library(finalfit)
@@
 ```{r}
-
-as.Date(PT_Data$Date, format= "%Y-%M-%D")
-
 PT_Data$Date<- as.numeric(PT_Data$Date)
@@
 
-#Date needs to be fixed in this DS
-
-FD_Data <- FD_Data %>%
-  mutate(Collection_Date=lubridate::dmy(Collection_Date))
-
 FD_Data$Collection_Date<- as.numeric(FD_Data$Collection_Date)
 
-coxph(Surv(Collection_Date, Absence_0_Presence_1 ~ Sample_Type, FD_Data))
+coxph(Surv(Collection_Date, Absence_0_Presence_1) ~ Feeder_Type, FD_Data)
 
-coxph(Surv(Date, Salmonella_Positive) ~ Sample_Type, PT_Data) %>%
+coxph(Surv(Collection_Date, Absence_0_Presence_1) ~ Feeder_Type, FD_Data) %>%
   tbl_regression(exp=TRUE)

### 2023-04-26 3. Analysis Script/exploratory_analysis.qmd
Edits Continued (Analysis)

@@
   ggtitle("Positive Salmonella Samples by Date and Quadrant") +
-  labs(y= "Salmonella Positive", x= "Collection Date")
+  labs(y= "Salmonella Positive", x= "Collection Date") +
 
@@
 
-FD_plot_final<- FD_plot + labs(fill="Feeder Type")
+FD_plot_final<- FD_plot + labs(fill="Feeder Type") + scale_fill_discrete(labels=c('Plastic', 'Plastic Coated', 'Wood', 'Wood Coated'))
 

### 2023-04-26 Project_Review_Template.qmd

# Overview

Title of project:Salmonella* Environmental Persistence Informs Management Relevant to Avian and Public Health

Name of project author(s): Kimberly M. Perez

Name of project reviewer: Connor Ross

# Specific project content evaluation
Evaluate the different parts of the project by filling in the sections below.

## Background, Context and Motivation
How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

Beautifully written introduction and description of data/data source(s). The writing in these sections were clear and succinct while providing plenty of background regarding the existence and context of existing work and why the project was undertaken.

### Summary assessment (PICK ONE, DELETE THE OTHERS)

* strong contextualization and motivation

## Question description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

Questions/hypotheses to be addressed in this project are clear and well formed. Questions relate to the data in their dataset(s).

### Summary assessment

* question/hypotheses fully clear

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

The data and the methods from which the data was collected are well described and give the reader a good understanding of the context in which the data was collected and a lense from which the analysis should be interpreted.

### Summary assessment

* source and overall structure of data well explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

Data was adequately cleaned and formatted for meaningful analysis. Exploratory analysis. Exploratory analysis results were meaningful to the reader and provided a good understanding and summary of the data in the dataset(s). I would consider providing a summary table of the dataset(s). This can be easily done using the gtsummary package with tbl_summary().

### Summary assessment

* essentially no weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

Analysis methods were appropriate for the data and appear to have been performed properly without errors. Performance measures were taken from built models and explained the results well.

### Summary assessment

* strong and reasonable analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

Overall the figures/results presented are easy to read and understand, however, on my end of things, the figures in the manuscript do not have axis titles or labels and do not match the figures in the StatisticalAnalysis.qmd or exploratory_analysis.qmd. Reloading the images to refresh them may be necessary. Outside of that, the figures are of high quality and explain the results well.

### Summary assessment

* results are very well presented

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

Sections 5.5 Full analysis, 6. Discussion, 6.1 Summary and Interpretation, 6.2 Strengths and Limitations, and 6.3 Conclusions are missing from the Manuscript.qmd and Manuscript.docx.

### Summary assessment

* minor parts wrong, missing or unclear

## Further comments

Awesome job! I can tell you put a lot of thought and effort into this project and it turned out really well.

# Overall project content evaluation
Evaluate overall features of the project  by filling in the sections below.

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

Overall the project is well structured, however, some cleaning is required to remove unused files and make the organization more clear. Further updating the README.md is necessary.

### Summary assessment

* well structured

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

The project was well documented in coding files as well as the manuscript. Adequate information was provided to understand the analysis and the decisions made by the author.

### Summary assessment

* fully and well documented

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

The results are fully reproducible, however, documentation (README.md) needs to be updated.Further, there was an error ("Error: object 'PTD_Quad' not found") in line 61 of the exploratory_analysis.qmd file. Neither of which was detrimental to reproducibility and did not require manual intervention.

### Summary assessment

* fully reproducible without issues

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

The author demonstrated a deep understanding of the subject matter and the analysis through the manuscript and the code (comments included) despite sections 5.5 and 6.0 - 6.3 of the manuscript being incomplete/missing. Code and write-up was well thought out and well worded.

### Summary assessment

* strong level of thorougness

## Further comments

Awesome job! I enjoyed reading about your project and look forward to its finalization.

### 2023-04-26 3. Analysis Script/StatisticalAnalysis.qmd
Auto stash before merge of "master" and "origin/master"

@@
 
-lmfit_PT<- lmog %>% #Training; outcome: Salmonella Pos and Main Predictor Table
+model_lm<- linear_reg()
+
+lmfit_PT<- model_lm %>% #Training; outcome: Salmonella Pos and Main Predictor Table
   fit(Salmonella_Positive~Date, PT_Data)
@@
          vline = geom_vline(xintercept = 0, color = "darkred", linetype = 4))
+```
 
+```{r}
 #Fitting other specific indicators
@@
 
-model_lm<- linear_reg()
+model_lm1<- linear_reg()
 
-lmfit_PT1<- lmog_1 %>% #Training; outcome: Salmonella Pos and Main Predictor Table
-  fit(Salmonella_Positive~Sample_Type+Date, PT_Data)
+lmfit_PT1<- model_lm1 %>% #Training; outcome: Salmonella Pos
+  fit(Salmonella_Positive~Sample_Type, PT_Data)
 
 lmfit_PT1
+
+tidy(lmfit_PT1)
 
@@
          vline = geom_vline(xintercept = 0, color = "darkred", linetype = 4))
+```
 
+```{r}
 #Trying ALL predictors of Picnic Table dataset 
@@
   fit(Salmonella_Positive~Quadrant+Table, PT_Data) 
+```
 
+```{r}
 lmfit_PTglm2
@@
 ```
+
 # **Generalized Linear Model: Picnic Table (Persistence & Prevalence)** 
@@
 ```
+
 # **Generalized Linear Model: Feeder Data**
@@
 ```
-#**Time to event**
+# **Time to event**
 ```{r}
@@
 
-as.Date(PT_Data$Date, format= "%Y-%M-%D")
-
 PT_Data$Date<- as.numeric(PT_Data$Date)
@@
   mutate(
-    Salmonella_Positive_os = ifelse(Salmonella_Positive == 0,1,0),
-      status_crr= case_when(
-        `Salmonella_Positive` == 1~0,
-        `Salmonella_Positive` == 1~1))
+      salmonella_crr= case_when(
+        Salmonella_Positive ==  0~1
+        Salmonella_Positive == 1~1
+        TRUE ~ 1))
 
@@
     rename("  " = all)
+
+survfit(Surv(Date, Salmonella_Positive)~1, data=PT_Data)
+
+PT_Data %>%
+  filter(status==1) %>%
+  summarize(median_surv=median(Date))
 ```

### 2023-04-26 5. Manuscript/Manuscript.qmd
Analysis Edits continued...
Auto stash before merge of "master" and "origin/master"

@@
 monitored and sampled daily for 8 days. To determine table prevalence, a
-pooled sample was taken daily from every quadrant by swabbing the picnic table surface with a sampling sponge, avoiding the designated "persistence pile". To maintain cleanliness for other park
-patrons, new piles were not smeared with the sampling sponge. Samples
-were placed into 20 mL of dulcitol selenite and maintained at room
-temperature prior to being shipped to the University of Georgia (UGA),
-every 2-4 days, for processing.
+pooled sample was taken daily from every quadrant by swabbing the picnic
+table surface with a sampling sponge, avoiding the designated
+"persistence pile". To maintain cleanliness for other park patrons, new
+piles were not smeared with the sampling sponge. Samples were placed
+into 20 mL of dulcitol selenite and maintained at room temperature prior
+to being shipped to the University of Georgia (UGA), every 2-4 days, for
+processing.
 
@@
 
-The data from the Picnic Table Trial in South Florida was collected by an
-undergraduate student of Dr. Sonia M. Hernandez and provided to me by Dr.
-Hernandez. All data from the Feeder Trial was collected by me, Kimberly
-Perez under the guidance of Dr. Sonia Hernandez.
+The data from the Picnic Table Trial in South Florida was collected by
+an undergraduate student of Dr. Sonia M. Hernandez and provided to me by
+Dr. Hernandez. All data from the Feeder Trial was collected by me,
+Kimberly Perez under the guidance of Dr. Sonia Hernandez.
 
@@
 
-Ensuring the data is uniform by removing unnecessary variables and renaming columns will assist in the analysis process downstream.
-To begin, "Yes" responses should be coded to produce a 1, while *NAs* would produce a 0.
-Given the variability in the column names, standardizing or renaming them a preferred name I will streamline future coding or analysis.
-With all of that, it is finally time to visualize the data by making some tables and/or graphs.
+Ensuring the data is uniform by removing unnecessary variables and
+renaming columns will assist in the analysis process downstream. To
+begin, "Yes" responses should be coded to produce a 1, while *NAs* would
+produce a 0. Given the variability in the column names, standardizing or
+renaming them a preferred name I will streamline future coding or
+analysis. With all of that, it is finally time to visualize the data by
+making some tables and/or graphs.
 
-More packages were loaded that would assist in the visualization process before begining this process. Before visualization began, the glimpse() function was utilized
-to view the recently cleaned data. Given the uncertainty of
-*Salmonella's* persistence on various surface materials and the
-information these trials hoped to provide, the summary() function was
-utilized to provide a quick glimpse of basic result summaries for
-positive *Salmonella* samples and if the pile present was the original
-pile or not. Exploration continued on these two variables of
-interest by plotting each using a stacked bar graph via ggplot().
+More packages were loaded that would assist in the visualization process
+before begining this process. Before visualization began, the glimpse()
+function was utilized to view the recently cleaned data. Given the
+uncertainty of *Salmonella's* persistence on various surface materials
+and the information these trials hoped to provide, the summary()
+function was utilized to provide a quick glimpse of basic result
+summaries for positive *Salmonella* samples and if the pile present was
+the original pile or not. Exploration continued on these two variables
+of interest by plotting each using a stacked bar graph via ggplot().
 
-The same workflow for the Feeder Data was utilized, however, this dataset
-required less wrangling and did not provide as many exploratory options
-as the main focus was on feeder type, date, and the presence/absence of
-*Salmonella* for each day the feeders were sampled.
-
-**Picnic Table Data Exploration and Visualization**
-
-![Positive Salmonella Samples by Type and
-Date](images/Picnic_Table_PosSamp-04.png) Positive *Salmonella* Samples
-by Sample Type and Date
-
-![Positive *Salmonella* Samples by Quadrant and
-Date](images/Picnic_Table_Quad-01.png)
-
-**Bird Feeder Data Exploration and Visualization**
-
-![Positive *Salmonella* Samples by feeder type and
-date.](images/Picnic_Table_PosSamp.png)
+The same workflow for the Feeder Data was utilized, however, this
+dataset required less wrangling and did not provide as many exploratory
+options as the main focus was on feeder type, date, and the
+presence/absence of *Salmonella* for each day the feeders were sampled.
 
@@
 
+**Picnic Table Data Exploration and Visualization**
+
+![Positive Salmonella Samples by Type and
+Date](images/Picnic_Table_PosSamp-05.png) Positive *Salmonella* Samples
+by Sample Type and Date
+
+![Positive *Salmonella* Samples by Quadrant and
+Date](images/Picnic_Table_Quad-02.png)
+
+**Bird Feeder Data Exploration and Visualization**
+
+![Positive *Salmonella* Samples by feeder type and
+date.](images/Feeder_PA.png)
+
 ## **Ibis Trial**
@@
 ### **Persistence**
-On Day 1 of the study, four fecal piles per table (one per quadrant, n = 12) were designated for tracking *Salmonella* persistence.
-**Salmonella** was isolated from three of the 12 designated persistence study piles on Day 1, allowing the persistence to be tracked on these three piles.
-Persistence, expressed as the number of days that a positive fecal pile on Day 1 remained positive, ranged from 1 to 4 days with an average persistence of 3 days.
-The probability of an already positive sample to remain positive the next day declined consistently throughout the study period.
+
+On Day 1 of the study, four fecal piles per table (one per quadrant, n =
+12) were designated for tracking *Salmonella* persistence.
+**Salmonella** was isolated from three of the 12 designated persistence
+study piles on Day 1, allowing the persistence to be tracked on these
+three piles. Persistence, expressed as the number of days that a
+positive fecal pile on Day 1 remained positive, ranged from 1 to 4 days
+with an average persistence of 3 days. The probability of an already
+positive sample to remain positive the next day declined consistently
+throughout the study period.
 
 ### **Prevalence**
-Of 96 deposited feces, we obtained 27% positive.
-*Salmonella* persisted for an average of 3 days, and maximum of 4 days on wooden picnic tables.
-*Salmonella* was isolated from all tables, at least once.
-While there appears to be a significant positive correlation between the number of new fecal piles and the probability of yielding a positive pooled sample.
-The probabilities of a positive sample varied throughout the week, with day 4 having the lowest probability of positive *Salmonella* isolation and day 7 having the most error.
+
+Of 96 deposited feces, we obtained 27% positive. *Salmonella* persisted
+for an average of 3 days, and maximum of 4 days on wooden picnic tables.
+*Salmonella* was isolated from all tables, at least once. While there
+appears to be a significant positive correlation between the number of
+new fecal piles and the probability of yielding a positive pooled
+sample. The probabilities of a positive sample varied throughout the
+week, with day 4 having the lowest probability of positive *Salmonella*
+isolation and day 7 having the most error.
 
 ## **Feeder Trial**
+
 We confirmed successful inoculation of all feeders with *Salmonella*
@@
 tables) with 4 (picnic tables) and 5 (feeders) being the maximum number
-of days persisted. The varied persistence during both trials could be
-a result of a lack of moisture, the fluctuation in ambient temperature
+of days persisted. The varied persistence during both trials could be a
+result of a lack of moisture, the fluctuation in ambient temperature
 during sampling, and water activity. Our studies suggest that

### 2023-04-26 5. Manuscript/Manuscript.qmd
Working update

@@
 
+Statistical analysis was conducted on both datasets, Picnic Table Data
+and Feeder Data, utilizing RStudio "Cherry Blossom" (2023-03-09).
+Analysis for both datasets were evaluated utilizing a linear regression
+model.
+
 # **Results**
@@
 ![Positive Salmonella Samples by Type and
-Date](images/Picnic_Table_PosSamp-05.png) Positive *Salmonella* Samples
-by Sample Type and Date
+Date](images/Picnic_Table_PosSamp-05.png)***Figure 1.*** Positive
+*Salmonella* Samples by Sample Type and Date
 
-![Positive *Salmonella* Samples by Quadrant and
+![***Figure 2.*** Positive *Salmonella* Samples by Quadrant and
 Date](images/Picnic_Table_Quad-02.png)
@@
 
-![Positive *Salmonella* Samples by feeder type and
+![***Figure 3.*** Positive *Salmonella* Samples by feeder type and
 date.](images/Feeder_PA.png)

### 2023-04-28 Project_Review_DeannaL.qmd
Deanna L Project Review

---
title: Project Review Template 
author: Deanna Lanier
date: "`r file.mtime(knitr::current_input())`"
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: Salmonella Enzironmental Persistence Informs Management Relevant to Avian and Public Health

Name of project author(s): Kimberly Perez

Name of project reviewer: Deanna Lanier

# Instructions

# Specific project content evaluation

## Background, Context and Motivation
You provided a very comprehensive background and included great sources. The previous and related work was included, as well as the gap in the current knowledge which was very helpful. This was well written and very informative. It is clear why this project was undertaken. This section does not explicitly state the new information you hope to provide, however it is implied and I was still able to determine the purpose of the project.

### Feedback and Comments

It would be helpful to add 1 or 2 sentences to your introduction about what you will do for this study and what new information you are providing by completing this analysis, or the goal of the study.

### Summary assessment (PICK ONE, DELETE THE OTHERS)
* strong contextualization and motivation

## Question description

The questions were stated and there relation to the data were clear. The questions tie in your description of the data and background very well and it is easy to make the connection.

### Summary assessment
* question/hypotheses fully clear

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

The data is described very clearly and the source of the data and collection was explained. Although the data source is explained and a description of the data is provided, there is no codebood in the repository. I would recommend adding some codebook just to help understand what data are included in these spreadsheets.

### Summary assessment
* source and overall structure of data somewhat explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

The data were processed fully and was explained in the manuscript well. I would have liked a little explanation about the rational behind certain processing steps in the manuscript, however the data processing code was well documented. All of the steps were reasonable and well explained. I do think there could have been more added to the exploratory visualizations. For figure 1, there is no reference to this figure in your manuscript. I believe you describe it in section 5.1.1, however you did not reference that you are describing that figure. For the figures, I would recommend larger text on the x and y axis just to make it easier to visualize. Also, All of the figures are exploratory/descriptive analysis so you should move that label (5.3 Exploratory/Descriptive analysis) to above the text corresponding to those figures.

### Summary assessment
* some weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

For the statistical analysis, I would recommend adding all packages at the beginning of the script to avoid having the install packages midway through. I also was not able to run the script all the way through to the end due to errors that should be fixed. The errors started at line 362 in the statistical analysis code. It looks like you are off to a good start with the analysis, however there is no description or explaination in the manuscript of code so I do not have alot of feedback for this section.

### Summary assessment
* defensible but not optimal analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

There are great figures for the exploratory analysis in the manuscript, but none included of the statistical analysis. I mentioned above that I would increase the axis text size and also include the figure number in the text that describes the findings.

### Summary assessment
* results are presented ok, with room for improvement

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

Sections 5.5, 6.1, 6.2, and 6.3 are not included in the analysis so I cannot provide adequate feedback.

### Summary assessment
* major parts of discussion missing or wrong

## Further comments

The project is off to a great start and the topic is very interesting, but I think there are still some major gaps in the analysis.

# Overall project content evaluation
Evaluate overall features of the project  by filling in the sections below.

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

The project is well structured and the README file explains the folder contents. There were no junk folders that were not needed for analysis. The folders were numbered which was very helpful.

### Summary assessment
* well structured

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

Project is well documented. The only file that I believe could benefit from additional comments is your statistical analysis script.

### Summary assessment
* fully and well documented

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

The statistical analysis is not fully reproducible as there are errors that prevent me from completing the analysis. All other aspects of the project are reproducible.

### Summary assessment
* small parts not reproducible or required manual intervention

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

There background and explanation of the project is extremely thorough, but there are gaps in the statistical analysis. There also was no additional information provided to address the questions you posed in the beginning of the analysis.

### Summary assessment
* decent level of thoroughness

### 2023-04-28 Reviews/Part 5 Review - Luisi.qmd

---
title: Part5 Project Review 
author: NICOLE LUISI
date: "`r file.mtime(knitr::current_input())`"
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: Salmonella Environmental Persistence Informs Management Relevant to Avian and Public Health

Name of project author(s): Kimberly Perez

Name of project reviewer: Nicole Luisi

# Specific project content evaluation

## Background, Context and Motivation

### Feedback and Comments

The background is well done; this topic is outside of my wheelhouse, but I did not have any trouble following the explanation provided. Good description of previous work with references to explain the rationale.

### Summary assessment

* strong contextualization and motivation

## Question description

### Feedback and Comments

Given the background provided, the proposed questions are very clear. I'm not sure why these particular data sources were selected (I assume convenience/access) but these seem like very reasonable data sources to address these hypotheses.

### Summary assessment

* question/hypotheses fully clear

## Data description

### Feedback and Comments

The data description is a sufficient overview, though it seems like there may still be some information still pending for the Picnic Table Trial data. If this information is not available, would recommend rewording this section a bit. I would reorganize this section so that everything about each source is together - after the opening line write all about the Florida data, then all about the Georgia data; it's a little difficult to follow when it goes from FL to GA to FL. Looking ahead to the methods, it seems like you may have actually put more info together about the data collection. There is some repetition throughout the manuscript document as far as the data sources; I think maybe you can combine and simplify these sections....within methods you can do the data source description, etc. I don't see a codebook; the FD data is pretty straight forward, some of the variables in the PTD set aren't as obvious. Could describe specifics about the variables and how they were coded in this section or within comments in the code as they are being processed.

### Summary assessment

* source and overall structure of data somewhat explained

## Data wrangling and exploratory analysis

### Feedback and Comments

Data cleaning file was simple and able to be run; comments explained the purpose for each step. Exploratory analysis file also clean ad easy to run. I think this one could use some more comments explaining the steps (or the thought process after viewing the output).

### Summary assessment

* essentially no weaknesses in wrangling and exploratory component

## Appropriateness of Analysis

### Feedback and Comments

I had some trouble following this section.  I think the progression in the statistical analysis code good and I was able to run those steps, but it seems like maybe you didn't have time to finish writing this up in the manuscript file. There are some notes about leaning towards GLM or including other data. The code looks good (and like you did more than you wrote up) but there aren't many comments, so I'm not sure I have a good handle on the thought process. I'm giving this a good rating because I think you have a lot of good work here, but the manuscript sections related to analysis are missing a lot of info.

### Summary assessment

* strong and reasonable analysis

## Presentation

### Feedback and Comments

The figures are nice, I would just recommend organizing more like a manuscript where you explain the results, then present the table or figure, etc. rather than having everything at the beginning before the text. None of the statistical analysis results seem to be included in tables, etc. or represented in the write-up. There are also some sections with headers and no content (full analysis, exploratory/descriptive analysis. Rating on this one is mostly just due to the missing info in the manuscript, I think it will be good once you get everything in there.

### Summary assessment

* results are poorly presented, hard to understand, poor quality

## Discussion/Conclusions

### Feedback and Comments

Unfortunately this section is missing from the manuscript so I'm not able to evaluate it.

### Summary assessment

* major parts of discussion missing or wrong

## Further comments

For summary/abstract at beginning, would recommend formatting more like a manuscript abstract, so adding in some of the primary findings/discussion points. Overall, the early sections are good shape, it looks like you may have run out of time on the results/discussion. Given that this was likely a time issue I don't much specific advice other than cleaning up these sections and pulling everything together.

# Overall project content evaluation

## Structure

### Feedback and Comments

Everything is mostly well structured, but there are a lot of extra things in the repo. It's clear that users should follow the numbered folder structure but it is a little confusing to have folders "4. Results" and "results", similar with "code" and the code within the numbered folders, "products" and "5. Manuscript". Some of the text within the QMD files also indicates that there were plans to move some things around and maybe that didn't get cleaned up, I would recommend reading through the QMD files again at the end to make sure placeholder language is removed.

### Summary assessment

* mostly clear, but some confusing parts (e.g. useless files, things in the wrong folders)

## Documentation

### Feedback and Comments

Some files could use additional comments, these are noted in the sections above. Overall, things are not difficult to follow. I think some of the README files throughout are a little out of date and need to be updated (may have been ones you made early on and then changed your files and structure).

### Summary assessment

* decently documented with some gaps

## Reproducibility

### Feedback and Comments

I was able to run everything provided without any issues. Some documentation updates would be helpful, these are noted above.

### Summary assessment

* fully reproducible without issues

## Thoroughness

### Feedback and Comments

Overall, I think this is well thought out, again it just seems like there was an issue with time. I think that all the pieces are mostly there, but more work is needed to summarize the findings and conclusion and pull everything together. Without some of this documentation it is difficult to get a sense of the whole thought process and whether or not the questions will be answered sufficiently.

### Summary assessment

* decent level of thoroughness

## Further comments

Early sections are strong, mostly need to spend more time on results and discussion. Overall, would recommend going through the repo to remove unnecessary folders/files and clean up old placeholder text. The manuscript document also has some overlapping sections that could be simplified. I don't have a lot of specific advice because everything you did is good, I think you just didn't have time to finish pulling it all together and double checking/cleaning up the language and files. Hopefully you have some time to make updates for part 6 it will all come together nicely!

### 2023-05-04 5. Manuscript/Manuscript.qmd
Updates for Final

@@
 
-The data selected for my MADA project is persistence and prevalence data
+The data selected for this project is persistence and prevalence data
 from two trial studies. The first trial was conducted in South Florida
@@
 The Picnic Table Trial data is composed of several worksheets embedded
-in one excel file. After viewing the file (and running it below), there
-seems to be 192 observations. I am not sure how the Picnic Table Trial
-was collected but hope to visit with Dr. Hernandez regarding data
-collection methods soon. Feeder Trial data was collected and noted upon
-confirmation of *Salmonella* from processed samples. There are 144
-observations for the Feeder Trial which spanned nearly 2 months. For
-both trials, we were attempting to measure *Salmonella* persistence. The
-Picnic Table Trial collected data in an attempt to measure prevalence as
-well. (Please Note: Aforementioned data was imported under "Data import
-and cleaning" section.)
+in one excel file. After viewing the file, 192 observations were
+returned. Feeder Trial data was collected and noted upon confirmation of
+*Salmonella* from processed samples. The Feeder Trial data returned 144
+observations for the nearly two months of sample collection. Both trials
+were designed to determine the persistence of *Salmonella*. The Picnic
+Table Trial also collected data in an attempt to determine prevalence.
 
@@
 for one week, and weekly for an additional week. Swabs were embedded
-into 10 mL of Dulcitol Selenite broth and transported back to UGA for
+into 10 mL of dulcitol selenite broth and transported back to UGA for
 processing.

### 2023-05-04 5. Manuscript/supplement/Supplementary_Material.qmd
Updates for Final

--- 2023-03-10 products/manuscript/supplement/Supplementary_Material.qmd
@@
 ---
-title: "Supplement to Example Manuscript Template for a Data Analysis Project"
+title: "Supplement to *Salmonella* Environmental Persistence Informs Management Relevant to Avian and Public Health for a Data Analysis Project"
 
 subtitle: ""
-author: Andreas Handel
+author: Kimberly Perez
 date: "`r Sys.Date()`"

### 2023-05-05 3. Analysis Script/StatisticalAnalysis.qmd
More updates

@@
 # **Analysis script**
-This script loads the cleaned data, does a simple analysis, and saves the results to the "4. Results" folder
+
+This script loads the cleaned data, does a series of analysis and saves the results to the "4. Results" folder.
 
 ## **Loading Packages**
+
 ```{r}
@@
 library(tidyverse)
-library(ggplot2) #for plotting
-library(broom) #for cleaning up output from lm()
-library(dotwhisker)
-library(here) #for data loading/saving
-library(tidymodels)
-library(tidyr)
-library(gtsummary)
+library(ggplot2) #For plotting/data vis.
+library(broom) #For cleaning up output from lm()
+library(dotwhisker) #Visualization for regression outputs
+library(here) #Setting paths/for data loading/saving
+library(tidymodels) #Great package for models and other resources used for fitting
+library(tidyr) #Helps with data wrangling
+library(gtsummary) #To create tables 
 library(dplyr)
 library(vip)
-library(performance)
+library(performance) #Model evaluation
 library(jtools)
 library(broom)
-library(broom.mixed)
+library(broom.mixed) #Converts fitted to tiddy dfs
 ```
 
-## **1. Loading: Picnic Table & Feeder Data**
+After reviewing both the Picnic Table and Feeder Trial datasets, I will run several analysis including linear regression, generalized linear model, time to event, and cox regression.
+Given that both datasets are exploring presence/absence of *Salmonella*, I believe the above models will provide the most flexibility with the continuous and categorical predictors contained in the data.
+
+# **1. Loading: Picnic Table & Feeder Data**
+
+Now that the packages we need are loaded, let's read in and view our cleaned data
+
 ```{r}
@@
 ```
-# 2. **Data Fitting/Statistical Analysis: Picnic Table**
+
+# 2. **Data Fitting/Statistical Analysis: Linear Regression \[Picnic Table Trial\]**
+
 Let's Begin with some Linear Model Fitting; I will try this multiple ways, with several predictors.
+
+To begin fitting, we will first need to define the linear regression.
+
 ```{r}
 lmog<-linear_reg() %>% 
-  set_engine("lm")
+  set_engine("lm") #Setting
 
 model_lm<- linear_reg()
+```
 
-lmfit_PT<- model_lm %>% #Training; outcome: Salmonella Pos and Main Predictor Table
+Now that we have defined our linear regression (LR), we will have to train the model to our given data.
+
+```{r}
+lmfit_PT<- model_lm %>% #Training outcome: Salmonella Positive and Main Predictor Date
   fit(Salmonella_Positive~Date, PT_Data)
 
-lmfit_PT
+lmfit_PT #With this function we can view what we just did
 
-glimpse(lmfit_PT)
-
-#Way 1
-glance(lmfit_PT)
-
-#Way 2
-tidy(lmfit_PT)
+tidy(lmfit_PT) #We can use the tidy() function to produce the output/ summary statistics
 
@@
 
+#Plotting the above 
 tidy(lmfit_PT) %>%
@@
 
+Again, we will need to define the linear regression for fitting of other specific indicators.
+
 ```{r}
@@
 lmog_1<-linear_reg() %>% 
-  set_engine("lm")
+  set_engine("lm") #Setting
 
 model_lm1<- linear_reg()
+```
 
+We have defined our LR, now we will train the model to our given data and new predictor.
+
+```{r}
 lmfit_PT1<- model_lm1 %>% #Training; outcome: Salmonella Pos
@@
 
-lmfit_PT1
+lmfit_PT1 #Let's view what we just did
 
-tidy(lmfit_PT1)
-
-glimpse(lmfit_PT1)
-
-#Way 1
-glance(lmfit_PT1)
-
-#Way 2
-tidy(lmfit_PT1)
+tidy(lmfit_PT1) #We can use the tidy() function to produce the output/ summary statistics
 
@@
 
+We will follow the same workflow as above, however, now we will utilize all predictors of the Picnic Table Trial Data.
+
 ```{r}
@@
 
-lmall_PT<- lmog_PT %>% #Training 
+model_lmall<- linear_reg()
+
+lmall_PT<- model_lmall %>% #Training 
   fit(Salmonella_Positive~., data=PT_Data)
@@
 tidy(lmall_PT)
+```
 
-#Do not want all predictors...
+I do not believe all predictors will be beneficial in this analysis. However, we will continue to check the model performance and include it in the supplemental section.
 
+```{r}
 #Checking model performance: Picnic Table
@@
 ```
+
+The code below \[lines132-154\] will likely be included in the supplemental section as it should produce similar values as the above code.
+I wanted to include this as an alternate way of writing the code and most importantly a resource to refer back to in the future.
 
@@
 
-#A different way Part Deux 
+#A different way 
 # I will do the same as above, however, I am now adding persistence pile
@@
 
-Second model fit
+# 2. **Data Fitting/Statistical Analysis: Linear Regression \[Feeder Trial\]**
+
+This section is a continuation of what was produced above, however, we will now be utilizing the Feeder Trial Data
+
 ```{r}
@@
 
-lmfit_FD<- lmog_FD %>% #Training; outcome: Salmonella Pos and Main Predictor Table
+model_lmFD<- linear_reg()
+
+lmfit_FD<- model_lmFD %>% #Training; outcome: Salmonella Pos and Main Predictor Table
   fit(Absence_0_Presence_1~Feeder_number, FD_Data)
 
-lmfit_FD
+lmfit_FD #Let's view what we just did
 
@@
 
-#Way 1
-glance(lmfit_FD)
-
-#Way 2
-tidy(lmfit_FD)
+tidy(lmfit_FD) #We can use the tidy() function to produce the output/ summary statistics
 
@@
 check_model(lmfit_FD$fit)
+```
 
+```{r}
 #More fitting with some other indicators
-
 lmog_FD2<-linear_reg() %>% 
@@
 
-lmfit_FD2<- lmog_FD2 %>% #Training; outcome: Salmonella Pos and Main Predictor Table
+lmfit_FD2<- lmog_FD2 %>% #Training; outcome: Salmonella Positive and Main Predictor Feeder Number and Type (e.g., wood v. plastic)
   fit(Absence_0_Presence_1~Feeder_number+Feeder_Type, FD_Data)
 
-lmfit_FD2
+lmfit_FD2 #Let's view what we just did
 
@@
 
-#Way 1
-glance(lmfit_FD2)
-
-#Way 2
-tidy(lmfit_FD2)
+tidy(lmfit_FD2) #We can use the tidy() function to produce the output/ summary statistics
 
@@
 
+check_model(lmfit_FD2$fit)
+```
+
+```{r}
 #Trying ALL predictors of Feeder dataset 
@@
 
-tidy(lmall_FD)
+tidy(lmall_FD) #We can use the tidy() function to produce the output/ summary statistics
 
-plot_summs(lmall_FD, robust=TRUE) #KP NOTE: I need to clean this up so it is a bit more readable, must fullscreen it
+plot_summs(lmall_FD, robust=TRUE)
+
+check_model(lmall_FD$fit)
 ```
-# **Generalized Linear Model: Picnic Table** 
+
+# 3. **Data Fitting/Statistical Analysis: Generalized Linear Model \[Picnic Table Trial\]**
+
+Unlike linear regressions, generalized linear models (GLMs) are often more flexible and allow for change in inputs. Below I will run GLMs for both datasets and check each model.
+
 ```{r}
@@
 
-#Way 1
-glance(lmfit_PT_glm)
-
-#Way 2
 tidy(lmfit_PT_glm)
@@
 plot_summs(lmfit_PT_glm)
+
+check_model(lmfit_PT_glm$fit)
 
@@
 
-lmfit_PTglm1<- lmog_glm1 %>% #Training; outcome: Salmonella Pos and Main Predictor Table
-  fit(Salmonella_Positive~Original_Persistence_Pile+Table, PT_Data)
+lmfit_PTglm1<- lmog_glm1 %>% #Training; outcome: Salmonella Positive and MPs: Pile, Table, and Date
+  fit(Salmonella_Positive~Original_Persistence_Pile+Table+Date, PT_Data)
 
@@
 
-glimpse(lmfit_PTglm1)
-
-#Way 1
-glance(lmfit_PTglm1)
-
-#Way 2
 tidy(lmfit_PTglm1)
@@
 plot_summs(lmfit_PTglm1)
+
+check_model(lmfit_PTglm1$fit)
 
@@
   fit(Salmonella_Positive~Quadrant+Table, PT_Data)
+
+check_model(lmfit_PTglm2$fit)
 ```
@@
 
-#Way 1
-glance(lmfit_PTglm2)
-
-#Way 2
 tidy(lmfit_PTglm2)
@@
 plot_summs(lmfit_PTglm2)
+
+check_model(lmfit_PTglm2$fit)
 
@@
 
-# **Generalized Linear Model: Picnic Table (Persistence & Prevalence)** 
+## a. **Generalized Linear Model: Picnic Table (Persistence & Prevalence)**
+Again, this is an alternative way to run a glm.
+
 ```{r}
-PT_GLM_PER <- glm(Salmonella_Positive ~ Date, data = PT_Data, family = binomial)
+PT_GLM_PER <- glm(Salmonella_Positive ~ Date, data = PT_Data, family = binomial) #Persistence
 
@@
 
-PT_GLM_PREV <- glm(Salmonella_Positive ~ Sample_Type + as.factor(Date) + Table, data= PT_Data, family = binomial)
+tidy(PT_GLM_PER)
+
+PT_GLM_PREV <- glm(Salmonella_Positive ~ Sample_Type + as.factor(Date) + Table, data= PT_Data, family = binomial) #Prevalence
 
 PT_GLM_PREV
+
+tidy(PT_GLM_PREV)
+
+lmall_PTGLM
 ```
 
-# **Generalized Linear Model: Feeder Data**
+# 4. **Data Fitting/Statistical Analysis: Generalized Linear Model [Feeder Trial]**
+
 ```{r}
@@
 
-#Way 1
-glance(lmfit_FD_glm)
-
 #Way 2
@@
 plot_summs(lmfit_FD_glm)
+
+check_model(lmfit_FD_glm$fit)
 
@@
 
-#Way 1
-glance(lmfit_FD_GLM)
-
-#Way 2
 tidy(lmfit_FD_GLM)
@@
 plot_summs(lmfit_FD_GLM)
+
+check_model(lmfit_FD_GLM$fit)
 
@@
 tidy(lmall_FDGLM_all1)
+
+check_model(lmall_FDGLM_all1$fit)
 ```
 
-# **Generalized Linear Model: Feeder Data (Persistence & Prevalence)**
+# **Generalized Linear Model: Feeder Data (Persistence)**
+
 ```{r}
-FD_GLM_PER <- glm(Absence_0_Presence_1 ~ Collection_Date, data = FD_Data, family = binomial)
+FD_GLM_PER <- glm(Absence_0_Presence_1 ~ Collection_Date, data = FD_Data, family = binomial) #Persistence
 
@@
 
-FD_GLM_PREV <- glm(Absence_0_Presence_1 ~ Feeder_Type + as.factor(Collection_Date), data= FD_Data, family = binomial)
+FD_GLM_PREV <- glm(Absence_0_Presence_1 ~ Feeder_Type + as.factor(Collection_Date), data= FD_Data, family = binomial) #Prevalence
 
@@
 ```
+
 # **Time to event**
+
 ```{r}
@@
 
-#**Cox regression: Picnic Table Data**
+# **Cox regression: Picnic Table Data**
+
 ```{r}
@@
 ```
-#**Cox regression: Feeder Data**
+
+\#**Cox regression: Feeder Data**
+
 ```{r}

### 2023-05-05 5. Manuscript/Manuscript.qmd
More updates

@@
 lined with antimicrobial coatings.
+
+## **Data acquisition**
+
+The data from the Picnic Table Trial in South Florida was collected by
+an undergraduate student of Dr. Sonia M. Hernandez and provided to me by
+Dr. Hernandez. All data from the Feeder Trial was collected by me,
+Kimberly Perez under the guidance of Dr. Sonia Hernandez.
 
@@
 
-After collection, 100 uL of the sample vortexed in dulcitol selentite
-was inoculated into 10 mL of Rappaport Vassiliadis (RV) broth. The broth
-was incubated for 24 hours at 37c. The RV broth was then streaked onto
-XLT-4 agar plates and incubated at 37c for 24 hours, then an additional
+After collection, dulcitol selenite was incubated overnight at 37°c,
+before 100 uL of the sample vortexed in dulcitol selentite was
+inoculated into 10 mL of Rappaport Vassiliadis (RV) broth. The broth was
+incubated for 24 hours at 37°c. The RV broth was then streaked onto
+XLT-4 agar plates and incubated at 37°c for 24 hours, then an additional
 24 hours on the lab bench at room temperature. If colonies
@@
 something other than *Salmonella* [@Maurer2015].
-
-## **Data acquisition**
-
-The data from the Picnic Table Trial in South Florida was collected by
-an undergraduate student of Dr. Sonia M. Hernandez and provided to me by
-Dr. Hernandez. All data from the Feeder Trial was collected by me,
-Kimberly Perez under the guidance of Dr. Sonia Hernandez.
 
@@
 **Picnic Table Data Exploration and Visualization**
+```{r, echo=FALSE, message= FALSE, fig.width=10, fig.height=8, fig.cap= "Figure. 1 Positive Salmonella Samples by Sample Type and Date"}
 
-![Positive Salmonella Samples by Type and
-Date](images/Picnic_Table_PosSamp-05.png)***Figure 1.*** Positive
-*Salmonella* Samples by Sample Type and Date
+library(here)
 
-![***Figure 2.*** Positive *Salmonella* Samples by Quadrant and
-Date](images/Picnic_Table_Quad-02.png)
+knitr::include_graphics(here("4. Results/Picnic_Table_PosSamp.png"))
+```
+
+```{r, echo=FALSE, message= FALSE, fig.width=10, fig.height=8, fig.cap= "Figure. 2 Positive Salmonella Samples by Quadrant and Date"}
+
+library(here)
+
+knitr::include_graphics(here("4. Results/Picnic_Table_Quad.png"))
+```
 
 **Bird Feeder Data Exploration and Visualization**
+```{r, echo=FALSE, message= FALSE, fig.width=10, fig.height=8, fig.cap= "Figure. 3 Positive Salmonella Samples by Feeder Type and Date"}
 
-![***Figure 3.*** Positive *Salmonella* Samples by feeder type and
-date.](images/Feeder_PA.png)
+library(here)
 
-## **Ibis Trial**
+knitr::include_graphics(here("4. Results/Feeder_PA.png"))
+```
+
+## **Picnic Table Trial**
 

### 2023-05-05 3. Analysis Script/StatisticalAnalysis.qmd
Final Pushes of MADA

@@
 library(broom.mixed) #Converts fitted to tiddy dfs
+library(finalfit) #Needed for Time to Event Analysis 
+library(forcats) #Needed for Time to Event Analysis 
+library(survival) #Needed for Time to Event Analysis 
+library(lubridate)
+library(ggsurvfit)
+library(gt)
+library(webshot) #To save tables
+library(knitr) #For tables
 ```
@@
 
-# 2. **Data Fitting/Statistical Analysis: Linear Regression \[Feeder Trial\]**
+# 2a. **Data Fitting/Statistical Analysis: Linear Regression \[Feeder Trial\]**
 
@@
 
-# 3. **Data Fitting/Statistical Analysis: Generalized Linear Model \[Picnic Table Trial\]**
+# 2b. **Data Fitting/Statistical Analysis: Generalized Linear Model \[Picnic Table Trial\]**
 
@@
 ```
+## 3a. **Generalized Linear Model: Picnic Table (Persistence & Prevalence)**
 
-## a. **Generalized Linear Model: Picnic Table (Persistence & Prevalence)**
 Again, this is an alternative way to run a glm.
@@
 
-# 4. **Data Fitting/Statistical Analysis: Generalized Linear Model [Feeder Trial]**
+# 3b. **Data Fitting/Statistical Analysis: Generalized Linear Model \[Feeder Trial\]**
 
@@
 ```
-
-# **Generalized Linear Model: Feeder Data (Persistence)**
+## 3b. **Generalized Linear Model: Feeder Data (Persistence)**
 
@@
 ```
+# 4a. **Time to Event Analysis: Picnic Table Trial**
+```{r}
 
-# **Time to event**
+PT_Data_TTE<- #First let's format the dates 
+  PT_Data %>% 
+  mutate(
+    Date= ymd(Date))
+
+PT_Data_TTE$Date<- as.numeric(PT_Data_TTE$Date) #Let's make these numeric
+
+PT_Data_TTE <- #I want to add a column displaying time between events (by day)
+  PT_Data_TTE %>%
+  mutate(
+    os_day= as.duration(Date)/ddays(1))
+
+PT_Data_TTE$os_day<- as.numeric(PT_Data_TTE$os_day) #Again let's make sure our new column is in the proper format
+
+surv1<- Surv(PT_Data_TTE$Date,PT_Data_TTE$Salmonella_Positive)
+
+#Let's visualize!
+sfit1<- survfit(Surv(Date, Salmonella_Positive) ~  1, data=PT_Data_TTE)  %>% 
+  ggsurvfit() + 
+  labs(
+    x = "Days",
+    y = "Overall Salmonella Persistence") + #Let's add some labels
+  add_confidence_interval() +
+  add_risktable()
+
+#I am not confident with this graphic so I will include this in my supplemental section
+```
 
 ```{r}
-library(finalfit)
-library(forcats)
-library(survival)
-
 glimpse(PT_Data)
-
-PT_Data$Date<- as.numeric(PT_Data$Date)
 
@@
   salmonella_crr= case_when(
-        Salmonella_Positive ==  0~1
-        Salmonella_Positive == 1~1
+   Salmonella_Positive ==  0~1,
+    Salmonella_Positive == 1~1,
      TRUE ~ 1))
 
-dependent_crr= "Surv(Date, status_crr)"
+dependent_crr= "Surv(Date, salmonella_crr)"
 explanatory= c("Table", "Quadrant", "Sample_Type")
+
+PT_Data$Date<- as.numeric(PT_Data$Date)
 
@@
 
-PT_Data %>%
-  filter(status==1) %>%
+PT_MedSurv<-PT_Data %>%
+ filter(salmonella_crr==1) %>%
   summarize(median_surv=median(Date))
+
+PT_MedSurv <- ((PT_MedSurv)/31556952000)*100
+
+PT_Surv_Table<- gt(PT_MedSurv) %>%
+   tab_header(
+    title = "Median Survival of Salmonella",
+    subtitle = ("01/02/2020 to 01/09/2020")) %>%
+  cols_label(
+    median_surv= "Median Salmonella Survival (days)")
+
+PT_Surv_Table
+
+PT_Surv_Table %>%
+  saveRDS(here("4. Results", "Picnic_Table_Median_Survival.rds")) #To include in my manuscript, persistence seems to be roughly 5 days
+```
+# 4b. **Time to Event Analysis: Feeder Trial**
+```{r}
+FD_Data_TTE<- #First let's format the dates 
+  FD_Data %>% 
+  mutate(
+    Date= ymd(Collection_Date))
+
+FD_Data_TTE$Collection_Date<- as.numeric(FD_Data_TTE$Collection_Date) #Let's make these numeric
+
+FD_Data_TTE <- #I want to add a column displaying time between events (by day)
+  FD_Data_TTE %>%
+  mutate(
+    os_day= as.duration(Collection_Date)/ddays(1))
+
+FD_Data_TTE$os_day<- as.numeric(FD_Data_TTE$os_day) #Again let's make sure our new column is in the proper format
+
+surv1_FD<- Surv(FD_Data_TTE$Collection_Date,FD_Data_TTE$Absence_0_Presence_1)
+
+#Let's visualize!
+sfit1<- survfit(Surv(Collection_Date, Absence_0_Presence_1) ~  1, data=FD_Data_TTE)  %>% 
+  ggsurvfit() + 
+  labs(
+    x = "Days",
+    y = "Overall Salmonella Persistence") + #Let's add some labels
+  add_confidence_interval() +
+  add_risktable()
+
+sfit1
+```
+```{r}
+FD_Data = FD_Data %>% 
+ mutate(
+  salmonella_crr= case_when(
+   Absence_0_Presence_1 ==  0~1,
+    Absence_0_Presence_1 == 1~1,
+     TRUE ~ 1))
+
+dependent_crr= "Surv(Collection_Date, salmonella_crr)"
+explanatory= c("Feeder_Type")
+
+FD_Data$Collection_Date<- as.numeric(FD_Data$Collection_Date)
+
+FD_Data %>% finalfit(dependent_crr, explanatory)
+
+FD_Data %>% 
+   finalfit(dependent_crr, explanatory, add_dependent_label = FALSE) %>% 
+    rename("Overall survival" = label) %>% 
+    rename(" " = levels) %>% 
+    rename("  " = all)
+
+survfit(Surv(Collection_Date, Absence_0_Presence_1)~1, data=FD_Data)
+
+FD_MedSurv<-FD_Data %>%
+ filter(salmonella_crr==1) %>%
+  summarize(median_surv=median(Collection_Date))
+
+FD_MedSurv <- ((FD_MedSurv)/604800)*100 #Persistence seems to be roughly 3 days for the FD
+
+FD_Surv_Table<- gt(FD_MedSurv) %>%
+   tab_header(
+    title = "Median Survival of Salmonella",
+    subtitle = ("02/22/2022 to 03/14/2022")) %>%
+  cols_label(
+    median_surv= "Median Salmonella Survival (days)")
+
+FD_Surv_Table
+
+FD_Surv_Table %>%
+  saveRDS(here("4. Results", "Feeder_Median_Survival.rds")) #To include in my manuscript, seems to be roughly 3 days
 ```
 
-# **Cox regression: Picnic Table Data**
+# 5a. **Cox regression: Picnic Table Trial**
 
@@
 
-coxph(Surv(Date, Salmonella_Positive) ~ Sample_Type, PT_Data) %>%
+CR_PT1<-coxph(Surv(Date, Salmonella_Positive) ~ Sample_Type, PT_Data) %>%
   tbl_regression(exp=TRUE)
+
+CR_PT1
+
+CR_PT1%>%
+  saveRDS(here("4. Results", "PT_Cox.rds"))
+
 ```
 
-\#**Cox regression: Feeder Data**
+# 5b. **Cox regression: Feeder Trial**
 
 ```{r}
-library(lubridate)
-
 FD_Data$Collection_Date<- as.numeric(FD_Data$Collection_Date)
@@
 
-coxph(Surv(Collection_Date, Absence_0_Presence_1) ~ Feeder_Type, FD_Data) %>%
+CR_FD1<- coxph(Surv(Collection_Date, Absence_0_Presence_1) ~ Feeder_Type, FD_Data) %>%
   tbl_regression(exp=TRUE)
+
+CR_FD1
+
+CR_FD1%>%
+  saveRDS(here("4. Results", "FD_Cox.rds"))
 ```
+It looks like the

### 2023-05-05 5. Manuscript/Manuscript.qmd
Final Pushes of MADA

@@
 
-## **Description of data and data source**
+### **Description of data and data source**
 
@@
 More packages were loaded that would assist in the visualization process
-before begining this process. Before visualization began, the glimpse()
+before beginning this process. Before visualization began, the glimpse()
 function was utilized to view the recently cleaned data. Given the
@@
 
-Statistical analysis was conducted on both datasets, Picnic Table Data
-and Feeder Data, utilizing RStudio "Cherry Blossom" (2023-03-09).
-Analysis for both datasets were evaluated utilizing a linear regression
-model.
+Statistical analysis was conducted on the Picnic Table Data and Feeder
+Data, utilizing RStudio "Cherry Blossom" (2023-03-09). Analysis for both
+datasets were evaluated utilizing several models including Linear
+Regression (LR), Generalized Linear Model (GLM), and Cox Regression
+(CR). Given that two datasets were utilized for this analysis, a
+holistic approach, considering various limitations of both trials, such
+as sample size, was taken into account when selecting a model. Because
+the majority of data is binary in nature (e.g., presence v. absences)
+and few variables exist for comparison, a LR was the first model
+selected to run on both datasets, followed by a GLM, and CR. Several
+predictors were selected to include in this the analysis including
+'Salmonella_Positive\~Date' , 'Salmonella_Positive\~Sample_Type' (e.g.,
+pooled v. persistent), and 'Salmonella_Positive\~.' with all predictors
+for the Picnic Table Trial and 'Absence_0\_Presence_1\~Feeder_number',
+'Absence_0\_Presence_1\~Feeder_number+Feeder_Type', and
+'Absence_0\_Presence_1\~.' and all predictors for the Feeder Trial.
+
+Both Linear Regression and Generalized Linear Models performed as
+expected when model performance was checked utilizing the check_models()
+function. Given the flexibility and model performance of the GLM, it was
+selected as the model for inclusion in our analysis between predictors
+(e.g., 'Date' \[PT\], 'Sample_Type' \[PD\], 'Feeder_Type' \[FD\]) and
+outcomes (e.g., 'Salmonella_Positive' \[PT\], 'Absence_0\_Presence_1'
+\[FD\]). To
+
+While LR and GLM provide insight into
 
@@
 **Picnic Table Data Exploration and Visualization**
+
 ```{r, echo=FALSE, message= FALSE, fig.width=10, fig.height=8, fig.cap= "Figure. 1 Positive Salmonella Samples by Sample Type and Date"}
@@
 **Bird Feeder Data Exploration and Visualization**
+
 ```{r, echo=FALSE, message= FALSE, fig.width=10, fig.height=8, fig.cap= "Figure. 3 Positive Salmonella Samples by Feeder Type and Date"}

### 2023-05-06 3. Analysis Script/StatisticalAnalysis.qmd
Good Night, and Good Luck

@@
 library(gt)
-library(webshot) #To save tables
 library(knitr) #For tables
@@
 ```
-It looks like the
+With our models run, let's see which one did the best. Because we have a small sample size, we will run an AICc.
+```{r}
+
+AIC_TAB_picnic<- compare_performance(lmfit_PT,lmfit_PT1,lmall_PT, lmfit_PT_glm, lmfit_PTglm1, lmfit_PTglm2, lmall_PTGLM, rank= TRUE, metrics="AICc")
+
+print(AIC_TAB_picnic)
+
+AIC_TAB_picnic %>%
+  saveRDS(here("4. Results", "LMModelAICPD.rds"))
+
+AIC_TAB_feeder<- compare_performance(lmfit_FD_glm, lmfit_FD, lmfit_FD2, lmall_FD, lmfit_FD_GLM, lmall_FDGLM_all1, rank= TRUE, metrics="AICc")
+
+print(AIC_TAB_feeder)
+
+AIC_TAB_feeder%>%
+  saveRDS(here("4. Results", "ModelAICFD.rds"))
+```

### 2023-05-06 5. Manuscript/Manuscript.qmd
Good Night, and Good Luck

@@
 
--   What is the prevalence of *Salmonella* over time?
+-   What is the persistence of *Salmonella* over time?
 
@@
 
-# **Statistical analysis**
-
-Statistical analysis was conducted on the Picnic Table Data and Feeder
-Data, utilizing RStudio "Cherry Blossom" (2023-03-09). Analysis for both
-datasets were evaluated utilizing several models including Linear
-Regression (LR), Generalized Linear Model (GLM), and Cox Regression
-(CR). Given that two datasets were utilized for this analysis, a
-holistic approach, considering various limitations of both trials, such
-as sample size, was taken into account when selecting a model. Because
-the majority of data is binary in nature (e.g., presence v. absences)
-and few variables exist for comparison, a LR was the first model
-selected to run on both datasets, followed by a GLM, and CR. Several
-predictors were selected to include in this the analysis including
-'Salmonella_Positive\~Date' , 'Salmonella_Positive\~Sample_Type' (e.g.,
-pooled v. persistent), and 'Salmonella_Positive\~.' with all predictors
-for the Picnic Table Trial and 'Absence_0\_Presence_1\~Feeder_number',
-'Absence_0\_Presence_1\~Feeder_number+Feeder_Type', and
-'Absence_0\_Presence_1\~.' and all predictors for the Feeder Trial.
-
-Both Linear Regression and Generalized Linear Models performed as
-expected when model performance was checked utilizing the check_models()
-function. Given the flexibility and model performance of the GLM, it was
-selected as the model for inclusion in our analysis between predictors
-(e.g., 'Date' \[PT\], 'Sample_Type' \[PD\], 'Feeder_Type' \[FD\]) and
-outcomes (e.g., 'Salmonella_Positive' \[PT\], 'Absence_0\_Presence_1'
-\[FD\]). To
-
-While LR and GLM provide insight into
-
 # **Results**
 
-**Picnic Table Data Exploration and Visualization**
-
-```{r, echo=FALSE, message= FALSE, fig.width=10, fig.height=8, fig.cap= "Figure. 1 Positive Salmonella Samples by Sample Type and Date"}
-
-library(here)
-
-knitr::include_graphics(here("4. Results/Picnic_Table_PosSamp.png"))
-```
-
-```{r, echo=FALSE, message= FALSE, fig.width=10, fig.height=8, fig.cap= "Figure. 2 Positive Salmonella Samples by Quadrant and Date"}
-
-library(here)
-
-knitr::include_graphics(here("4. Results/Picnic_Table_Quad.png"))
-```
-
-**Bird Feeder Data Exploration and Visualization**
-
-```{r, echo=FALSE, message= FALSE, fig.width=10, fig.height=8, fig.cap= "Figure. 3 Positive Salmonella Samples by Feeder Type and Date"}
-
-library(here)
-
-knitr::include_graphics(here("4. Results/Feeder_PA.png"))
-```
-
-## **Picnic Table Trial**
-
-### **Persistence**
-
-On Day 1 of the study, four fecal piles per table (one per quadrant, n =
-12) were designated for tracking *Salmonella* persistence.
-**Salmonella** was isolated from three of the 12 designated persistence
-study piles on Day 1, allowing the persistence to be tracked on these
-three piles. Persistence, expressed as the number of days that a
-positive fecal pile on Day 1 remained positive, ranged from 1 to 4 days
-with an average persistence of 3 days. The probability of an already
-positive sample to remain positive the next day declined consistently
-throughout the study period.
-
-### **Prevalence**
-
-Of 96 deposited feces, we obtained 27% positive. *Salmonella* persisted
-for an average of 3 days, and maximum of 4 days on wooden picnic tables.
-*Salmonella* was isolated from all tables, at least once. While there
-appears to be a significant positive correlation between the number of
-new fecal piles and the probability of yielding a positive pooled
-sample. The probabilities of a positive sample varied throughout the
-week, with day 4 having the lowest probability of positive *Salmonella*
-isolation and day 7 having the most error.
-
-## **Feeder Trial**
-
-We confirmed successful inoculation of all feeders with *Salmonella*
-(presence on day post inoculation \[DPI\] 0). Further, results indicate
-that *Salmonella* may persist on some feeders up to DPI 5 and current
-analysis indicates that *Salmonella* persisted on both coated and
-uncoated plastic feeders at a higher rate than wood and wood coated.
-Figure 2 depicts the percent of samples positive based on feeder type
-throughout the study period. While plastic and plastic coated feeders
-had an overall higher percent positivity, a spike in percent positivity
-in wood feeders occurred on DPI 5 after a rain event. Further analyses
-should be conducted to determine whether environmental conditions (e.g.,
-rain event) had a significant influence on persistence.
+## **Exploratory/Descriptive Analysis**
 
@@
 
-## **Exploratory/Descriptive analysis**
+### **Picnic Table Data: Exploration and Visualization**
 
-## **Basic statistical analysis**
+```{r, echo=FALSE, message= FALSE, fig.width=10, fig.height=8, fig.cap= "Figure. 1 Positive Salmonella Samples by Sample Type and Date"}
 
-After the modules following Part 1, I am leaning towards running a GLM
-for both given the binary responses for *Salmonella* persistence (e.g.,
-presence v. absence). I need to meet with Dr. Hernandez to clear up some
-questions I have on the picnic table data, specifically regarding the
-introduction of new feces. I am curious if I should include other data
-into the mix to make my analysis more robust. At the moment, the
-variables of interest for both are limited (e.g., two for the picnic
-table data, and one primary variable of interest for the feeder trial).
+library(here)
 
-## **Full analysis**
+knitr::include_graphics(here("4. Results/Picnic_Table_PosSamp.png"))
+```
+
+Figure 1 depicts the the number of positive *Sallmonella* samples by
+sample type, persistence or pooled. Pooled samples were collected from
+every quadrant daily, by swabbing a sampling sponge along the picnic
+table surface. Persistence piles, on the other hand were, selected based
+on two criteria, quadrant and size of fecal pile. The persistence pile
+was sampled everyday for 8 days. While the information presented in
+Figure 1 depicts count data, it is a useful visual to convey numbers of
+positive samples for each sample type based on day.
+
+```{r, echo=FALSE, message= FALSE, fig.width=10, fig.height=8, fig.cap= "Figure. 2 Positive Salmonella Samples by Quadrant and Date"}
+
+library(here)
+
+knitr::include_graphics(here("4. Results/Picnic_Table_Quad.png"))
+```
+
+Figure 2 depicts the the number of positive *Sallmonella* samples by
+quadrant. While the information presented in Figure 2 depicts count
+data, it is a useful visual to convey which quadrants may have been "hot
+spots" for the presence of *Salmonella*.
+
+**Persistence** On Day 1 of the study, four fecal piles per table (one
+per quadrant, n = 12) were designated for tracking *Salmonella*
+persistence. **Salmonella** was isolated from three of the 12 designated
+persistence study piles on Day 1, allowing the persistence to be tracked
+on these three piles. Persistence, expressed as the number of days that
+a positive fecal pile on Day 1 remained positive, varied with an average
+persistence of 3 days.
+
+**Prevalence** Of 96 deposited feces, we obtained 27% positive (26/96).
+*Salmonella* persisted for an average of 3 days, and maximum of 4 days
+on wooden picnic tables. *Salmonella* was isolated from all tables, at
+least once.
+
+### **Bird Feeder Data Exploration and Visualization**
+
+```{r, echo=FALSE, message= FALSE, fig.width=10, fig.height=8, fig.cap= "Figure. 3 Positive Salmonella Samples by Feeder Type and Date"}
+
+library(here)
+
+knitr::include_graphics(here("4. Results/Feeder_PA.png"))
+```
+
+Figure 3 depicts the the number of bird feeders that remained
+*Salmonella* positive throughout the trial. Stacked bars in various
+colors indicate the number of feeders (plastic, plastic + antimicrobial
+coating, wood, and wood + antimicrobial coating) from which we isolated
+*Salmonella*. We confirmed successful inoculation of all feeders with
+Salmonella (presence on Day Post Inoculation (DPI) 0- 2/22/2022).
+Further, results indicate that Salmonella may persist on some feeders up
+to five DPI and current analysis indicates that Salmonella persisted on
+both coated and uncoated feeders. With persistence occurring with both
+plastic and plastic coated feeders.
+
+# **Statistical Analysis**
+
+Statistical analysis was conducted on the Picnic Table Data and Feeder
+Data, utilizing RStudio "Cherry Blossom" (2023-03-09). Analysis for both
+datasets were evaluated utilizing several models including Linear
+Regression (LR) and Generalized Linear Model (GLM). Given that two
+datasets were utilized for this analysis, a holistic approach,
+considering various limitations of both trials, such as sample size, was
+taken into account when selecting a model. Because the majority of data
+is binary in nature (e.g., presence v. absence) and few variables exist
+for comparison, a LR was the first model selected to run on both
+datasets, followed by a GLM. Several predictors were selected to include
+in this the analysis including 'Salmonella_Positive\~Date' ,
+'Salmonella_Positive\~Sample_Type' (e.g., pooled v. persistent), and
+'Salmonella_Positive\~.' with all predictors for the Picnic Table Trial
+and 'Absence_0\_Presence_1\~Feeder_number',
+'Absence_0\_Presence_1\~Feeder_number+Feeder_Type', and
+'Absence_0\_Presence_1\~.' and all predictors for the Feeder Trial.
+
+## **Regression Models**
+
+### **Picnic Table Trial**
+
+```{r}
+#| label: Table 1
+#| tbl-cap: "Small sample adjusted Akaike's Information Criteria (AICc) model weights and performance scores for linear and generalized linear models predicting Salmonella Persistance on Picnic Tables.
+"
+#| echo: FALSE
+LMModelAICPD = readRDS(here("4. Results", "LMModelAICPD.rds"))
+knitr::kable(LMModelAICPD)
+```
+
+Of the models evaluated, the highest performing of all contained the
+predictors original persistence pile \[Original_Persistence_Pile\],
+table, and date and the outcome of *Salmonella* \[Salmonella_Positive\].
+Given the AICc_wt along with performance score, a GLM was selected for
+the Picnic Table Trial data, as depicted in Table 1. When evaluating a
+model using AICc, we would ideally like to select a model with a low
+AICc and a high performance score. Out of all of the models run, the
+model labeled "lmfit_PTglm1" was selected. In this model, original
+persistence pile along with table and date was associated with a higher
+likelihood of Salmonella persistence.
+
+### **Feeder Trial**
+
+```{r}
+#| label: Table 2
+#| tbl-cap: "Small sample adjusted Akaike's Information Criteria (AICc) model weights and performance scores for linear and generalized linear models predicting Salmonella Persistance on Various Feeder Surfaces."
+#| echo: FALSE
+ModelAICFD= readRDS(here("4. Results", "ModelAICFD.rds"))
+knitr::kable(ModelAICFD)
+```
+
+Of the models evaluated, the highest performing of all contained all
+predictors. Given the AICc_wt along with performance score, a LM was
+selected for the Feeder Trial data, as depicted in Table 2. When
+evaluating a model using AICc, we would ideally like to select a model
+with a low AICc and a high performance score, as mentioned above.
+However, because this model can be considered "global" as it includes
+all predictors, it is not the most ideal as it may overfit the data.
+Despite this, the model labeled "lmall_FD" was selected. In this model,
+feeder number, collection data, and feeder type was associated with a
+higher likelihood of Salmonella persistence on feeders.
 
 # **Discussion**
+As cases of environmental salmonellosis increase throughout the United
+States, environs that individuals come into contact with on a regular
+basis need to be evaluated to fill knowledge gaps relating to microbial
+transmission [@Hernandez2016]. With these trials we will research the
+factors that influence *Salmonella* persistence, which will allow us to
+better inform public health agencies and mitigate seasonal outbreaks in
+humans and wildlife.
 
 ## **Summary and Interpretation**
+Based on the modeling conducted above, there seems to be positive
+effects of the various predictors on *Salmonella* absence/presence. The
+Picnic Table Data showed that the original pile, date, and table all had
+statistically significant effects on *Salmonella* presence on picnic
+tables in South Florida. Persistence of *Salmonella* varied, however,
+the maximum number of days it persisted was four
+days (12 positive persistence piles/32 piles total).
+
+Feeder data also indicate that *Salmonella* can persist on various surfaces for several days.
+*Salmonella* persisted for an average of ~2 days, (5 days max).
+Surprisingly, however, *Salmonella* persisted on plastic and plastic coated feeders longer than wooden and wooden coated feeders and had a prevalence of 15% on plastic coated feeders.
+
+The results from both trials indicate that *Salmonella* can persist on various surfaces.
+As human and wildlife interactions increase due to urbanization, precautions should be taken to avoid potential pathogen transmission from the wildlife directly or through interactions with anthropogenic environs that may harbor wildlife excrement.
+While the data collected was mainly absence/presence, through data visualization, statistical analysis, and performance evaluations, gaps in knowledge surrounding *Salmonella* persistence outside of the poultry industry can be closed.
 
 ## **Strengths and Limitations**
+While information on *Salmonella* persistence outside of the commercial poultry industry is relatively minimal, our work with the Picnic Table and Feeder Trials, provides some interesting insight into the persistence of this pathogen and possible sources of transmission during avian outbreaks (e.g., feeders).
+Because both experiments were trials, limited funding was allocated to each and a small sample size was a result of such.
+Small sample sizes limit data interpretation.
+Despite the small sample size, all models were analyzed with the model_check() function and most produced the expected results.
+However, some residuals in certain models were not normally distributed which can lead to over fitting.
+
+Despite the limitations of these trials, steps were taken to close knowledge gaps.
+Trials, such as the ones conducted, can lay the groundwork for future studies that improve upon our sampling and collection methodology, as well as statistical analysis.
+Through the information garnered from these trials as well as future studies, more insight into *Salmonella* persistence and ways to mitigate it in our communities will inevitably be developed and enacted.
 
 ## **Conclusions**
+Globally, salmonellosis outbreaks are common in aquatic and songbird populations and Salmonella continues to be a significant pathogen for public health.
+Notably, in 2021 a Salmonella Typhimurium outbreak in people was associated with the use and handling of bird feeders.
+Research is needed to understand the interaction between humans, birds, and their interface.
+Key to preventing outbreaks is understanding transmission dynamics, specifically, how long Salmonella persists on various materials.
+Persistence has been studied in food systems and commercial poultry operations—neither are relevant to free-living birds or people in contact with shared areas.
+Through further research and investment some of these pressing questions may be answered.
+Saving lives of various species globally.
 

### 2023-05-06 5. Manuscript/supplement/Supplementary_Material.qmd
Good Night, and Good Luck

--- 2023-03-10 products/manuscript/supplement/Supplementary_Material.qmd
@@
 ---
-title: "Supplement to Example Manuscript Template for a Data Analysis Project"
-
-subtitle: ""
-author: Andreas Handel
+title: "Supplement to *Salmonella* Environmental Persistence Informs Management Relevant to Avian and Public Health for a Data Analysis Project"
+author: Kimberly Perez
 date: "`r Sys.Date()`"
-format:
-  pdf:
-    toc: false
-    number-sections: true
-    highlight-style: github
-bibliography: ../../dataanalysis_template_references.bib
-csl: ../../vancouver.csl
+output: html_document
 ---
@@
 # **Overview**
-Overview of what I will include for readers needs to be included here.
+The information included on this .qmd file is code that was written but not utilized in final version of my manuscript.
+The majority of what you will find on this page comes directly from my "StatisticalAnalysis.qmd".
 
-# **Code and file information**
+Loading in the packages we may need...
+```{r}
+#Load needed packages
+library(tidyverse)
+library(ggplot2) #For plotting/data vis.
+library(broom) #For cleaning up output from lm()
+library(dotwhisker) #Visualization for regression outputs
+library(here) #Setting paths/for data loading/saving
+library(tidymodels) #Great package for models and other resources used for fitting
+library(tidyr) #Helps with data wrangling
+library(gtsummary) #To create tables 
+library(dplyr)
+library(vip)
+library(performance) #Model evaluation
+library(jtools)
+library(broom)
+library(broom.mixed) #Converts fitted to tiddy dfs
+library(finalfit) #Needed for Time to Event Analysis 
+library(forcats) #Needed for Time to Event Analysis 
+library(survival) #Needed for Time to Event Analysis 
+library(lubridate)
+library(ggsurvfit)
+library(gt)
+library(knitr) #For tables
+```
 
-Here is where I will provide a step-by-step document so others will be able to follow along and easily reproduce my code.
-I will utilize some of the text I included in the main text/Manuscript.qmd and will expand upon it as I work through this project.
+We need the data from both trials, so let's load them in using here() 
+```{r}
+#Picnic Table Data
+PT_Data<- readRDS(here::here("2. Clean Data","PTprocessed.rds"))
 
-{{< pagebreak >}}
+#Feeder Data 
+FD_Data <-readRDS(here::here("2. Clean Data","FDprocessed.rds"))
 
-# **Additional Method Details**
-{{< pagebreak >}}
+glimpse(PT_Data)
 
-# **Additional results**
-{{< pagebreak >}}
+glimpse (FD_Data)
+```
 
-# **Discussion**
-{{< pagebreak >}}
+```{r}
+#### Model fit another way
+# fit linear model using Salmonella_Pos as outcome, Table as predictor
+lmfit3.0 <- lm(Salmonella_Positive~Table, PT_Data) 
+lmfit3.0
+plot(lmfit3.0)
 
-# **References**
+# Here we place results from the above fit into a data frame via tidy function
+lmtable1 <- broom::tidy(lmfit3.0)
+
+#Let's check out the results
+print(lmtable1)
+
+#A different way 
+# I will do the same as above, however, I am now adding persistence pile
+lmfit2 <- lm(Salmonella_Positive~Original_Persistence_Pile+Table, PT_Data) 
+plot(lmfit2)
+
+# Convert to a data frame via tidy function
+lmtable2 <- broom::tidy(lmfit2)
+
+#Results
+print(lmtable2)
+```
+
+This section is a continuation of what was produced above, however, we will now be utilizing the Feeder Trial Data
+
+```{r}
+#Doing the same for FEEDER DATA
+lmog_FD<-linear_reg() %>% 
+  set_engine("lm")
+
+model_lmFD<- linear_reg()
+
+lmfit_FD<- model_lmFD %>% #Training; outcome: Salmonella Pos and Main Predictor Table
+  fit(Absence_0_Presence_1~Feeder_number, FD_Data)
+
+lmfit_FD #Let's view what we just did
+
+glimpse(lmfit_FD)
+
+tidy(lmfit_FD) #We can use the tidy() function to produce the output/ summary statistics
+
+plot_summs(lmfit_FD)
+
+#Checking model performance: FEEDER
+check_model(lmfit_FD$fit)
+```
+
+**Time to Event Analysis: Picnic Table Trial**
+```{r}
+
+PT_Data_TTE<- #First let's format the dates 
+  PT_Data %>% 
+  mutate(
+    Date= ymd(Date))
+
+PT_Data_TTE$Date<- as.numeric(PT_Data_TTE$Date) #Let's make these numeric
+
+PT_Data_TTE <- #I want to add a column displaying time between events (by day)
+  PT_Data_TTE %>%
+  mutate(
+    os_day= as.duration(Date)/ddays(1))
+
+PT_Data_TTE$os_day<- as.numeric(PT_Data_TTE$os_day) #Again let's make sure our new column is in the proper format
+
+surv1<- Surv(PT_Data_TTE$Date,PT_Data_TTE$Salmonella_Positive)
+
+#Let's visualize!
+sfit1<- survfit(Surv(Date, Salmonella_Positive) ~  1, data=PT_Data_TTE)  %>% 
+  ggsurvfit() + 
+  labs(
+    x = "Days",
+    y = "Overall Salmonella Persistence") + #Let's add some labels
+  add_confidence_interval() +
+  add_risktable()
+
+#I am not confident with this graphic so I will include this in my supplemental section
+```
+
+```{r}
+glimpse(PT_Data)
+
+PT_Data = PT_Data %>% 
+ mutate(
+  salmonella_crr= case_when(
+   Salmonella_Positive ==  0~1,
+    Salmonella_Positive == 1~1,
+     TRUE ~ 1))
+
+dependent_crr= "Surv(Date, salmonella_crr)"
+explanatory= c("Table", "Quadrant", "Sample_Type")
+
+PT_Data$Date<- as.numeric(PT_Data$Date)
+
+PT_Data %>% finalfit(dependent_crr, explanatory)
+
+PT_Data %>% 
+   finalfit(dependent_crr, explanatory, add_dependent_label = FALSE) %>% 
+    rename("Overall survival" = label) %>% 
+    rename(" " = levels) %>% 
+    rename("  " = all)
+
+survfit(Surv(Date, Salmonella_Positive)~1, data=PT_Data)
+
+PT_MedSurv<-PT_Data %>%
+ filter(salmonella_crr==1) %>%
+  summarize(median_surv=median(Date))
+
+PT_MedSurv <- ((PT_MedSurv)/31556952000)*100
+
+PT_Surv_Table<- gt(PT_MedSurv) %>%
+   tab_header(
+    title = "Median Survival of Salmonella",
+    subtitle = ("01/02/2020 to 01/09/2020")) %>%
+  cols_label(
+    median_surv= "Median Salmonella Survival (days)")
+
+PT_Surv_Table # persistence seems to be roughly 5 days
+```
+
+**Time to Event Analysis: Feeder Trial**
+```{r}
+FD_Data_TTE<- #First let's format the dates 
+  FD_Data %>% 
+  mutate(
+    Date= ymd(Collection_Date))
+
+FD_Data_TTE$Collection_Date<- as.numeric(FD_Data_TTE$Collection_Date) #Let's make these numeric
+
+FD_Data_TTE <- #I want to add a column displaying time between events (by day)
+  FD_Data_TTE %>%
+  mutate(
+    os_day= as.duration(Collection_Date)/ddays(1))
+
+FD_Data_TTE$os_day<- as.numeric(FD_Data_TTE$os_day) #Again let's make sure our new column is in the proper format
+
+surv1_FD<- Surv(FD_Data_TTE$Collection_Date,FD_Data_TTE$Absence_0_Presence_1)
+
+#Let's visualize!
+sfit1<- survfit(Surv(Collection_Date, Absence_0_Presence_1) ~  1, data=FD_Data_TTE)  %>% 
+  ggsurvfit() + 
+  labs(
+    x = "Days",
+    y = "Overall Salmonella Persistence") + #Let's add some labels
+  add_confidence_interval() +
+  add_risktable()
+
+sfit1
+```
+
+```{r}
+FD_Data = FD_Data %>% 
+ mutate(
+  salmonella_crr= case_when(
+   Absence_0_Presence_1 ==  0~1,
+    Absence_0_Presence_1 == 1~1,
+     TRUE ~ 1))
+
+dependent_crr= "Surv(Collection_Date, salmonella_crr)"
+explanatory= c("Feeder_Type")
+
+FD_Data$Collection_Date<- as.numeric(FD_Data$Collection_Date)
+
+FD_Data %>% finalfit(dependent_crr, explanatory)
+
+FD_Data %>% 
+   finalfit(dependent_crr, explanatory, add_dependent_label = FALSE) %>% 
+    rename("Overall survival" = label) %>% 
+    rename(" " = levels) %>% 
+    rename("  " = all)
+
+survfit(Surv(Collection_Date, Absence_0_Presence_1)~1, data=FD_Data)
+
+FD_MedSurv<-FD_Data %>%
+ filter(salmonella_crr==1) %>%
+  summarize(median_surv=median(Collection_Date))
+
+FD_MedSurv <- ((FD_MedSurv)/604800)*100 #Persistence seems to be roughly 3 days for the FD
+
+FD_Surv_Table<- gt(FD_MedSurv) %>%
+   tab_header(
+    title = "Median Survival of Salmonella",
+    subtitle = ("02/22/2022 to 03/14/2022")) %>%
+  cols_label(
+    median_surv= "Median Salmonella Survival (days)")
+
+FD_Surv_Table # seems to be roughly 3 days
+```
+
+**Cox regression: Picnic Table Trial**
+
+```{r}
+PT_Data$Date<- as.numeric(PT_Data$Date)
+
+coxph(Surv(Date, Salmonella_Positive) ~ Sample_Type, PT_Data)
+
+CR_PT1<-coxph(Surv(Date, Salmonella_Positive) ~ Sample_Type, PT_Data) %>%
+  tbl_regression(exp=TRUE)
+
+CR_PT1
+```
+
+**Cox regression: Feeder Trial**
+
+```{r}
+FD_Data$Collection_Date<- as.numeric(FD_Data$Collection_Date)
+
+coxph(Surv(Collection_Date, Absence_0_Presence_1) ~ Feeder_Type, FD_Data)
+
+CR_FD1<- coxph(Surv(Collection_Date, Absence_0_Presence_1) ~ Feeder_Type, FD_Data) %>%
+  tbl_regression(exp=TRUE)
+
+CR_FD1
+
+CR_FD1%>%
+  saveRDS(here("4. Results", "FD_Cox.rds"))
+```

### 2023-05-06 5. Manuscript/supplement/Supplementary_Material.tex
Good Night, and Good Luck

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.33,0.33}{\textbf{#1}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.42,0.45,0.49}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.84,0.23,0.29}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.36,0.77}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.84,0.23,0.29}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.01,0.18,0.38}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.42,0.45,0.49}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.42,0.45,0.49}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.36,0.77}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.84,0.23,0.29}{#1}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.84,0.23,0.29}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.36,0.77}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.42,0.45,0.49}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.33,0.33}{\underline{#1}}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.84,0.23,0.29}{\textbf{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.36,0.77}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.44,0.26,0.76}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.01,0.18,0.38}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.42,0.45,0.49}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.84,0.23,0.29}{#1}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.14,0.16,0.18}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.14,0.16,0.18}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.44,0.26,0.76}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.84,0.23,0.29}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.42,0.45,0.49}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.36,0.77}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.01,0.18,0.38}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.01,0.18,0.38}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.89,0.38,0.04}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.01,0.18,0.38}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{1.00,0.33,0.33}{#1}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Supplement to Salmonella Environmental Persistence Informs Management Relevant to Avian and Public Health for a Data Analysis Project},
  pdfauthor={Kimberly Perez},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Supplement to \emph{Salmonella} Environmental Persistence Informs
Management Relevant to Avian and Public Health for a Data Analysis
Project}
\author{Kimberly Perez}
\date{5/5/23}

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden, frame hidden, sharp corners]}{\end{tcolorbox}}\fi

\hypertarget{overview}{%
\section{\texorpdfstring{\textbf{Overview}}{Overview}}\label{overview}}

The information included on this .qmd file is code that was written but
not utilized in final version of my manuscript. The majority of what you
will find on this page comes directly from my
``StatisticalAnalysis.qmd''.

Loading in the packages we may need\ldots{}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#Load needed packages}
\FunctionTok{library}\NormalTok{(tidyverse)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
-- Attaching packages --------------------------------------- tidyverse 1.3.2 --
v ggplot2 3.4.0      v purrr   1.0.1 
v tibble  3.1.8      v dplyr   1.0.10
v tidyr   1.3.0      v stringr 1.5.0 
v readr   2.1.4      v forcats 0.5.2 
\end{verbatim}

\begin{verbatim}
Warning: package 'tidyr' was built under R version 4.2.3
\end{verbatim}

\begin{verbatim}
-- Conflicts ------------------------------------------ tidyverse_conflicts() --
x dplyr::filter() masks stats::filter()
x dplyr::lag()    masks stats::lag()
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(ggplot2) }\CommentTok{\#For plotting/data vis.}
\FunctionTok{library}\NormalTok{(broom) }\CommentTok{\#For cleaning up output from lm()}
\FunctionTok{library}\NormalTok{(dotwhisker) }\CommentTok{\#Visualization for regression outputs}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
Warning: package 'dotwhisker' was built under R version 4.2.3
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(here) }\CommentTok{\#Setting paths/for data loading/saving}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
here() starts at C:/GitHub/MADA/KimberlyPerez-MADA-project
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(tidymodels) }\CommentTok{\#Great package for models and other resources used for fitting}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
-- Attaching packages -------------------------------------- tidymodels 1.0.0 --
v dials        1.1.0     v rsample      1.1.1
v infer        1.0.4     v tune         1.0.1
v modeldata    1.1.0     v workflows    1.1.2
v parsnip      1.0.3     v workflowsets 1.0.0
v recipes      1.0.4     v yardstick    1.1.0
-- Conflicts ----------------------------------------- tidymodels_conflicts() --
x scales::discard() masks purrr::discard()
x dplyr::filter()   masks stats::filter()
x recipes::fixed()  masks stringr::fixed()
x dplyr::lag()      masks stats::lag()
x yardstick::spec() masks readr::spec()
x recipes::step()   masks stats::step()
* Dig deeper into tidy modeling with R at https://www.tmwr.org
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(tidyr) }\CommentTok{\#Helps with data wrangling}
\FunctionTok{library}\NormalTok{(gtsummary) }\CommentTok{\#To create tables }
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
#BlackLivesMatter

Attaching package: 'gtsummary'

The following objects are masked from 'package:recipes':

    all_double, all_factor, all_integer, all_logical, all_numeric
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(dplyr)}
\FunctionTok{library}\NormalTok{(vip)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}

Attaching package: 'vip'

The following object is masked from 'package:utils':

    vi
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(performance) }\CommentTok{\#Model evaluation}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}

Attaching package: 'performance'

The following objects are masked from 'package:yardstick':

    mae, rmse
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(jtools)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}

Attaching package: 'jtools'

The following object is masked from 'package:yardstick':

    get_weights
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(broom)}
\FunctionTok{library}\NormalTok{(broom.mixed) }\CommentTok{\#Converts fitted to tiddy dfs}
\FunctionTok{library}\NormalTok{(finalfit) }\CommentTok{\#Needed for Time to Event Analysis }
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
Warning: package 'finalfit' was built under R version 4.2.3
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(forcats) }\CommentTok{\#Needed for Time to Event Analysis }
\FunctionTok{library}\NormalTok{(survival) }\CommentTok{\#Needed for Time to Event Analysis }
\FunctionTok{library}\NormalTok{(lubridate)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
Loading required package: timechange

Attaching package: 'lubridate'

The following objects are masked from 'package:base':

    date, intersect, setdiff, union
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(ggsurvfit)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
Warning: package 'ggsurvfit' was built under R version 4.2.3
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(gt)}
\FunctionTok{library}\NormalTok{(knitr) }\CommentTok{\#For tables}
\end{Highlighting}
\end{Shaded}

We need the data from both trials, so let's load them in using here()

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#Picnic Table Data}
\NormalTok{PT\_Data}\OtherTok{\textless{}{-}} \FunctionTok{readRDS}\NormalTok{(here}\SpecialCharTok{::}\FunctionTok{here}\NormalTok{(}\StringTok{"2. Clean Data"}\NormalTok{,}\StringTok{"PTprocessed.rds"}\NormalTok{))}

\CommentTok{\#Feeder Data }
\NormalTok{FD\_Data }\OtherTok{\textless{}{-}}\FunctionTok{readRDS}\NormalTok{(here}\SpecialCharTok{::}\FunctionTok{here}\NormalTok{(}\StringTok{"2. Clean Data"}\NormalTok{,}\StringTok{"FDprocessed.rds"}\NormalTok{))}

\FunctionTok{glimpse}\NormalTok{(PT\_Data)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
Rows: 192
Columns: 7
$ ID                        <chr> "DPT1Q1 PER 1-2-2020", "DPT1Q2 PER 1-2-2020"~
$ Date                      <dttm> 2020-01-02, 2020-01-02, 2020-01-02, 2020-01~
$ Table                     <dbl> 1, 1, 1, 1, 1, 1, 1, 1, 2, 2, 2, 2, 2, 2, 2,~
$ Quadrant                  <dbl> 1, 2, 3, 4, 1, 2, 3, 4, 1, 2, 3, 4, 1, 2, 3,~
$ Sample_Type               <chr> "Persistence", "Persistence", "Persistence",~
$ Original_Persistence_Pile <dbl> 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,~
$ Salmonella_Positive       <dbl> 0, 0, 0, 0, 0, 1, 1, 0, 0, 0, 0, 1, 1, 0, 0,~
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{glimpse}\NormalTok{ (FD\_Data)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
Rows: 144
Columns: 4
$ Feeder_number        <dbl> 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 1, 2, 3, 4~
$ Collection_Date      <date> 2022-02-22, 2022-02-22, 2022-02-22, 2022-02-22, ~
$ Absence_0_Presence_1 <dbl> 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 1, 1, 0~
$ Feeder_Type          <chr> "Wood_coated", "Wood", "Plastic_coated", "Plastic~
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\DocumentationTok{\#\#\#\# Model fit another way}
\CommentTok{\# fit linear model using Salmonella\_Pos as outcome, Table as predictor}
\NormalTok{lmfit3}\FloatTok{.0} \OtherTok{\textless{}{-}} \FunctionTok{lm}\NormalTok{(Salmonella\_Positive}\SpecialCharTok{\textasciitilde{}}\NormalTok{Table, PT\_Data) }
\NormalTok{lmfit3}\FloatTok{.0}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}

Call:
lm(formula = Salmonella_Positive ~ Table, data = PT_Data)

Coefficients:
(Intercept)        Table  
    0.15625      0.02344  
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot}\NormalTok{(lmfit3}\FloatTok{.0}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

{\centering \includegraphics{Supplementary_Material_files/figure-pdf/unnamed-chunk-3-1.pdf}

}

\end{figure}

\begin{figure}[H]

{\centering \includegraphics{Supplementary_Material_files/figure-pdf/unnamed-chunk-3-2.pdf}

}

\end{figure}

\begin{figure}[H]

{\centering \includegraphics{Supplementary_Material_files/figure-pdf/unnamed-chunk-3-3.pdf}

}

\end{figure}

\begin{figure}[H]

{\centering \includegraphics{Supplementary_Material_files/figure-pdf/unnamed-chunk-3-4.pdf}

}

\end{figure}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Here we place results from the above fit into a data frame via tidy function}
\NormalTok{lmtable1 }\OtherTok{\textless{}{-}}\NormalTok{ broom}\SpecialCharTok{::}\FunctionTok{tidy}\NormalTok{(lmfit3}\FloatTok{.0}\NormalTok{)}

\CommentTok{\#Let\textquotesingle{}s check out the results}
\FunctionTok{print}\NormalTok{(lmtable1)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 2 x 5
  term        estimate std.error statistic p.value
  <chr>          <dbl>     <dbl>     <dbl>   <dbl>
1 (Intercept)   0.156     0.0771     2.03   0.0442
2 Table         0.0234    0.0357     0.656  0.512 
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#A different way }
\CommentTok{\# I will do the same as above, however, I am now adding persistence pile}
\NormalTok{lmfit2 }\OtherTok{\textless{}{-}} \FunctionTok{lm}\NormalTok{(Salmonella\_Positive}\SpecialCharTok{\textasciitilde{}}\NormalTok{Original\_Persistence\_Pile}\SpecialCharTok{+}\NormalTok{Table, PT\_Data) }
\FunctionTok{plot}\NormalTok{(lmfit2)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

{\centering \includegraphics{Supplementary_Material_files/figure-pdf/unnamed-chunk-3-5.pdf}

}

\end{figure}

\begin{figure}[H]

{\centering \includegraphics{Supplementary_Material_files/figure-pdf/unnamed-chunk-3-6.pdf}

}

\end{figure}

\begin{figure}[H]

{\centering \includegraphics{Supplementary_Material_files/figure-pdf/unnamed-chunk-3-7.pdf}

}

\end{figure}

\begin{figure}[H]

{\centering \includegraphics{Supplementary_Material_files/figure-pdf/unnamed-chunk-3-8.pdf}

}

\end{figure}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Convert to a data frame via tidy function}
\NormalTok{lmtable2 }\OtherTok{\textless{}{-}}\NormalTok{ broom}\SpecialCharTok{::}\FunctionTok{tidy}\NormalTok{(lmfit2)}

\CommentTok{\#Results}
\FunctionTok{print}\NormalTok{(lmtable2)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 3 x 5
  term                      estimate std.error statistic p.value
  <chr>                        <dbl>     <dbl>     <dbl>   <dbl>
1 (Intercept)                 0.231     0.0803     2.87  0.00456
2 Original_Persistence_Pile  -0.162     0.0579    -2.80  0.00567
3 Table                       0.0209    0.0351     0.596 0.552  
\end{verbatim}

This section is a continuation of what was produced above, however, we
will now be utilizing the Feeder Trial Data

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#Doing the same for FEEDER DATA}
\NormalTok{lmog\_FD}\OtherTok{\textless{}{-}}\FunctionTok{linear\_reg}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{set\_engine}\NormalTok{(}\StringTok{"lm"}\NormalTok{)}

\NormalTok{model\_lmFD}\OtherTok{\textless{}{-}} \FunctionTok{linear\_reg}\NormalTok{()}

\NormalTok{lmfit\_FD}\OtherTok{\textless{}{-}}\NormalTok{ model\_lmFD }\SpecialCharTok{\%\textgreater{}\%} \CommentTok{\#Training; outcome: Salmonella Pos and Main Predictor Table}
  \FunctionTok{fit}\NormalTok{(Absence\_0\_Presence\_1}\SpecialCharTok{\textasciitilde{}}\NormalTok{Feeder\_number, FD\_Data) }

\NormalTok{lmfit\_FD }\CommentTok{\#Let\textquotesingle{}s view what we just did}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
parsnip model object

Call:
stats::lm(formula = Absence_0_Presence_1 ~ Feeder_number, data = data)

Coefficients:
  (Intercept)  Feeder_number  
      0.06439        0.01253  
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{glimpse}\NormalTok{(lmfit\_FD)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
List of 5
 $ lvl    : NULL
 $ spec   :List of 6
  ..$ args                 :List of 2
  .. ..$ penalty: language ~NULL
  .. .. ..- attr(*, ".Environment")=<environment: R_EmptyEnv> 
  .. ..$ mixture: language ~NULL
  .. .. ..- attr(*, ".Environment")=<environment: R_EmptyEnv> 
  ..$ mode                 : chr "regression"
  ..$ user_specified_mode  : logi FALSE
  ..$ method               :List of 3
  .. ..$ libs: chr "stats"
  .. ..$ fit :List of 5
  .. ..$ pred:List of 4
  ..$ engine               : chr "lm"
  ..$ user_specified_engine: logi FALSE
  ..- attr(*, "class")= chr [1:2] "linear_reg" "model_spec"
 $ fit    :List of 12
  ..$ coefficients : Named num [1:2] 0.0644 0.0125
  .. ..- attr(*, "names")= chr [1:2] "(Intercept)" "Feeder_number"
  ..$ residuals    : Named num [1:144] 0.923 0.911 0.898 0.885 0.873 ...
  .. ..- attr(*, "names")= chr [1:144] "1" "2" "3" "4" ...
  ..$ effects      : Named num [1:144] -1.75 0.519 0.77 0.776 0.782 ...
  .. ..- attr(*, "names")= chr [1:144] "(Intercept)" "Feeder_number" "" "" ...
  ..$ rank         : int 2
  ..$ fitted.values: Named num [1:144] 0.0769 0.0895 0.102 0.1145 0.127 ...
  .. ..- attr(*, "names")= chr [1:144] "1" "2" "3" "4" ...
  ..$ assign       : int [1:2] 0 1
  ..$ qr           :List of 5
  .. ..$ qr   : num [1:144, 1:2] -12 0.0833 0.0833 0.0833 0.0833 ...
  .. .. ..- attr(*, "dimnames")=List of 2
  .. .. ..- attr(*, "assign")= int [1:2] 0 1
  .. ..$ qraux: num [1:2] 1.08 1.1
  .. ..$ pivot: int [1:2] 1 2
  .. ..$ tol  : num 1e-07
  .. ..$ rank : int 2
  .. ..- attr(*, "class")= chr "qr"
  ..$ df.residual  : int 142
  ..$ xlevels      : Named list()
  ..$ call         : language stats::lm(formula = Absence_0_Presence_1 ~ Feeder_number, data = data)
  ..$ terms        :Classes 'terms', 'formula'  language Absence_0_Presence_1 ~ Feeder_number
  .. .. ..- attr(*, "variables")= language list(Absence_0_Presence_1, Feeder_number)
  .. .. ..- attr(*, "factors")= int [1:2, 1] 0 1
  .. .. .. ..- attr(*, "dimnames")=List of 2
  .. .. ..- attr(*, "term.labels")= chr "Feeder_number"
  .. .. ..- attr(*, "order")= int 1
  .. .. ..- attr(*, "intercept")= int 1
  .. .. ..- attr(*, "response")= int 1
  .. .. ..- attr(*, ".Environment")=<environment: 0x00000271ddcb4e18> 
  .. .. ..- attr(*, "predvars")= language list(Absence_0_Presence_1, Feeder_number)
  .. .. ..- attr(*, "dataClasses")= Named chr [1:2] "numeric" "numeric"
  .. .. .. ..- attr(*, "names")= chr [1:2] "Absence_0_Presence_1" "Feeder_number"
  ..$ model        :'data.frame':   144 obs. of  2 variables:
  .. ..$ Absence_0_Presence_1: num [1:144] 1 1 1 1 1 1 1 1 1 1 ...
  .. ..$ Feeder_number       : num [1:144] 1 2 3 4 5 6 7 8 9 10 ...
  .. ..- attr(*, "terms")=Classes 'terms', 'formula'  language Absence_0_Presence_1 ~ Feeder_number
  .. .. .. ..- attr(*, "variables")= language list(Absence_0_Presence_1, Feeder_number)
  .. .. .. ..- attr(*, "factors")= int [1:2, 1] 0 1
  .. .. .. .. ..- attr(*, "dimnames")=List of 2
  .. .. .. ..- attr(*, "term.labels")= chr "Feeder_number"
  .. .. .. ..- attr(*, "order")= int 1
  .. .. .. ..- attr(*, "intercept")= int 1
  .. .. .. ..- attr(*, "response")= int 1
  .. .. .. ..- attr(*, ".Environment")=<environment: 0x00000271ddcb4e18> 
  .. .. .. ..- attr(*, "predvars")= language list(Absence_0_Presence_1, Feeder_number)
  .. .. .. ..- attr(*, "dataClasses")= Named chr [1:2] "numeric" "numeric"
  .. .. .. .. ..- attr(*, "names")= chr [1:2] "Absence_0_Presence_1" "Feeder_number"
  ..- attr(*, "class")= chr "lm"
 $ preproc:List of 1
  ..$ y_var: chr "Absence_0_Presence_1"
 $ elapsed:List of 1
  ..$ elapsed: num NA
 - attr(*, "class")= chr [1:2] "_lm" "model_fit"
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{tidy}\NormalTok{(lmfit\_FD) }\CommentTok{\#We can use the tidy() function to produce the output/ summary statistics}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 2 x 5
  term          estimate std.error statistic p.value
  <chr>            <dbl>     <dbl>     <dbl>   <dbl>
1 (Intercept)     0.0644   0.0627       1.03   0.306
2 Feeder_number   0.0125   0.00852      1.47   0.143
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot\_summs}\NormalTok{(lmfit\_FD)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

{\centering \includegraphics{Supplementary_Material_files/figure-pdf/unnamed-chunk-4-1.pdf}

}

\end{figure}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#Checking model performance: FEEDER}
\FunctionTok{check\_model}\NormalTok{(lmfit\_FD}\SpecialCharTok{$}\NormalTok{fit)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

{\centering \includegraphics{Supplementary_Material_files/figure-pdf/unnamed-chunk-4-2.pdf}

}

\end{figure}

\textbf{Time to Event Analysis: Picnic Table Trial}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{PT\_Data\_TTE}\OtherTok{\textless{}{-}} \CommentTok{\#First let\textquotesingle{}s format the dates }
\NormalTok{  PT\_Data }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{mutate}\NormalTok{(}
    \AttributeTok{Date=} \FunctionTok{ymd}\NormalTok{(Date))}

\NormalTok{PT\_Data\_TTE}\SpecialCharTok{$}\NormalTok{Date}\OtherTok{\textless{}{-}} \FunctionTok{as.numeric}\NormalTok{(PT\_Data\_TTE}\SpecialCharTok{$}\NormalTok{Date) }\CommentTok{\#Let\textquotesingle{}s make these numeric}

\NormalTok{PT\_Data\_TTE }\OtherTok{\textless{}{-}} \CommentTok{\#I want to add a column displaying time between events (by day)}
\NormalTok{  PT\_Data\_TTE }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}
    \AttributeTok{os\_day=} \FunctionTok{as.duration}\NormalTok{(Date)}\SpecialCharTok{/}\FunctionTok{ddays}\NormalTok{(}\DecValTok{1}\NormalTok{))}

\NormalTok{PT\_Data\_TTE}\SpecialCharTok{$}\NormalTok{os\_day}\OtherTok{\textless{}{-}} \FunctionTok{as.numeric}\NormalTok{(PT\_Data\_TTE}\SpecialCharTok{$}\NormalTok{os\_day) }\CommentTok{\#Again let\textquotesingle{}s make sure our new column is in the proper format}

\NormalTok{surv1}\OtherTok{\textless{}{-}} \FunctionTok{Surv}\NormalTok{(PT\_Data\_TTE}\SpecialCharTok{$}\NormalTok{Date,PT\_Data\_TTE}\SpecialCharTok{$}\NormalTok{Salmonella\_Positive) }

\CommentTok{\#Let\textquotesingle{}s visualize!}
\NormalTok{sfit1}\OtherTok{\textless{}{-}} \FunctionTok{survfit}\NormalTok{(}\FunctionTok{Surv}\NormalTok{(Date, Salmonella\_Positive) }\SpecialCharTok{\textasciitilde{}}  \DecValTok{1}\NormalTok{, }\AttributeTok{data=}\NormalTok{PT\_Data\_TTE)  }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{ggsurvfit}\NormalTok{() }\SpecialCharTok{+} 
  \FunctionTok{labs}\NormalTok{(}
    \AttributeTok{x =} \StringTok{"Days"}\NormalTok{,}
    \AttributeTok{y =} \StringTok{"Overall Salmonella Persistence"}\NormalTok{) }\SpecialCharTok{+} \CommentTok{\#Let\textquotesingle{}s add some labels}
  \FunctionTok{add\_confidence\_interval}\NormalTok{() }\SpecialCharTok{+}
  \FunctionTok{add\_risktable}\NormalTok{()}

\CommentTok{\#I am not confident with this graphic so I will include this in my supplemental section}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{glimpse}\NormalTok{(PT\_Data)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
Rows: 192
Columns: 7
$ ID                        <chr> "DPT1Q1 PER 1-2-2020", "DPT1Q2 PER 1-2-2020"~
$ Date                      <dttm> 2020-01-02, 2020-01-02, 2020-01-02, 2020-01~
$ Table                     <dbl> 1, 1, 1, 1, 1, 1, 1, 1, 2, 2, 2, 2, 2, 2, 2,~
$ Quadrant                  <dbl> 1, 2, 3, 4, 1, 2, 3, 4, 1, 2, 3, 4, 1, 2, 3,~
$ Sample_Type               <chr> "Persistence", "Persistence", "Persistence",~
$ Original_Persistence_Pile <dbl> 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,~
$ Salmonella_Positive       <dbl> 0, 0, 0, 0, 0, 1, 1, 0, 0, 0, 0, 1, 1, 0, 0,~
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{PT\_Data }\OtherTok{=}\NormalTok{ PT\_Data }\SpecialCharTok{\%\textgreater{}\%} 
 \FunctionTok{mutate}\NormalTok{(}
  \AttributeTok{salmonella\_crr=} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{   Salmonella\_Positive }\SpecialCharTok{==}  \DecValTok{0}\SpecialCharTok{\textasciitilde{}}\DecValTok{1}\NormalTok{,}
\NormalTok{    Salmonella\_Positive }\SpecialCharTok{==} \DecValTok{1}\SpecialCharTok{\textasciitilde{}}\DecValTok{1}\NormalTok{,}
     \ConstantTok{TRUE} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{))}

\NormalTok{dependent\_crr}\OtherTok{=} \StringTok{"Surv(Date, salmonella\_crr)"}
\NormalTok{explanatory}\OtherTok{=} \FunctionTok{c}\NormalTok{(}\StringTok{"Table"}\NormalTok{, }\StringTok{"Quadrant"}\NormalTok{, }\StringTok{"Sample\_Type"}\NormalTok{)}

\NormalTok{PT\_Data}\SpecialCharTok{$}\NormalTok{Date}\OtherTok{\textless{}{-}} \FunctionTok{as.numeric}\NormalTok{(PT\_Data}\SpecialCharTok{$}\NormalTok{Date)}

\NormalTok{PT\_Data }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{finalfit}\NormalTok{(dependent\_crr, explanatory)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
 Dependent: Surv(Date, salmonella_crr)                   all
                                 Table           1 64 (33.3)
                                                 2 64 (33.3)
                                                 3 64 (33.3)
                              Quadrant           1 48 (25.0)
                                                 2 48 (25.0)
                                                 3 48 (25.0)
                                                 4 48 (25.0)
                           Sample_Type Persistence 96 (50.0)
                                            Pooled 96 (50.0)
          HR (univariable)        HR (multivariable)
                         -                         -
 1.00 (0.71-1.41, p=1.000) 1.00 (0.71-1.41, p=1.000)
 1.00 (0.71-1.41, p=1.000) 1.00 (0.71-1.41, p=1.000)
                         -                         -
 1.00 (0.67-1.49, p=1.000) 1.00 (0.67-1.49, p=1.000)
 1.00 (0.67-1.49, p=1.000) 1.00 (0.67-1.49, p=1.000)
 1.00 (0.67-1.49, p=1.000) 1.00 (0.67-1.49, p=1.000)
                         -                         -
 1.00 (0.75-1.33, p=1.000) 1.00 (0.75-1.33, p=1.000)
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{PT\_Data }\SpecialCharTok{\%\textgreater{}\%} 
   \FunctionTok{finalfit}\NormalTok{(dependent\_crr, explanatory, }\AttributeTok{add\_dependent\_label =} \ConstantTok{FALSE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
    \FunctionTok{rename}\NormalTok{(}\StringTok{"Overall survival"} \OtherTok{=}\NormalTok{ label) }\SpecialCharTok{\%\textgreater{}\%} 
    \FunctionTok{rename}\NormalTok{(}\StringTok{" "} \OtherTok{=}\NormalTok{ levels) }\SpecialCharTok{\%\textgreater{}\%} 
    \FunctionTok{rename}\NormalTok{(}\StringTok{"  "} \OtherTok{=}\NormalTok{ all) }
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
 Overall survival                                HR (univariable)
            Table           1 64 (33.3)                         -
                            2 64 (33.3) 1.00 (0.71-1.41, p=1.000)
                            3 64 (33.3) 1.00 (0.71-1.41, p=1.000)
         Quadrant           1 48 (25.0)                         -
                            2 48 (25.0) 1.00 (0.67-1.49, p=1.000)
                            3 48 (25.0) 1.00 (0.67-1.49, p=1.000)
                            4 48 (25.0) 1.00 (0.67-1.49, p=1.000)
      Sample_Type Persistence 96 (50.0)                         -
                       Pooled 96 (50.0) 1.00 (0.75-1.33, p=1.000)
        HR (multivariable)
                         -
 1.00 (0.71-1.41, p=1.000)
 1.00 (0.71-1.41, p=1.000)
                         -
 1.00 (0.67-1.49, p=1.000)
 1.00 (0.67-1.49, p=1.000)
 1.00 (0.67-1.49, p=1.000)
                         -
 1.00 (0.75-1.33, p=1.000)
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{survfit}\NormalTok{(}\FunctionTok{Surv}\NormalTok{(Date, Salmonella\_Positive)}\SpecialCharTok{\textasciitilde{}}\DecValTok{1}\NormalTok{, }\AttributeTok{data=}\NormalTok{PT\_Data)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
Call: survfit(formula = Surv(Date, Salmonella_Positive) ~ 1, data = PT_Data)

       n events median 0.95LCL 0.95UCL
[1,] 192     39     NA      NA      NA
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{PT\_MedSurv}\OtherTok{\textless{}{-}}\NormalTok{PT\_Data }\SpecialCharTok{\%\textgreater{}\%}
 \FunctionTok{filter}\NormalTok{(salmonella\_crr}\SpecialCharTok{==}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{summarize}\NormalTok{(}\AttributeTok{median\_surv=}\FunctionTok{median}\NormalTok{(Date))}

\NormalTok{PT\_MedSurv }\OtherTok{\textless{}{-}}\NormalTok{ ((PT\_MedSurv)}\SpecialCharTok{/}\DecValTok{31556952000}\NormalTok{)}\SpecialCharTok{*}\DecValTok{100}

\NormalTok{PT\_Surv\_Table}\OtherTok{\textless{}{-}} \FunctionTok{gt}\NormalTok{(PT\_MedSurv) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{tab\_header}\NormalTok{(}
    \AttributeTok{title =} \StringTok{"Median Survival of Salmonella"}\NormalTok{,}
    \AttributeTok{subtitle =}\NormalTok{ (}\StringTok{"01/02/2020 to 01/09/2020"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{cols\_label}\NormalTok{(}
    \AttributeTok{median\_surv=} \StringTok{"Median Salmonella Survival (days)"}\NormalTok{) }

\NormalTok{PT\_Surv\_Table }\CommentTok{\# persistence seems to be roughly 5 days}
\end{Highlighting}
\end{Shaded}

\begin{longtable}{r}
\caption*{
{\large Median Survival of Salmonella} \\ 
{\small 01/02/2020 to 01/09/2020}
} \\ 
\toprule
Median Salmonella Survival (days) \\ 
\midrule
5.001198 \\ 
\bottomrule
\end{longtable}

\textbf{Time to Event Analysis: Feeder Trial}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{FD\_Data\_TTE}\OtherTok{\textless{}{-}} \CommentTok{\#First let\textquotesingle{}s format the dates }
\NormalTok{  FD\_Data }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{mutate}\NormalTok{(}
    \AttributeTok{Date=} \FunctionTok{ymd}\NormalTok{(Collection\_Date))}

\NormalTok{FD\_Data\_TTE}\SpecialCharTok{$}\NormalTok{Collection\_Date}\OtherTok{\textless{}{-}} \FunctionTok{as.numeric}\NormalTok{(FD\_Data\_TTE}\SpecialCharTok{$}\NormalTok{Collection\_Date) }\CommentTok{\#Let\textquotesingle{}s make these numeric}

\NormalTok{FD\_Data\_TTE }\OtherTok{\textless{}{-}} \CommentTok{\#I want to add a column displaying time between events (by day)}
\NormalTok{  FD\_Data\_TTE }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}
    \AttributeTok{os\_day=} \FunctionTok{as.duration}\NormalTok{(Collection\_Date)}\SpecialCharTok{/}\FunctionTok{ddays}\NormalTok{(}\DecValTok{1}\NormalTok{))}

\NormalTok{FD\_Data\_TTE}\SpecialCharTok{$}\NormalTok{os\_day}\OtherTok{\textless{}{-}} \FunctionTok{as.numeric}\NormalTok{(FD\_Data\_TTE}\SpecialCharTok{$}\NormalTok{os\_day) }\CommentTok{\#Again let\textquotesingle{}s make sure our new column is in the proper format}

\NormalTok{surv1\_FD}\OtherTok{\textless{}{-}} \FunctionTok{Surv}\NormalTok{(FD\_Data\_TTE}\SpecialCharTok{$}\NormalTok{Collection\_Date,FD\_Data\_TTE}\SpecialCharTok{$}\NormalTok{Absence\_0\_Presence\_1) }

\CommentTok{\#Let\textquotesingle{}s visualize!}
\NormalTok{sfit1}\OtherTok{\textless{}{-}} \FunctionTok{survfit}\NormalTok{(}\FunctionTok{Surv}\NormalTok{(Collection\_Date, Absence\_0\_Presence\_1) }\SpecialCharTok{\textasciitilde{}}  \DecValTok{1}\NormalTok{, }\AttributeTok{data=}\NormalTok{FD\_Data\_TTE)  }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{ggsurvfit}\NormalTok{() }\SpecialCharTok{+} 
  \FunctionTok{labs}\NormalTok{(}
    \AttributeTok{x =} \StringTok{"Days"}\NormalTok{,}
    \AttributeTok{y =} \StringTok{"Overall Salmonella Persistence"}\NormalTok{) }\SpecialCharTok{+} \CommentTok{\#Let\textquotesingle{}s add some labels}
  \FunctionTok{add\_confidence\_interval}\NormalTok{() }\SpecialCharTok{+}
  \FunctionTok{add\_risktable}\NormalTok{()}

\NormalTok{sfit1}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

{\centering \includegraphics{Supplementary_Material_files/figure-pdf/unnamed-chunk-7-1.pdf}

}

\end{figure}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{FD\_Data }\OtherTok{=}\NormalTok{ FD\_Data }\SpecialCharTok{\%\textgreater{}\%} 
 \FunctionTok{mutate}\NormalTok{(}
  \AttributeTok{salmonella\_crr=} \FunctionTok{case\_when}\NormalTok{(}
\NormalTok{   Absence\_0\_Presence\_1 }\SpecialCharTok{==}  \DecValTok{0}\SpecialCharTok{\textasciitilde{}}\DecValTok{1}\NormalTok{,}
\NormalTok{    Absence\_0\_Presence\_1 }\SpecialCharTok{==} \DecValTok{1}\SpecialCharTok{\textasciitilde{}}\DecValTok{1}\NormalTok{,}
     \ConstantTok{TRUE} \SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{))}

\NormalTok{dependent\_crr}\OtherTok{=} \StringTok{"Surv(Collection\_Date, salmonella\_crr)"}
\NormalTok{explanatory}\OtherTok{=} \FunctionTok{c}\NormalTok{(}\StringTok{"Feeder\_Type"}\NormalTok{)}

\NormalTok{FD\_Data}\SpecialCharTok{$}\NormalTok{Collection\_Date}\OtherTok{\textless{}{-}} \FunctionTok{as.numeric}\NormalTok{(FD\_Data}\SpecialCharTok{$}\NormalTok{Collection\_Date)}

\NormalTok{FD\_Data }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{finalfit}\NormalTok{(dependent\_crr, explanatory)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
 Dependent: Surv(Collection_Date, salmonella_crr)                      all
                                      Feeder_Type        Plastic 36 (25.0)
                                                  Plastic_coated 36 (25.0)
                                                            Wood 36 (25.0)
                                                     Wood_coated 36 (25.0)
          HR (univariable)        HR (multivariable)
                         -                         -
 1.00 (0.63-1.59, p=1.000) 1.00 (0.63-1.59, p=1.000)
 1.00 (0.63-1.59, p=1.000) 1.00 (0.63-1.59, p=1.000)
 1.00 (0.63-1.59, p=1.000) 1.00 (0.63-1.59, p=1.000)
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{FD\_Data }\SpecialCharTok{\%\textgreater{}\%} 
   \FunctionTok{finalfit}\NormalTok{(dependent\_crr, explanatory, }\AttributeTok{add\_dependent\_label =} \ConstantTok{FALSE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
    \FunctionTok{rename}\NormalTok{(}\StringTok{"Overall survival"} \OtherTok{=}\NormalTok{ label) }\SpecialCharTok{\%\textgreater{}\%} 
    \FunctionTok{rename}\NormalTok{(}\StringTok{" "} \OtherTok{=}\NormalTok{ levels) }\SpecialCharTok{\%\textgreater{}\%} 
    \FunctionTok{rename}\NormalTok{(}\StringTok{"  "} \OtherTok{=}\NormalTok{ all) }
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
 Overall survival                                   HR (univariable)
      Feeder_Type        Plastic 36 (25.0)                         -
                  Plastic_coated 36 (25.0) 1.00 (0.63-1.59, p=1.000)
                            Wood 36 (25.0) 1.00 (0.63-1.59, p=1.000)
                     Wood_coated 36 (25.0) 1.00 (0.63-1.59, p=1.000)
        HR (multivariable)
                         -
 1.00 (0.63-1.59, p=1.000)
 1.00 (0.63-1.59, p=1.000)
 1.00 (0.63-1.59, p=1.000)
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{survfit}\NormalTok{(}\FunctionTok{Surv}\NormalTok{(Collection\_Date, Absence\_0\_Presence\_1)}\SpecialCharTok{\textasciitilde{}}\DecValTok{1}\NormalTok{, }\AttributeTok{data=}\NormalTok{FD\_Data)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
Call: survfit(formula = Surv(Collection_Date, Absence_0_Presence_1) ~ 
    1, data = FD_Data)

       n events median 0.95LCL 0.95UCL
[1,] 144     21     NA      NA      NA
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{FD\_MedSurv}\OtherTok{\textless{}{-}}\NormalTok{FD\_Data }\SpecialCharTok{\%\textgreater{}\%}
 \FunctionTok{filter}\NormalTok{(salmonella\_crr}\SpecialCharTok{==}\DecValTok{1}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{summarize}\NormalTok{(}\AttributeTok{median\_surv=}\FunctionTok{median}\NormalTok{(Collection\_Date))}

\NormalTok{FD\_MedSurv }\OtherTok{\textless{}{-}}\NormalTok{ ((FD\_MedSurv)}\SpecialCharTok{/}\DecValTok{604800}\NormalTok{)}\SpecialCharTok{*}\DecValTok{100} \CommentTok{\#Persistence seems to be roughly 3 days for the FD}

\NormalTok{FD\_Surv\_Table}\OtherTok{\textless{}{-}} \FunctionTok{gt}\NormalTok{(FD\_MedSurv) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{tab\_header}\NormalTok{(}
    \AttributeTok{title =} \StringTok{"Median Survival of Salmonella"}\NormalTok{,}
    \AttributeTok{subtitle =}\NormalTok{ (}\StringTok{"02/22/2022 to 03/14/2022"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{cols\_label}\NormalTok{(}
    \AttributeTok{median\_surv=} \StringTok{"Median Salmonella Survival (days)"}\NormalTok{) }

\NormalTok{FD\_Surv\_Table }\CommentTok{\# seems to be roughly 3 days}
\end{Highlighting}
\end{Shaded}

\begin{longtable}{r}
\caption*{
{\large Median Survival of Salmonella} \\ 
{\small 02/22/2022 to 03/14/2022}
} \\ 
\toprule
Median Salmonella Survival (days) \\ 
\midrule
3.149884 \\ 
\bottomrule
\end{longtable}

\textbf{Cox regression: Picnic Table Trial}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{PT\_Data}\SpecialCharTok{$}\NormalTok{Date}\OtherTok{\textless{}{-}} \FunctionTok{as.numeric}\NormalTok{(PT\_Data}\SpecialCharTok{$}\NormalTok{Date)}

\FunctionTok{coxph}\NormalTok{(}\FunctionTok{Surv}\NormalTok{(Date, Salmonella\_Positive) }\SpecialCharTok{\textasciitilde{}}\NormalTok{ Sample\_Type, PT\_Data)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
Call:
coxph(formula = Surv(Date, Salmonella_Positive) ~ Sample_Type, 
    data = PT_Data)

                    coef exp(coef) se(coef)     z      p
Sample_TypePooled 0.8344    2.3034   0.3470 2.405 0.0162

Likelihood ratio test=6.28  on 1 df, p=0.01224
n= 192, number of events= 39 
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{CR\_PT1}\OtherTok{\textless{}{-}}\FunctionTok{coxph}\NormalTok{(}\FunctionTok{Surv}\NormalTok{(Date, Salmonella\_Positive) }\SpecialCharTok{\textasciitilde{}}\NormalTok{ Sample\_Type, PT\_Data) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{tbl\_regression}\NormalTok{(}\AttributeTok{exp=}\ConstantTok{TRUE}\NormalTok{)}

\NormalTok{CR\_PT1}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
Table printed with `knitr::kable()`, not {gt}. Learn why at
https://www.danieldsjoberg.com/gtsummary/articles/rmarkdown.html
To suppress this message, include `message = FALSE` in code chunk header.
\end{verbatim}

\begin{longtable}[]{@{}lccc@{}}
\toprule()
\textbf{Characteristic} & \textbf{HR} & \textbf{95\% CI} &
\textbf{p-value} \\
\midrule()
\endhead
Sample\_Type & & & \\
Persistence & --- & --- & \\
Pooled & 2.30 & 1.17, 4.55 & 0.016 \\
\bottomrule()
\end{longtable}

\textbf{Cox regression: Feeder Trial}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{FD\_Data}\SpecialCharTok{$}\NormalTok{Collection\_Date}\OtherTok{\textless{}{-}} \FunctionTok{as.numeric}\NormalTok{(FD\_Data}\SpecialCharTok{$}\NormalTok{Collection\_Date)}

\FunctionTok{coxph}\NormalTok{(}\FunctionTok{Surv}\NormalTok{(Collection\_Date, Absence\_0\_Presence\_1) }\SpecialCharTok{\textasciitilde{}}\NormalTok{ Feeder\_Type, FD\_Data)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
Call:
coxph(formula = Surv(Collection_Date, Absence_0_Presence_1) ~ 
    Feeder_Type, data = FD_Data)

                             coef exp(coef) se(coef)      z     p
Feeder_TypePlastic_coated  0.2918    1.3389   0.5401  0.540 0.589
Feeder_TypeWood           -0.4111    0.6629   0.6455 -0.637 0.524
Feeder_TypeWood_coated    -0.6999    0.4966   0.7071 -0.990 0.322

Likelihood ratio test=2.88  on 3 df, p=0.4112
n= 144, number of events= 21 
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{CR\_FD1}\OtherTok{\textless{}{-}} \FunctionTok{coxph}\NormalTok{(}\FunctionTok{Surv}\NormalTok{(Collection\_Date, Absence\_0\_Presence\_1) }\SpecialCharTok{\textasciitilde{}}\NormalTok{ Feeder\_Type, FD\_Data) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{tbl\_regression}\NormalTok{(}\AttributeTok{exp=}\ConstantTok{TRUE}\NormalTok{)}

\NormalTok{CR\_FD1}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
Table printed with `knitr::kable()`, not {gt}. Learn why at
https://www.danieldsjoberg.com/gtsummary/articles/rmarkdown.html
To suppress this message, include `message = FALSE` in code chunk header.
\end{verbatim}

\begin{longtable}[]{@{}lccc@{}}
\toprule()
\textbf{Characteristic} & \textbf{HR} & \textbf{95\% CI} &
\textbf{p-value} \\
\midrule()
\endhead
Feeder\_Type & & & \\
Plastic & --- & --- & \\
Plastic\_coated & 1.34 & 0.46, 3.86 & 0.6 \\
Wood & 0.66 & 0.19, 2.35 & 0.5 \\
Wood\_coated & 0.50 & 0.12, 1.99 & 0.3 \\
\bottomrule()
\end{longtable}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{CR\_FD1}\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{saveRDS}\NormalTok{(}\FunctionTok{here}\NormalTok{(}\StringTok{"4. Results"}\NormalTok{, }\StringTok{"FD\_Cox.rds"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\end{document}
